---
title: "Coral Growth and Symbionts Analysis"
author: "Callie Stephenson"
date: "2023-08-29"
output: github_document
---
## Introduction

This is a R Markdown file in which I hope to write out my analyses. I will use this document to process all the R scripts found in the R folder and complete my project. 

#### Loading the data
```{r load packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(fishualize)
library(lme4)
library(lmerTest) #figure out which lme package you need
library(ggpp)
library(ggplot2)
library(GGally)
library(ggeffects)
library(arm) #discrete hist
library(MuMIn) #dredge
library(effects) #effects plots for dredged models
library(curl) #curl
library(gridExtra)
```


```{r load data, echo=FALSE, message=FALSE}
#explanatory data
all_nut <- read.csv("data/All_Nutrients_Processed.csv")
nut <- read.csv("data/March_nutrients_processed.csv")
nut_wide <- read.csv("data/Nutrient_data_wide.csv")
explanatory_seasonal <- read.csv("data/Explanatory_variables.csv")
turb <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Biogeochem/August2021/Turb_NC.csv")

#explanatory 2
cv_low_wet <- read.csv("data/Wet_Nutrients_Processed_CV_Low.csv")
cv_low_all <- read.csv("data/All_Nutrients_Processed_CV_Low.csv")

#Response data
sa <- read.csv("data/Surface_Area.csv")
BW <- read.csv("data/BW_tidy.csv") 
BW <- subset(BW, select=-c(1,4,5)) #remove some erroneous columns
TLE <- read.csv("data/TLE_summary.csv")
FCM <- read.csv("data/FCM_tidy.csv") %>% 
  rename(Placement_Code = PLACEMENT)
SI <- read.csv("data/SI_Tidy.csv")

#metadata
meta <- read.csv("data/coral_metadata.csv")

#Mean Low Tide Value PCA 1 Axis
PCA_Pulse_variables_wide <- read.csv("data/PCA_Pulse_variables_wide.csv")
#explan <- read.csv("data/Explanatory_variables_wide.csv")
```

```{r clean turb}
turb1 <- turb %>% 
  filter(startsWith(CowTagID, "V")& CowTagID != "VSEEP")

turb1 <- turb1 %>%
  dplyr::select(c(CowTagID,
                   N_percent,
                   C_N))
#use C:N 
#can also use N percent
#plot(turb$C_N, turb$N_percent)
#hist(turb$C_N)
#hist(turb$N_percent)
#N_percent looks better. Will use N_percent for models
```

```{r build explanatory}
all_nut_long <- pivot_longer(
  data = all_nut,
  cols = c(Minimum_Salinity:Low_Tide_Mean_Ammonia_umolL)) 
names(all_nut_long)[names(all_nut_long) == "name"] <- "Parameter"

join <- explanatory_seasonal[,c(2:7)]
join <- unique(join)

explanatory_all <- left_join(join,all_nut_long, by = join_by(CowTagID))
explanatory_all_wide <- left_join(join, all_nut,by = join_by(CowTagID))
explanatory_all_wide <- left_join(explanatory_all_wide, turb1,by = join_by(CowTagID))
explanatory_all_wide <- left_join(explanatory_all_wide, PCA_Pulse_variables_wide,by = join_by(CowTagID))
```


```{r build response, echo=FALSE}
#make a master response variable data sheet
response_data <- left_join(meta[,c(2:4,7,8)], BW[,c(2,4,5,6)], by = join_by(Placement_Code)) %>% 
  #select(!X) %>% #check that this doesn't need to be here
  left_join(FCM[,c(1,16,17)], by = join_by(Placement_Code)) %>% 
  left_join(TLE[,c(3:7)], by = join_by(Placement_Code))
response_data$CowTagID <- paste0("V",response_data$Pin_Number)

PRU_response_data <- response_data %>% 
  filter(Species == "Porites rus")

PRU_response_data_caged <- PRU_response_data %>% 
  filter(Cage_Uncaged == "C")

PAC_response_data <- response_data %>% 
  filter(Species == "Pocillopora acuta")

PAC_response_data_caged <- PAC_response_data %>% 
  filter(Cage_Uncaged == "C")
```


```{r all data seasonal, warning=FALSE}
#now make an all data that has both explanatory and response variables:
all_data_seasonal <- left_join(response_data, explanatory_seasonal, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
all_data_seasonal$Genotype <- as.factor(all_data_seasonal$Genotype)
all_PAC_data_seasonal <- all_data_seasonal %>% 
  filter(Species == "Pocillopora acuta")
```

```{r all data both seasons, warning=FALSE}
#now make an all data that has both explanatory and response variables:
all_data <- left_join(response_data, explanatory_all, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
all_data$Genotype <- as.factor(all_data$Genotype)
all_PAC_data <- all_data %>% 
  filter(Species == "Pocillopora acuta")

all_data_wide <- left_join(response_data, explanatory_all_wide, by = join_by(CowTagID))
```


```{r columns to dredge}
minimum_columns <- names(all_data_wide)[grepl("Minimum_", names(all_data_wide)) & 
                                     grepl("Salinity|Temperature|pH", names(all_data_wide))]
#LOOK OUT YOU CHANGED THIS ON APRIL 17
maximum_columns <- names(all_data_wide)[grepl("Maximum_", names(all_data_wide)) & 
                                  !grepl("Salinity|Temperature|Ammonia_umol",names(all_data_wide))]
#mean_columns <- names(all_data_wide)[grepl("Mean_", names(all_data_wide)) & 
#                                  !grepl("Salinity|Temperature|pH|Ammonia_umol",names(all_data_wide))]
CV_columns <- names(all_data_wide)[grepl("CV_", names(all_data_wide), names(all_data_wide))]

mean_low_columns <- names(all_data_wide)[grepl("Low_Tide_Mean_", names(all_data_wide), names(all_data_wide))]

#PAC_model_data$CV_Salinity_scaled <- scale(PAC_model_data$CV_Salinity, TRUE, TRUE)
#PAC_model_data$CV_Temperature_scaled  <-  scale(PAC_model_data$CV_Temperature, TRUE, TRUE)
#PAC_model_data$CV_TA_scaled  <-  scale(PAC_model_data$CV_TA, TRUE, TRUE)
#PAC_model_data$CV_pH_scaled  <-  scale(PAC_model_data$CV_pH, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_Phosphate_umolL_scaled<-
#  scale(PAC_model_data$Low_Tide_Mean_Phosphate_umolL, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_NN_umolL_scaled  <-  
#  scale(PAC_model_data$Low_Tide_Mean_NN_umolL, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_Ammonia_umolL_scaled  <-
#  scale(PAC_model_data$Low_Tide_Mean_Ammonia_umolL, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_NNP_umolL_scaled <- 
#  scale(PAC_model_data$Low_Tide_Mean_Ammonia_umolL, TRUE, TRUE)


for (col in CV_columns) {
  scaled_variable <- paste0(col, "_scaled")
  scaled_values <- scale(all_data_wide[[col]], TRUE, TRUE)
  all_data_wide[[scaled_variable]] <- scaled_values[, 1]
}

for (col in mean_low_columns) {
  scaled_variable <- paste0(col, "_scaled")
  scaled_values <- scale(all_data_wide[[col]], TRUE, TRUE)
  all_data_wide[[scaled_variable]] <- scaled_values[, 1]
}

for (col in minimum_columns) {
  scaled_variable <- paste0(col, "_scaled")
  scaled_values <- scale(all_data_wide[[col]], TRUE, TRUE)
  all_data_wide[[scaled_variable]] <- scaled_values[, 1]
}

for (col in maximum_columns) {
  scaled_variable <- paste0(col, "_scaled")
  scaled_values <- scale(all_data_wide[[col]], TRUE, TRUE)
  all_data_wide[[scaled_variable]] <- scaled_values[, 1]
}

scaled_cv_columns <- names(all_data_wide)[grepl("^CV.*_scaled$", names(all_data_wide))]
scaled_ml_columns <- names(all_data_wide)[grepl("^Low.*_scaled$", names(all_data_wide))]
scaled_min_columns <- names(all_data_wide)[grepl("^Minimum.*_scaled$", names(all_data_wide))]
scaled_max_columns <- names(all_data_wide)[grepl("^Maximum.*_scaled$", names(all_data_wide))]

# Extract column names for filtering
selected_columns <- c("FSC.Events_per_cm_2", "Percent_Change", "Genotype", "Cage_Uncaged", "Species",
                      minimum_columns, 
                      maximum_columns,
                      CV_columns,
                      mean_low_columns,
                      scaled_min_columns, 
                      scaled_max_columns,
                      scaled_cv_columns,
                      scaled_ml_columns)

model_data <- all_data_wide[, selected_columns]


PAC_model_data <- na.omit(model_data %>% 
                            filter(Species == "Pocillopora acuta")) #%>%
# filter(Cage_Uncaged == "C")

PRU_model_data <- na.omit(model_data %>% 
        filter(Species == "Porites rus")) # %>% 
#  filter(Cage_Uncaged == "C")

dredge.data <- na.omit(model_data[, c(maximum_columns, 
                                    minimum_columns #,
#                                    mean_columns, #removed bc not enough data for the amount of variables
#                                    CV_columns,
#                      CV_columns,
#                      mean_low_columns,
#                      scaled_min_columns, 
#                      scaled_max_columns,
#                      scaled_cv_columns,
#                      scaled_ml_columns
)])
```

## Data Exploration:
Data exploration focuses on the following points:
1. Outliers
2. Collinearity
3. Independence (Relationships between response and explanatory variables) 

```{r response data hist, fig.width=10, fig.height=10, echo=FALSE}
#png("output/response_data_hist.png", width = 1100, height = 1100)
par(mfrow = c(4,4), oma=c(1,1,3,1))
hist(PAC_response_data$Change_Over_Area, main="Change Over Area", col="lightblue", xlab="Change_Over_Area")
hist_columns <- c(7:13)
for (i in 1:length(hist_columns)) {
  hist(PAC_response_data[, hist_columns[i]], col="lightblue", 
       main = paste("PAC", colnames(PAC_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
hist(PRU_response_data$Change_Over_Area, main="Change Over Area", col="darkolivegreen3", xlab="Change_Over_Area")
for (i in 1:length(hist_columns)) {
  hist(PRU_response_data[, hist_columns[i]], col="darkolivegreen3", 
       main = paste("PRU", colnames(PRU_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
main_title <- "Response Data Exploration"
title(main = main_title, outer = TRUE)
#dev.off()
```

Ok so FSC.Events are both pretty right-skewed. CN says log10 transform these.

```{r wet season raw growth data plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
#  ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
#         plot = plot, 
#         width = 15, height = 15)
}
```

Now, doing this again but for both seasons combined:
```{r both season raw growth data plots, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```

Same but let's do for FCM data

```{r wet season raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = FSC.Events_per_cm_2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
#  ggsave(filename = paste("output/wet_raw_symbiong_count_response_plot_", sp, ".png", sep = ""),
#         plot = plot, 
#         width = 15, height = 15)
}
```

And again, now for both seasons:
```{r both seasons raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = log10(FSC.Events_per_cm_2))) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```



```{r growth change by symbionts}
response_data %>% 
  filter(Cage_Uncaged == "C") %>% 
  ggplot(aes(y = Percent_Change, x = log10(FSC.Events_per_cm_2)))+
  geom_point() +
  geom_smooth(method = "lm", formula = "y~poly(x,2)") +
  ggtitle("Change in BW by Number of Symbionts") +
  geom_text(aes(label = Pin_Number)) +
  facet_wrap(~ Species, scales = "free") -> growth_v_symb_plot

growth_v_symb_plot
#ggsave("output/caged_growth_v_symb.png", width = 9, height = 6, plot = growth_v_symb_plot)


symb_growth_model <- response_data %>%
  filter(Cage_Uncaged == "C") %>%
  group_by(Species) %>% 
  do(model = lmer(Percent_Change ~ log10(FSC.Events_per_cm_2) + (1|Genotype), data = .))

# Extract model objects and print the anovas
model_objects <- symb_growth_model$model
anova_tables <- lapply(model_objects, function(model) {
  car::Anova(model)
})
anova_tables
#car::Anova(symb_growth_model$model[[1]]) #less code with small number of models
car::Anova(symb_growth_model$model[[2]])
```

OK, I don't want this relationship driven by 14 for rus, which I know had a good B coral and a diseased C coral

```{r}
response_data_no_14C_14B <- response_data %>%
  filter(Species == "Porites rus" & 
         (Cage_Uncaged == "C" | (Cage_Uncaged == "B" & Pin_Number == 14)))%>% 
  filter(!(Pin_Number == 14 & Cage_Uncaged == "C"))

response_data_no_14C <- response_data %>%
  filter(Species == "Porites rus" & 
         (Cage_Uncaged == "C")) %>%
  filter(!(Pin_Number == 14 & Cage_Uncaged == "C"))

model = lmer(Percent_Change ~ log10(FSC.Events_per_cm_2) + (1|Genotype), data = response_data_no_14C_14B)
car::Anova(model)

plot1 <- response_data_no_14C_14B %>% ggplot(aes(y = Percent_Change, x = log10(FSC.Events_per_cm_2)))+
  geom_point() +
  geom_smooth(method = "lm", formula = "y~poly(x,2)") +
  ggtitle("BW by Number of Symbionts no 14C with 14B")

plot2 <- response_data_no_14C %>% ggplot(aes(y = Percent_Change, x = log10(FSC.Events_per_cm_2)))+
  geom_point() +
  geom_smooth(method = "lm", formula = "y~poly(x,2)") +
  ggtitle("BW by Number of Symbionts no 14C")

grid.arrange(plot1, plot2, ncol = 2)

```




```{r bw change in rus vs acuta}
BW_long <- response_data %>% 
  filter(Cage_Uncaged == "C") %>%
  dplyr::select(Species, Percent_Change,Pin_Number ) %>%
  group_by(Species, Pin_Number)%>%
  summarise(Percent_ChangeM  = mean(Percent_Change, na.rm = TRUE)) %>% 
  pivot_wider(names_from = "Species",
              values_from = "Percent_ChangeM")

growth_v_sp_plot <- ggplot(data = BW_long, aes(x = `Pocillopora acuta`, y = `Porites rus`))+
  geom_point() +
  labs(title = "% ∆ BW between species by Pin Number")
#  geom_smooth(method = "lm") +
  
#ggsave("output/caged_growth_v_sp_plot.png", width = 9, height = 6, plot = growth_v_sp_plot)
```


#### Look for Outliers:
Buoyant Weight
It looks like that 13 A is really off. We know it broke, so I am removing it and replotted next to it:
```{r BW Outlier Tidy, echo=FALSE}
# BW Plot
BW <- merge(BW, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
BW_no <- BW[!(BW$Placement_Code == "PRU V13 A"), ]
```

Now making a plot to look at outliers:
```{r BW Outlier Plot, echo=FALSE}
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
#dotchart(BW$Change_Over_Area, main = "∆BW", 
#         groups = BW$Cage_Uncaged)
#dotchart(BW_no$Change_Over_Area, main = "∆BW removed erroneous data point", 
#         groups = BW_no$Cage_Uncaged)
layout(1)
```
I removed PRU V13 A placement because this was broken in the field, creating a large negative change in buoyant weight. 

TLE
```{r TLE Plot, echo=FALSE}
TLE <- merge(TLE, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
layout(matrix(c(1, 2, 3, 4), 2, 2, byrow = TRUE)) # Setting up the layout
#dotchart(na.omit(TLE$Max_L_AVG), main = "Max L", groups = na.omit(TLE$Cage_Uncaged)) # Side Max L Plot
#dotchart(TLE$Max_L_Basal, main = "Max L Basal", groups = na.omit(TLE$Cage_Uncaged)) # Basal Max L plot
#dotchart(TLE$Area_AVG, main = "Area", groups = na.omit(TLE$Cage_Uncaged)) # Area plot
#dotchart(TLE$Area_Basal, main = "Basal Area", groups = na.omit(TLE$Cage_Uncaged)) # Basal Area plot
layout(1) # Reset the layout
#title("TLE Outlier Dotchart", line = -1, outer = TRUE, cex.main = 1.5)
```

### Symbionts

```{r Sym plot, echo=FALSE}
FCM <- FCM[complete.cases(FCM[, "Cage_Uncaged"]), ] #Remove T0 samples by only including rows where the "Cage_Uncaged" exists
FCM$Cage_Uncaged <- as.factor(FCM$Cage_Uncaged) 
FCM_no <- FCM[!(FCM$Placement_Code == "PRU V17 C"), ]
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
dotchart(log10(FCM$FSC.Events_per_cm_2), main = "Symbionts", groups = FCM$Cage_Uncaged)
dotchart(log10(FCM_no$FSC.Events_per_cm_2), main = "Symbionts without erroneous point", groups = FCM_no$Cage_Uncaged)
layout(1)
```
Jess will rerun PRU 49 (PRU V17 C) with her samples - I likely added twice as much as needed to the sample, but for now I will omit in my analyses (using dataframe FCM_no for "no outliers")

```{r FCM Violin Plot}
FCM_violin <- ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_cm_2), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
  geom_text(aes(label = Pin_Number)) +
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1)+
  facet_wrap(~Species)+
  labs(title = expression(paste("Symbionts per ", cm^2, " by caging treatment")))+
  theme(plot.title = element_text(hjust = 0.5))

FCM_violin
#ggsave("output/FCM_violin.png", width = 9, height = 6, plot = FCM_violin)
```

same but other normalization:

```{r FCM Violin Plot Dry Weight}
ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_g_dry_weight), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1) +
  facet_wrap(~Species)+
  labs(title = "Symbionts per g dry weight by caging treatment")+
  theme(plot.title = element_text(hjust = 0.5))
```

### Collinearity
need to make a dataframe that's long that has all the nutrient metrics and independent variables
Here's the thing: we KNOW the nutrients are collinear
For example, here's a pairs plot for the maximum of all the nutrient values:

```{r pairs plot collinearity}
nut_wide_no_seep<- nut_wide[nut_wide$CowTagID != 'VSEEP', ]
maximum_columns <- names(nut_wide)[grepl("Maximum_", names(nut_wide)) & !grepl("Salinity|Temperature|pH|Ammonia_umol", names(nut_wide))]

ggpairs(nut_wide_no_seep[, c(maximum_columns)],progress = FALSE)
```

### Independence
```{r independence}

```

### Data Analysis


FCM Data Analysis:
```{r FCM vis, echo=FALSE}
anova_PRU <- aov(FSC.Events_per_cm_2 ~ Cage_Uncaged, 
                 data = FCM %>% 
                   filter(Species=="Porites_rus"))
summary(anova_PRU)

#caging did not effect P. rus

anova_PAC <- aov(FSC.Events_per_cm_2 ~ Cage_Uncaged, 
                 data = FCM %>% 
                   filter(Species=="Pocillopora_acuta"))
summary(anova_PAC)
#caging did not effect symbionts in P. acuta
#this I have a bit of a hard time believing 
```

### UNIVARIATE Models
```{r CALLIE THIS IS YOUR LOOPING CODE Singular model looping, warning=FALSE}
#make the PAC data frame longer
PAC_model_scaled_data_long <- pivot_longer(
  data = PAC_model_data[,c(1:5,32:57)],
  cols = c(Minimum_Salinity_scaled:Low_Tide_Mean_Ammonia_umolL_scaled)) 
names(PAC_model_scaled_data_long)[names(PAC_model_scaled_data_long) == "name"] <- "Parameter"

PAC_model_data_long <- pivot_longer(
  data = PAC_model_data[,c(1:5,6:31)],
  cols = c(Minimum_Salinity:Low_Tide_Mean_Ammonia_umolL)) 
names(PAC_model_data_long)[names(PAC_model_data_long) == "name"] <- "Parameter"


#make the PRU data frame longer
PRU_model_scaled_data_long <- pivot_longer(
  data = PRU_model_data[,c(1:5,32:57)],
  cols = c(Minimum_Salinity_scaled:Low_Tide_Mean_Ammonia_umolL_scaled)) 
names(PRU_model_scaled_data_long)[names(PRU_model_scaled_data_long) == "name"] <- "Parameter"

```

#### Pac growth linear
```{r PAC linear growth model all scaled parameteres against}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)

plotlist <- list()

parameters <- unique(PAC_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PAC_model_scaled_data_long, Parameter == parameter)
  
  # Fit the mixed-effects model
  model <- lmer(Percent_Change ~ value + (1|Genotype) + (1|Cage_Uncaged), data = subset_data)
  
  # Get AICc and p-value
    AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["value", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["value", "Estimate"]
  
  # Calculate marginal and conditional R-squared values
    rsq <- as.numeric(r.squaredGLMM(model))
    marginal_R2 <- rsq[1]
    conditional_R2 <- rsq[2]
  
  if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate,
                                                     Marginal_R2 = marginal_R2,
                                                     Conditional_R2 = conditional_R2))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value[all]")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }
print(model_results)
grid.arrange(grobs = plotlist, ncol = 2)

model_results$Species <- "Pocillopora acuta"
model_results$type <- "linear"

PAC_growth_model_results <- model_results

```

#### Pac growth polynomial
```{r PAC polynomial model all scaled parameteres against growth}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)
plotlist <- list()

parameters <- unique(PAC_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PAC_model_scaled_data_long, Parameter == parameter)
  
  # Fit the mixed-effects model
  model <- lmer(Percent_Change ~ poly(value,2) + (1|Genotype) + (1|Cage_Uncaged), data = subset_data)
  
  # Get AICc and p-value
  AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["poly(value, 2)1", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["poly(value, 2)1", "Estimate"]
  
  # Calculate marginal and conditional R-squared values
    rsq <- as.numeric(r.squaredGLMM(model))
    marginal_R2 <- rsq[1]
    conditional_R2 <- rsq[2]
  
  if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate,
                                                     Marginal_R2 = marginal_R2,
                                                     Conditional_R2 = conditional_R2))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value[all]")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }

model_results %>% filter(p.value < 0.05)
grid.arrange(grobs = plotlist, ncol = 2)

model_results$Species <- "Pocillopora acuta"
model_results$type <- "polynomial"

PAC_growth_model_results <- rbind(model_results, PAC_growth_model_results)
```

#### PAC growth results
```{r}
PAC_growth_model_results[order(PAC_growth_model_results$AICc), c(1, 8, 2:7)]
```

```{r PAC growth multivariate dredge}
PAC_growth_big_model <- lmer(Percent_Change ~ 
                       CV_pH_scaled +
                        Minimum_pH_scaled +
                         CV_TA_scaled +
                         CV_NN_umolL_scaled +
                         Maximum_TA_scaled +
                       (1 | Genotype)+ (1|Cage_Uncaged), 
                                 data = PAC_model_data, 
                                 na.action = na.pass)

PAC_growth_big_model_dredge <- dredge(PAC_growth_big_model, trace = 2)
#model.sel(PAC_cv_low_growth_model_dredge)
PAC_growth.avg = model.avg(PAC_growth_big_model_dredge, delta < 4)
summary(PAC_growth.avg)
PAC_growth.avg$sw
```


## PAC GROWTH FINAL:
```{r}
FINAL_PAC_growth <- subset(PAC_growth_model_results, abs(AICc - min(AICc)) < 2)

FINAL_PAC_growth
```


#### Pru growth linear

```{r PRU all variables scaled against growth, message=FALSE}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)
plotlist <- list()

parameters <- unique(PRU_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PRU_model_scaled_data_long, Parameter == parameter)
  
  # Fit the mixed-effects model
  model <- lmer(Percent_Change ~ value + (1|Genotype) + (1|Cage_Uncaged), data = subset_data)
  
  # Get AICc
  AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["value", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["value", "Estimate"]
  
 if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value[all]")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }
model_results %>% filter(p.value < 0.05)
grid.arrange(grobs = plotlist, ncol = 2)

model_results$Species <- "Porites rus"
model_results$type <- "linear"

PRU_growth_model_results <- model_results
```

#### Pru growth polynomial

```{r PRU polynomial all variables scaled against growth, message=FALSE}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)
plotlist <- list()

parameters <- unique(PRU_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PRU_model_scaled_data_long, Parameter == parameter)
  
  # Fit the mixed-effects model
  model <- lmer(Percent_Change ~ poly(value,2) + (1|Genotype) + (1|Cage_Uncaged), data = subset_data)
  
  # Get AICc and p-value
  AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["poly(value, 2)1", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["poly(value, 2)1", "Estimate"]
  
  if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }

model_results %>% filter(p.value < 0.05)
grid.arrange(grobs = plotlist, ncol = 2)

model_results$Species <- "Porites rus"
model_results$type <- "polynomial"

PRU_growth_model_results <- rbind(model_results,PRU_growth_model_results)
```

#### PRU growth results
```{r}
PRU_growth_model_results[order(PRU_growth_model_results$AICc), c(1, 6, 2:5)]
```

```{r multivariate dredge}
PRU_growth_big_model <- lmer(Percent_Change ~ 
                       Low_Tide_Mean_Salinity_scaled + 
                       Low_Tide_Mean_Temperature_scaled +
                        CV_Salinity_scaled +
                       (1 | Genotype)+ (1|Cage_Uncaged), 
                                 data = PRU_model_data, 
                                 na.action = na.pass)

PRU_growth_big_model_dredge <- dredge(PRU_growth_big_model, trace = 2)
#model.sel(PAC_cv_low_growth_model_dredge)
PRU_growth.avg = model.avg(PRU_growth_big_model_dredge, delta < 10)
summary(PRU_growth.avg)
PRU_growth.avg$sw
```

#### Pac symb linear

```{r model all parameteres scaled against symbiont, warning=FALSE, message=FALSE}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)

plotlist <- list()

parameters <- unique(PAC_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PAC_model_scaled_data_long, Parameter == parameter)
  
  # Fit the mixed-effects model
  model <- lmer(log10(FSC.Events_per_cm_2) ~ value + (1|Genotype) + (1|Cage_Uncaged), data = subset_data)
  
  # Get AICc and p-value
    AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["value", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["value", "Estimate"]
  
if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value[all]")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }
model_results %>% filter(p.value < 0.05)
#grid.arrange(grobs = plotlist, ncol = 2)

#model_results$Species <- "Pocillopora acuta"
#model_results$type <- "linear"
PAC_symb_model_results <- model_results
```

#### Pac symb polynomial

```{r PAC polynomialmodel all parameteres scaled against symbiont, warning=FALSE, message=FALSE}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)
plotlist <- list()

parameters <- unique(PAC_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PAC_model_scaled_data_long, Parameter == parameter)

  # Fit the mixed-effects model
  model <- lmer(log10(FSC.Events_per_cm_2) ~ poly(value,2) + (1|Genotype) + (1|Cage_Uncaged), data = subset_data)
  
  # Get AICc and p-value
  AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["poly(value, 2)1", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["poly(value, 2)1", "Estimate"]
  
  if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }

model_results %>% filter(p.value < 0.05)
#grid.arrange(grobs = plotlist, ncol = 1)

#model_results$Species <- "Pocillopora acuta"
#model_results$type <- "polynomial"

PAC_symb_model_results <- rbind(model_results, PAC_symb_model_results)
```

#### Pac symb results

```{r}
PAC_symb_model_results
```



#### Pru symb linear

```{r PRU symbionts}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)
plotlist <- list()

parameters <- unique(PRU_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PRU_model_scaled_data_long, Parameter == parameter)
  
  # Fit the mixed-effects model
  model <- lmer(log10(FSC.Events_per_cm_2) ~ value + (1|Genotype) + Cage_Uncaged, data = subset_data)
  
  # Get AICc and p-value
    AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["value", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["value", "Estimate"]
  
 if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value[all]")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }
model_results %>% filter(p.value < 0.05)
grid.arrange(grobs = plotlist, ncol = 2)


model_results$Species <- "Porites rus"
model_results$type <- "linear"

PRU_symb_model_results <- model_results
```

#### Pru symb polynomial

```{r PRU symbionts polynomial, message=FALSE}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)

plotlist <- list()

parameters <- unique(PRU_model_scaled_data_long$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(PRU_model_scaled_data_long, Parameter == parameter)
  
  # Fit the mixed-effects model
  model <- lmer(log10(FSC.Events_per_cm_2) ~ poly(value,2) + (1|Genotype) + (1|Cage_Uncaged), data = subset_data)
  
  # Get AICc and p-value
  AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["poly(value, 2)1", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["poly(value, 2)1", "Estimate"]
  
  if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(model_results, data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }

model_results %>% filter(p.value < 0.05)
grid.arrange(grobs = plotlist, ncol = 1)

model_results$Species <- "Porites rus"
model_results$type <- "polynomial"

PRU_symb_model_results <- rbind(model_results, PRU_symb_model_results)
```

#### Pru symb results

```{r}
PRU_symb_model_results
```

```{r multivariate dredge}
PRU_symb_big_model <- lmer(log10(FSC.Events_per_cm_2) ~ 
                       Maximum_Phosphate_umolL_scaled + 
                       Low_Tide_Mean_Phosphate_umolL_scaled +
                       (1 | Genotype)+ (1|Cage_Uncaged), 
                                 data = PRU_model_data, 
                                 na.action = na.pass)

PRU_symb_big_model_dredge <- dredge(PRU_symb_big_model, trace = 2)
#model.sel(PAC_cv_low_growth_model_dredge)
PRU_symb.avg = model.avg(PRU_symb_big_model_dredge, delta < 2)
summary(PRU_symb.avg)
PRU_symb.avg$sw
```


## ISOTOPES

### Exploratory plots:
If you could plot their 15N and 13C host and symbiont values on 2 different plots that would be a good place to start. One plot for 15N and one for 13C and then facet by whatever your categories are (location on gradient?). Then we can look at how overlapped their isotopes are:

```{r}
CV <- names(explanatory_all_wide)[grepl("CV_", names(explanatory_all_wide), names(explanatory_all_wide))]
mean_low <- names(explanatory_all_wide)[grepl("Low_Tide_Mean_", names(explanatory_all_wide), names(explanatory_all_wide))]

filter_values <- c("Minimum_Salinity",
                   "Minimum_Temperature",
                   "Maximum_TA",
                   "Minimum_pH",
                   "Maximum_NN_umolL",
                   "Maximum_Phosphate_umolL",
                   "Maximum_NNP_umolL")

SI_data1 <- left_join(SI, explanatory_all,by = join_by(CowTagID))
SI_data_wide <- left_join(SI, explanatory_all_wide,by = join_by(CowTagID))

# Filter the dataframe SI_data
SI_data <- SI_data1[SI_data1$Parameter %in% filter_values, ] 
```


```{r 15N host, fig.width=15, fig.height=15}
# species <- unique(SI_data$Species) # has NA
species <- c("Pocillopora acuta", "Porites rus" )
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- SI_data %>%
    filter(Species == sp, HS == "host")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = δ15N)) +
    geom_point(aes(shape = Genotype)) +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```

```{r EXPLORATORY 13C host, fig.width=15, fig.height=15}
species <- c("Pocillopora acuta", "Porites rus" )
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- SI_data %>%
    filter(Species == sp, HS == "host")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = δ13C)) +
    geom_point(aes(shape = Genotype)) +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```

```{r EXPLORATORY 13C symb, fig.width=15, fig.height=15}
species <- c("Pocillopora acuta", "Porites rus" )
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- SI_data %>%
    filter(Species == sp, HS == "symb")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = δ13C)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```


```{r EXPLORATORY 15N symb, fig.width=15, fig.height=15}
species <- c("Pocillopora acuta", "Porites rus" )
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- SI_data %>%
    filter(Species == sp, HS == "symb")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = δ15N)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```


# YOUR DATA FRAME IS REPEATING - GO FIX THIS

```{r pc plots, warning=FALSE, message=FALSE, fig.width=15, fig.height=8}
create_plot <- function(data, species, HS, y_variable, title_suffix) {
  data %>%
    filter(Species == species, HS == HS) %>% 
    ggplot(aes(x = pulse_pc1, y = !!sym(y_variable))) +
    geom_point(aes(shape = Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle(paste(species, HS, title_suffix)) +  
    theme_bw() +
    theme(legend.position = "none")
}

plot1 <- create_plot(SI_data_wide, "Porites rus", "symb", "δ15N", "δ15N")
plot2 <- create_plot(SI_data_wide, "Porites rus", "host", "δ15N", "δ15N")
plot3 <- create_plot(SI_data_wide, "Porites rus", "host", "Δ15N", "minus symbiont, Δ15N")
plot4 <- create_plot(SI_data_wide, "Porites rus", "symb", "δ13C", "δ13C")
plot5 <- create_plot(SI_data_wide, "Porites rus", "host", "δ13C", "δ13C")
plot6 <- create_plot(SI_data_wide, "Porites rus", "host", "Δ13C", "minus symbiont, Δ13C")


legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(plot1 + theme(legend.position = "bottom"))), "guide-box")

rus_pc_plots <- grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, legend, ncol = 3)

#ggsave(filename = paste("output/exploratory_isotope_rus_pc_plots.png", sep = ""),
#         plot = rus_pc_plots, 
#         width = 15, height = 8)
```

```{r acuta pc plots, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
create_SI_plot <- function(data, species, HS, y_variable, title_suffix) {
  data %>%
    filter(Species == species, HS == HS) %>% 
    ggplot(aes(x = pulse_pc1, y = !!sym(y_variable))) +
    geom_point(aes(shape = Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle(paste(species, HS, title_suffix)) +  
    theme_bw() +
    theme(legend.position = "none")
}

plot1 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "symb", "δ15N", "δ15N")
plot2 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "δ15N", "δ15N")
plot3 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "Δ15N", "minus symbiont, Δ15N")
plot4 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "symb", "δ13C", "δ13C")
plot5 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "δ13C", "δ13C")
plot6 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "Δ13C", "minus symbiont, Δ13C")


legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(plot1 + theme(legend.position = "bottom"))), "guide-box")

acuta_pc_plots <- grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol = 3)

#ggsave(filename = paste("output/exploratory_isotope_acuta_pc_plots.png", sep = ""),
#         plot = acuta_pc_plots, 
#         width = 15, height = 8)

```



```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
plot1 <- SI_data_wide %>%
    filter(Species == "Porites rus", HS == "symb") %>% 
    ggplot(aes(x = pulse_pc1, y = δ15N)) +
    geom_point(aes(shape=Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle("Porites rus symb δ15N") +  #Add species name to the plot title
    theme_bw() +
  theme(legend.position = "none")

plot2 <- SI_data_wide %>%
    filter(Species == "Porites rus", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = δ15N)) +
    geom_point(aes(shape=Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle("Porites rus host δ15N") +  #Add species name to the plot title
    theme_bw()+
  theme(legend.position = "none")

plot3 <- SI_data_wide %>%
    filter(Species == "Porites rus", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = Δ15N)) +
    geom_point(aes(shape=Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle("Porites rus Δ15N") +  #Add species name to the plot title
    theme_bw()+
  theme(legend.position = "none")

plot4 <- SI_data_wide %>%
    filter(Species == "Pocillopora acuta", HS == "symb") %>% 
    ggplot(aes(x = pulse_pc1, y = δ15N)) +
    geom_point(aes(shape=Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle("Pocillopora acuta symb δ15N") +  #Add species name to the plot title
    theme_bw()+
  theme(legend.position = "none")

plot5 <- SI_data_wide %>%
    filter(Species == "Pocillopora acuta", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = δ15N)) +
    geom_point(aes(shape=Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle("Pocillopora acuta host δ15N") +  #Add species name to the plot title
    theme_bw()+
  theme(legend.position = "none")

plot6 <- SI_data_wide %>%
    filter(Species == "Pocillopora acuta", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = Δ15N)) +
    geom_point(aes(shape=Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle("Pocillopora acuta Δ15N") +  #Add species name to the plot title
    theme_bw()+
  theme(legend.position = "none")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol = 3)
```

```{r, eval=FALSE, include=FALSE}
plot1 <- SI_data_wide %>%
    filter(Species == "Porites rus", HS == "symb") %>% 
    ggplot(aes(x = pulse_pc1, y = δ13C)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle("Porites rus symb δ13C") +  #Add species name to the plot title
    theme_bw()

plot2 <- SI_data_wide %>%
    filter(Species == "Porites rus", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = δ13C)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle("Porites rus host δ13C") +  #Add species name to the plot title
    theme_bw()

plot3 <- SI_data_wide %>%
    filter(Species == "Porites rus", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = Δ13C)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle("Porites rus host δ13C") +  #Add species name to the plot title
    theme_bw()

plot4 <- SI_data_wide %>%
    filter(Species == "Pocillopora acuta", HS == "symb") %>% 
    ggplot(aes(x = pulse_pc1, y = δ13C)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle("Pocillopora acuta symb δ13C") +  #Add species name to the plot title
    theme_bw()

plot5 <- SI_data_wide %>%
    filter(Species == "Pocillopora acuta", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = δ13C)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle("Pocillopora acuta host δ13C") +  #Add species name to the plot title
    theme_bw()

plot6 <- SI_data_wide %>%
    filter(Species == "Pocillopora acuta", HS == "host") %>% 
    ggplot(aes(x = pulse_pc1, y = Δ13C)) +
    geom_point() +
    geom_smooth(method = "lm") +
    ggtitle("Porites rus host δ13C") +  #Add species name to the plot title
    theme_bw()

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol = 3)
```

### Exploring plots of symbionts and growth vs. isotope values
```{r}
SI_data_wide <- left_join(SI_data_wide, FCM[,c("Placement_Code", "FSC.Events_per_cm_2")])
SI_data_wide$log10symbcount <- log10(SI_data_wide$FSC.Events_per_cm_2)
```

If you remember, the P. acuta did not have anything affecting its symbionts, but rus did. 

```{r Acuta vs symbionts, fig.width=12, fig.height=6}

create_SI_plot <- function(data, species, HS, x_variable, y_variable, title_suffix) {
  data %>%
    filter(Species == species, HS == HS) %>% 
    ggplot(aes(x = !!sym(x_variable), y = !!sym(y_variable))) +
    geom_point(aes(shape = Genotype)) +
    geom_smooth(method = "lm") +
    ggtitle(paste(species, HS, title_suffix)) +  
    theme_bw() +
    theme(legend.position = "none")
}

plot1 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "symb", "log10symbcount","δ15N", "δ15N")
plot2 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "log10symbcount", "δ15N", "δ15N")
plot3 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "log10symbcount", "Δ15N", "minus symbiont, Δ15N")
plot4 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "symb", "log10symbcount", "δ13C", "δ13C")
plot5 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "log10symbcount", "δ13C", "δ13C")
plot6 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "log10symbcount", "Δ13C", "minus symbiont, Δ13C")


legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(plot1 + theme(legend.position = "top"))), "guide-box")

acuta_SIvssymb_plots <- grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, legend, ncol = 3)
```

I'm not surprised these mostly look like flat lines. Let's look at the symbionts versus the rus:

```{r Prus symb vs isotopes plots, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
plot1 <- create_SI_plot(SI_data_wide, "Porites rus", "symb", "log10symbcount","δ15N", "δ15N")
plot2 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "log10symbcount", "δ15N", "δ15N")
plot3 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "log10symbcount", "Δ15N", "minus symbiont, Δ15N")
plot4 <- create_SI_plot(SI_data_wide, "Porites rus", "symb", "log10symbcount", "δ13C", "δ13C")
plot5 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "log10symbcount", "δ13C", "δ13C")
plot6 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "log10symbcount", "Δ13C", "minus symbiont, Δ13C")


legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(plot1 + theme(legend.position = "top"))), "guide-box")

rus_SIvssymb_plots <- grid.arrange(plot1, plot2, plot4, plot5, plot3, plot6, legend, ncol = 2)
```

Ok, we expect Δ13C(h-s) to increase with increasing symbiont populations. 

#### Just running a model to check that bc the last plot looks very significant:

```{r rus symbiont vs isotope C}
Rus_data <- SI_data_wide %>%
  filter(Species == "Porites rus")
model <- lmer(Δ13C ~ log10(FSC.Events_per_cm_2) + (1|Genotype), data = Rus_data)
car::Anova(model)
```
Ok so the host is going to have more 13C than the symbiont at higher symbiont populations in Porties rus corals. 

Increased contributions of heterotrophically-derived carbon results in more pronounced differences between δ13C levels in host and Symbiodiniaceae (∆13C = δ13Chost - δ13Csymbiont) (Muscatine et al., 1989; M. E. A. Santos et al., 2021).

“During periods of high coral photosynthesis the internal carbon pool becomes enriched in 13C (higher δ13C values) as dissolved inorganic carbon enriched in 12C is preferentially fixed” [Wall et al., 2020, p. 954]. 

```{r Rus Δ13C ~ Maximum_Phosphate_umolL}
model <- lmer(Δ13C ~ Maximum_Phosphate_umolL+ (1|Genotype), data = Rus_data)
car::Anova(model)
```
```{r Rus Δ13C ~ Low_Tide_Mean_Phosphate_umolL}
model <- lmer(Δ13C ~ Low_Tide_Mean_Phosphate_umolL+ (1|Genotype), data = Rus_data)
car::Anova(model)
```


```{r rus symbiont vs isotope N}
Rus_data <- SI_data_wide %>%
  filter(Species == "Porites rus")
model <- lmer(Δ15N ~ log10(FSC.Events_per_cm_2) + (1|Genotype), data = Rus_data)
car::Anova(model)
```

But, as expected from the plot, this is not congruent between C and N isotopes. 15N normally comes from more heterotrophy, whereas 14N is normally fixed from the environment.

## Some univariate models with isotopes:

```{r PAC_Δ15N_model_results}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)

plotlist <- list()

parameters <- unique(SI_data$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(SI_data, Parameter == parameter)
  subset_data <- subset(subset_data, Species == "Pocillopora acuta")
  
  # Fit the mixed-effects model
  model <- lmer(Δ15N ~ value + (1|Genotype), data = subset_data)
  
  # Get AICc and p-value
    AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["value", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["value", "Estimate"]
  
  # Calculate marginal and conditional R-squared values
    rsq <- as.numeric(r.squaredGLMM(model))
    marginal_R2 <- rsq[1]
    conditional_R2 <- rsq[2]
  
  if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate,
                                                     Marginal_R2 = marginal_R2,
                                                     Conditional_R2 = conditional_R2))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value[all]")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }
print(model_results)
grid.arrange(grobs = plotlist, ncol = 2)

model_results$Species <- "Pocillopora acuta"
model_results$type <- "linear"

PAC_Δ15N_model_results <- model_results
```

```{r PAC_Δ13C_model_results}
model_results <- data.frame(Parameter = character(),
                            AICc = numeric(),
                            p.value = numeric(),
                            Estimate = numeric(),
                            stringsAsFactors = FALSE)

plotlist <- list()

parameters <- unique(SI_data$Parameter)

for (parameter in parameters) {
  # Create a subset of data for the current parameter
  subset_data <- subset(SI_data, Parameter == parameter)
  subset_data <- subset(subset_data, Species == "Pocillopora acuta")
  
  # Fit the mixed-effects model
  model <- lmer(Δ13C ~ value + (1|Genotype), data = subset_data)
  
  # Get AICc and p-value
    AICc <- AICc(model)
  
  # Get p-value from summary
  model_summary <- summary(model)
  p.value <- model_summary$coefficients["value", "Pr(>|t|)"]
  Estimate <- model_summary$coefficients["value", "Estimate"]
  
  # Calculate marginal and conditional R-squared values
    rsq <- as.numeric(r.squaredGLMM(model))
    marginal_R2 <- rsq[1]
    conditional_R2 <- rsq[2]
  
  if (p.value < 0.05) {
    # Add results to the dataframe
    model_results <- rbind(data.frame(Parameter = parameter,
                                                     AICc = AICc,
                                                     p.value = p.value,
                                                     Estimate = Estimate,
                                                     Marginal_R2 = marginal_R2,
                                                     Conditional_R2 = conditional_R2))
    
    # Create an effects plot using ggpredict
    effect_plot <- ggpredict(model, terms = "value[all]")
    
    # Extract the ggplot object from the ggpredict object
    ggplot_obj <- plot(effect_plot)
    
    # Add the ggplot object to the plotlist
    plotlist[[parameter]] <- ggplot_obj + ggtitle(parameter)
  }
  }
print(model_results)
grid.arrange(grobs = plotlist, ncol = 2)

model_results$Species <- "Pocillopora acuta"
model_results$type <- "linear"

PAC_Δ13C_model_results <- model_results
```


### Versus growth

```{r}
SI_data_wide <- left_join(SI_data_wide, BW[,c("Placement_Code", "Percent_Change")])
SI_data_wide$Percent_Change <-100*SI_data_wide$Percent_Change
```

```{r, warning=FALSE, message=FALSE}
plot1 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "symb", "Percent_Change","δ15N", "δ15N")
plot2 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "Percent_Change", "δ15N", "δ15N")
plot3 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "Percent_Change", "Δ15N", "minus symbiont, Δ15N")
plot4 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "symb", "Percent_Change", "δ13C", "δ13C")
plot5 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "Percent_Change", "δ13C", "δ13C")
plot6 <- create_SI_plot(SI_data_wide, "Pocillopora acuta", "host", "Percent_Change", "Δ13C", "minus symbiont, Δ13C")


legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(plot1 + theme(legend.position = "top"))), "guide-box")

acuta_SIvsgrowth_plots <- grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, legend, ncol = 3)
```

```{r}
plot1 <- create_SI_plot(SI_data_wide, "Porites rus", "symb", "Percent_Change","δ15N", "δ15N")
plot2 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "Percent_Change", "δ15N", "δ15N")
plot3 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "Percent_Change", "Δ15N", "minus symbiont, Δ15N")
plot4 <- create_SI_plot(SI_data_wide, "Porites rus", "symb", "Percent_Change", "δ13C", "δ13C")
plot5 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "Percent_Change", "δ13C", "δ13C")
plot6 <- create_SI_plot(SI_data_wide, "Porites rus", "host", "Percent_Change", "Δ13C", "minus symbiont, Δ13C")


legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(plot1 + theme(legend.position = "top"))), "guide-box")

rus_SIvsgrowth_plots <- grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, legend, ncol = 3)
```

