---
title: "Coral Growth and Symbionts Analysis"
author: "Callie Stephenson"
date: "2023-08-29"
output: github_document
---
## Introduction

This is a R Markdown file in which I hope to write out my analyses. I will use this document to process all the R scripts found in the R folder and complete my project. 

### NS reccomendation

 * Grab both dataframes and make the same version of nut
 * make graphs that have the min/max (like you've put into the PCA) and then the PC axes and then the bw
 * use BW just caged and do model selection like Kyle says

#### Loading the data
```{r load packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(fishualize)
library(lme4)
library(lmerTest) #figure out which lme package you need
library(ggpp)
library(ggplot2)
library(GGally)
library(ggeffects)
library(arm) #discrete hist
library(MuMIn) #dredge
library(effects) #effects plots for dredged models
library(curl) #curl
```

```{r load data, echo=FALSE, message=FALSE}
#explanatory data
all_nut <- read.csv("data/All_Nutrients_Processed.csv")
nut <- read.csv("data/March_nutrients_processed.csv")
nut_wide <- read.csv("data/Nutrient_data_wide.csv")
explanatory_seasonal <- read.csv("data/Explanatory_variables.csv")
turb <- read_csv(curl("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Biogeochem/August2021/Turb_NC.csv"))

#explanatory 2
cv_low_wet <- read.csv("data/Wet_Nutrients_Processed_CV_Low.csv")
cv_low_all <- read.csv("data/All_Nutrients_Processed_CV_Low.csv")

#Response data
sa <- read.csv("data/Surface_Area.csv")
BW <- read.csv("data/BW_tidy.csv") 
BW <- subset(BW, select=-c(1,4,5)) #remove some erroneous columns
TLE <- read.csv("data/TLE_summary.csv")
FCM <- read.csv("data/FCM_tidy.csv") %>% 
  rename(Placement_Code = PLACEMENT)

#metadata
PAC_meta <- read.csv("data/PAC_Coral_Codes.csv") 
PRU_meta <- read.csv("data/PRU_Coral_Codes.csv") 
```

```{r clean turb}
turb1 <- turb %>% 
  filter(startsWith(CowTagID, "V")& CowTagID != "VSEEP")

turb1 <- turb1 %>%
  dplyr::select(c(CowTagID,
                   N_percent,
                   C_N))
#use C:N 
#can also use N percent
#plot(turb$C_N, turb$N_percent)
#hist(turb$C_N)
#hist(turb$N_percent)
#N_percent looks better. Will use N_percent for models
```

```{r explanatory}
all_nut_long <- pivot_longer(
  data = all_nut,
  cols = c(Minimum_Salinity:CV_Ammonia_umolL)) 
names(all_nut_long)[names(all_nut_long) == "name"] <- "Parameter"

join <- explanatory_seasonal[,c(2:7)]
join <- unique(join)

explanatory_all <- left_join(join,all_nut_long, by = join_by(CowTagID))
explanatory_all_wide <- left_join(join, all_nut,by = join_by(CowTagID))
explanatory_all_wide <- left_join(explanatory_all_wide, turb1,by = join_by(CowTagID))
```

```{r response variables, echo=FALSE}
#meta takes a little massaging
PAC_meta <- PAC_meta[,c(3:10)]
PAC_meta$Species <- "Pocillopora acuta"
PAC_meta$Genotype <- gsub('PR21', 'PA21', PAC_meta$Genotype) #fix a typo in the genotypes
PRU_meta <- PRU_meta[,c(2:9)]
PRU_meta$Species <- "Porites rus"

meta <- rbind(PAC_meta, PRU_meta)
meta$Cage_Uncaged <- as.factor(meta$Cage_Uncaged)

#make a master response variable data sheet
response_data <- left_join(meta[,c(2:4,7,8)], BW[,c(2,4,5)], by = join_by(Placement_Code)) %>% 
  #select(!X) %>% #check that this doesn't need to be here
  left_join(FCM[,c(1,15,17)], by = join_by(Placement_Code)) %>% 
  left_join(TLE[,c(3:7)], by = join_by(Placement_Code)) 
response_data$CowTagID <- paste0("V",response_data$Pin_Number)

PRU_response_data <- response_data %>% 
  filter(Species == "Porites rus")

PRU_response_data_caged <- PRU_response_data %>% 
  filter(Cage_Uncaged == "C")

PAC_response_data <- response_data %>% 
  filter(Species == "Pocillopora acuta")

PAC_response_data_caged <- PAC_response_data %>% 
  filter(Cage_Uncaged == "C")
```

```{r all data seasonal, warning=FALSE}
#now make an all data that has both explanatory and response variables:
all_data_seasonal <- left_join(response_data, explanatory_seasonal, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
all_data_seasonal$Genotype <- as.factor(all_data_seasonal$Genotype)
all_PAC_data_seasonal <- all_data_seasonal %>% 
  filter(Species == "Pocillopora acuta")
```

```{r all data both seasons, warning=FALSE}
#now make an all data that has both explanatory and response variables:
all_data <- left_join(response_data, explanatory_all, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
all_data$Genotype <- as.factor(all_data$Genotype)
all_PAC_data <- all_data %>% 
  filter(Species == "Pocillopora acuta")

all_data_wide <- left_join(response_data, explanatory_all_wide, by = join_by(CowTagID))
```

## Data Exploration:
Data exploration focuses on the following points:
1. Outliers
2. Collinearity
3. Independence (Relationships between response and explanatory variables) 

```{r response data hist, fig.width=10, fig.height=10, echo=FALSE}
#png("output/response_data_hist.png", width = 1100, height = 1100)
par(mfrow = c(4,4), oma=c(1,1,3,1))
hist(PAC_response_data$Change_Over_Area, main="Change Over Area", col="lightblue", xlab="Change_Over_Area")
hist_columns <- c(7:13)
for (i in 1:length(hist_columns)) {
  hist(PAC_response_data[, hist_columns[i]], col="lightblue", 
       main = paste("PAC", colnames(PAC_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
hist(PRU_response_data$Change_Over_Area, main="Change Over Area", col="darkolivegreen3", xlab="Change_Over_Area")
for (i in 1:length(hist_columns)) {
  hist(PRU_response_data[, hist_columns[i]], col="darkolivegreen3", 
       main = paste("PRU", colnames(PRU_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
main_title <- "Response Data Exploration"
title(main = main_title, outer = TRUE)
#dev.off()
```

Ok so FSC.Events are both pretty right-skewed. CN says log10 transform these.

```{r wet season raw growth data plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
         plot = plot, 
         width = 15, height = 15)
}
```
Now, doing this again but for both seasons combined:
```{r both season raw growth data plots, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```

Same but let's do for FCM data

```{r wet season raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = FSC.Events_per_cm_2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  ggsave(filename = paste("output/wet_raw_symbiong_count_response_plot_", sp, ".png", sep = ""),
         plot = plot, 
         width = 15, height = 15)
}
```

And again, now for both seasons:
```{r both seasons raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = FSC.Events_per_cm_2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```


#### Look for Outliers:
Buoyant Weight
It looks like that 13 A is really off. We know it broke, so I am removing it and replotted next to it:
```{r BW Outlier Tidy, echo=FALSE}
# BW Plot
BW <- merge(BW, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
BW_no <- BW[!(BW$Placement_Code == "PRU V13 A"), ]
```

Now making a plot to look at outliers:

```{r BW Outlier Plot, echo=FALSE}
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
dotchart(BW$Change_Over_Area, main = "∆BW", 
         groups = BW$Cage_Uncaged)
dotchart(BW_no$Change_Over_Area, main = "∆BW removed erroneous data point", 
         groups = BW_no$Cage_Uncaged)
layout(1)
```
I removed PRU V13 A placement because this was broken in the field, creating a large negative change in buoyant weight. 

TLE
```{r TLE Plot, echo=FALSE}
TLE <- merge(TLE, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
layout(matrix(c(1, 2, 3, 4), 2, 2, byrow = TRUE)) # Setting up the layout
dotchart(TLE$Max_L_AVG, main = "Max L", groups = TLE$Cage_Uncaged) # Side Max L Plot
dotchart(TLE$Max_L_Basal, main = "Max L Basal", groups = TLE$Cage_Uncaged) # Basal Max L plot
dotchart(TLE$Area_AVG, main = "Area", groups = TLE$Cage_Uncaged) # Area plot
dotchart(TLE$Area_Basal, main = "Basal Area", groups = TLE$Cage_Uncaged) # Basal Area plot
layout(1) # Reset the layout
title("TLE Outlier Dotchart", line = -1, outer = TRUE, cex.main = 1.5)
```

### Symbionts

```{r Sym plot, echo=FALSE}
FCM <- FCM[complete.cases(FCM[, "Cage_Uncaged"]), ] #Remove T0 samples by only including rows where the "Cage_Uncaged" exists
FCM$Cage_Uncaged <- as.factor(FCM$Cage_Uncaged) 
FCM_no <- FCM[!(FCM$Placement_Code == "PRU V17 C"), ]
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
dotchart(FCM$FSC.Events_per_cm_2, main = "Symbionts", groups = FCM$Cage_Uncaged)
dotchart(FCM_no$FSC.Events_per_cm_2, main = "Symbionts without erroneous point", groups = FCM_no$Cage_Uncaged)
layout(1)
```
Jess will rerun PRU 49 (PRU V17 C) with her samples - I likely added twice as much as needed to the sample, but for now I will omit in my analyses (using dataframe FCM_no for "no outliers")

```{r FCM Violin Plot}
ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_cm_2), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1)+
  facet_wrap(~Species)+
  labs(title = expression(paste("Symbionts per ", cm^2, " by caging treatment")))+
  theme(plot.title = element_text(hjust = 0.5))
```
same but other normalization:
```{r FCM Violin Plot Dry Weight}
ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_g_dry_weight), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1) +
  facet_wrap(~Species)+
  labs(title = "Symbionts per g dry weight by caging treatment")+
  theme(plot.title = element_text(hjust = 0.5))
```

### Collinearity
need to make a dataframe that's long that has all the nutrient metrics and independent variables
Here's the thing: we KNOW the nutrients are collinear
For example, here's a pairs plot for the maximum of all the nutrient values:

```{r pairs plot collinearity}
nut_wide_no_seep<- nut_wide[nut_wide$CowTagID != 'VSEEP', ]
maximum_columns <- names(nut_wide)[grepl("Maximum_", names(nut_wide)) & !grepl("Salinity|Temperature|pH|Ammonia_umol", names(nut_wide))]

ggpairs(nut_wide_no_seep[, c(maximum_columns)],progress = FALSE)
```

```{r}

```


### Independence
```{r independence}

```

### Data Analysis


FCM Data Analysis:
```{r FCM vis, echo=FALSE}
anova_PRU <- aov(FSC.Events_per_cm_2 ~ Cage_Uncaged, 
                 data = FCM_no %>% 
                   filter(Species=="Porites_rus"))
summary(anova_PRU)

#caging did not effect P. rus

anova_PAC <- aov(FSC.Events_per_cm_2 ~ Cage_Uncaged, 
                 data = FCM_no %>% 
                   filter(Species=="Pocillopora_acuta"))
summary(anova_PAC)
#caging did not effect symbionts in P. acuta
#this I have a bit of a hard time believing 
```

### Growth Models

```{r columns to dredge}
#minimum_columns <- names(all_data_wide)[grepl("Minimum_", names(all_data_wide)) & 
#                                     grepl("Salinity|Temperature|pH", names(all_data_wide))]
#maximum_columns <- names(all_data_wide)[grepl("Maximum_", names(all_data_wide)) & 
#                                  !grepl("Salinity|Temperature|pH|Ammonia_umol",names(all_data_wide))]
#mean_columns <- names(all_data_wide)[grepl("Mean_", names(all_data_wide)) & 
#                                  !grepl("Salinity|Temperature|pH|Ammonia_umol",names(all_data_wide))]
CV_columns <- names(all_data_wide)[grepl("CV_", names(all_data_wide), names(all_data_wide))]

mean_low_columns <- names(all_data_wide)[grepl("Low_Tide_Mean_", names(all_data_wide), names(all_data_wide))]

#dredge.data <- na.omit(all_data_wide[, c(maximum_columns, 
#                                    minimum_columns,
#                                    mean_columns, #removed bc not enough data for the amount of variables
#                                    CV_columns)])

PAC_model_data <- na.omit(all_data_wide %>% 
        filter(Species == "Pocillopora acuta"))

# Extract column names for filtering
selected_columns <- c("FSC.Events_per_cm_2", "Percent_Change", "Genotype", "Cage_Uncaged", 
                      CV_columns, mean_low_columns)

# Filter the data and select the relevant columns
PAC_model_data <- PAC_model_data[, selected_columns]

#PAC_model_data$CV_Salinity_scaled <- scale(PAC_model_data$CV_Salinity, TRUE, TRUE)
#PAC_model_data$CV_Temperature_scaled  <-  scale(PAC_model_data$CV_Temperature, TRUE, TRUE)
#PAC_model_data$CV_TA_scaled  <-  scale(PAC_model_data$CV_TA, TRUE, TRUE)
#PAC_model_data$CV_pH_scaled  <-  scale(PAC_model_data$CV_pH, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_Phosphate_umolL_scaled<-
#  scale(PAC_model_data$Low_Tide_Mean_Phosphate_umolL, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_NN_umolL_scaled  <-  
#  scale(PAC_model_data$Low_Tide_Mean_NN_umolL, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_Ammonia_umolL_scaled  <-
#  scale(PAC_model_data$Low_Tide_Mean_Ammonia_umolL, TRUE, TRUE)
#PAC_model_data$Low_Tide_Mean_NNP_umolL_scaled <- 
#  scale(PAC_model_data$Low_Tide_Mean_Ammonia_umolL, TRUE, TRUE)


for (col in CV_columns) {
  scaled_variable <- paste0(col, "_scaled")
  scaled_values <- scale(PAC_model_data[[col]], TRUE, TRUE)
  PAC_model_data[[scaled_variable]] <- scaled_values[, 1]
}

for (col in mean_low_columns) {
  scaled_variable <- paste0(col, "_scaled")
  scaled_values <- scale(PAC_model_data[[col]], TRUE, TRUE)
  PAC_model_data[[scaled_variable]] <- scaled_values[, 1]
}

scaled_cv_columns <- names(PAC_model_data)[grepl("^CV.*_scaled$", names(PAC_model_data))]
scaled_ml_columns <- names(PAC_model_data)[grepl("^Low.*_scaled$", names(PAC_model_data))]
```

```{r collinearity in cv dredge parameters}
ggpairs(PAC_model_data, columns = c("CV_Salinity_scaled",  
                       "CV_Temperature_scaled",  
                       "CV_TA_scaled",  
                       "CV_pH_scaled"), progress = FALSE)
```

```{r collinearity in dredge parameters}
ggpairs(PAC_model_data, columns = c("Low_Tide_Mean_Phosphate_umolL",  
                       "Low_Tide_Mean_NN_umolL", 
                       "Low_Tide_Mean_Ammonia_umolL",
                       "Low_Tide_Mean_NNP_umolL"), progress = FALSE)
```

Extra exploratory plots:
```{r exploratory plots with bw change}
PAC_model_data_long <- pivot_longer(
  data = PAC_model_data,
  cols = c("Low_Tide_Mean_Phosphate_umolL",  
                       "Low_Tide_Mean_NN_umolL", 
                       "Low_Tide_Mean_Ammonia_umolL",
                       "Low_Tide_Mean_NNP_umolL",
           "CV_Salinity_scaled",  
                       "CV_Temperature_scaled",  
                       "CV_TA_scaled",  
                       "CV_pH_scaled"),
  names_to = "Parameter",
  values_to = "Value")

ggplot(PAC_model_data_long, aes(x = Value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    theme_bw()
```


```{r exploratory plots with symb}
ggplot(PAC_model_data_long, aes(x = Value, y = FSC.Events_per_cm_2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    theme_bw()
```


```{r Growth Dredge CV or Low Tide Pulse, warning=FALSE, message=FALSE}
#PAC_cv_model <- glm(Percent_Change ~ ., data = PAC_cv_model_data, na.action = na.pass)
#the above line works if you only include response and explanatory in the dataframe, but dredge can't handle it with random effects

# Build the GLM model
#PAC_cv_model <- glm(Percent_Change ~ ., data = PAC_cv_model_data, na.action = na.pass)
PAC_cv_low_growth_model <- lmer(Percent_Change ~ 
                       CV_Salinity_scaled + 
                       CV_Temperature_scaled + 
                       CV_TA_scaled + 
                       CV_pH_scaled +
                       Low_Tide_Mean_Phosphate_umolL_scaled + 
                       Low_Tide_Mean_NN_umolL_scaled + 
                       Low_Tide_Mean_Ammonia_umolL_scaled +
                       Low_Tide_Mean_NNP_umolL_scaled +
                       (1 | Genotype) + (1 | Cage_Uncaged), 
                                 data = PAC_model_data, 
                                 na.action = na.pass)

PAC_cv_low_growth_model_dredge <- dredge(PAC_cv_low_growth_model, trace = 2)
model.sel(PAC_cv_low_growth_model_dredge)
PAC.avg_cv_low_growth_model = model.avg(PAC_cv_low_growth_model_dredge, delta < 5)
summary(PAC.avg_cv_low_growth_model)
PAC.avg_cv_low_growth_model$sw
```

Make the effects plots for the model with the most support:
```{r Growth CV + Low tide best model effects plots}
PAC_growth_cv_plus_lowtide_model <- lmer(Percent_Change ~ 
#                       CV_Salinity_scaled + 
#                       CV_Temperature_scaled + 
                       CV_TA_scaled + 
                       CV_pH_scaled +
                       (1 | Genotype) + (1 | Cage_Uncaged), 
                                 data = PAC_model_data, 
                                 na.action = na.pass)
plot(allEffects(PAC_growth_cv_plus_lowtide_model))
```

### Symbiont Count Models

```{r Symbiont data CV AND low tide data dredge, warning=FALSE, message=FALSE}
# Build the GLM model
#PAC_cv_model <- glm(Percent_Change ~ ., data = PAC_cv_model_data, na.action = na.pass)
PAC_cv_low_symb_model <- lmer(log10(FSC.Events_per_cm_2) ~
                       Low_Tide_Mean_Phosphate_umolL_scaled + 
                       Low_Tide_Mean_NN_umolL_scaled + 
                       Low_Tide_Mean_Ammonia_umolL_scaled +
                       Low_Tide_Mean_NNP_umolL_scaled +
                       (1 | Genotype) + (1 | Cage_Uncaged), 
                                 data = PAC_model_data, 
                                 na.action = na.pass)

PAC_cv_low_symb_model_dredge <- dredge(PAC_cv_low_symb_model, trace = 2)
model.sel(PAC_cv_low_symb_model_dredge)
PAC_cv_low_symb_model_dredge_model.avg = model.avg(PAC_cv_low_symb_model_dredge, delta < 8)
summary(PAC_cv_low_symb_model_dredge_model.avg)
PAC_cv_low_symb_model_dredge_model.avg$sw
```


```{r Symbiont cv and low tide best model effects plots}
PAC_symb_cv_plus_lowtide_model <- lmer(Percent_Change ~ 
                       CV_Salinity + 
                       CV_Temperature + 
                       CV_TA + 
                       CV_pH +
                       (1 | Genotype) + (1 | Cage_Uncaged), 
                                 data = PAC_model_data, 
                                 na.action = na.pass)
plot(allEffects(PAC_symb_cv_plus_lowtide_model))
```

