---
title: "Coral"
author: "Callie Stephenson"
date: "2023-08-29"
output: github_document
---
## Introduction

This is a R Markdown file in which I hope to write out my analyses. I will use this document to process all the R scripts found in the R folder and complete my project. 

### NS reccomendation

 * Grab both dataframes and make the same version of nut
 * make graphs that have the min/max (like you've put into the PCA) and then the PC axes and then the bw
 * use BW just caged and do model selection like Kyle says

#### Loading the data
```{r load packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(fishualize)
library(lme4)
library(lmerTest) #figure out which lme package you need
library(ggpp)
library(ggplot2)
library(GGally)
library(ggeffects)
library(arm) #discrete hist
library(MuMIn) #dredge
```

```{r load data, echo=FALSE}
#explanatory data
all_nut <- read.csv("data/All_Nutrients_Processed.csv")
nut <- read.csv("data/March_nutrients_processed.csv")
nut_wide <- read.csv("data/Nutrient_data_wide.csv")
explanatory_seasonal <- read.csv("data/Explanatory_variables.csv")

#Response data
sa <- read.csv("data/Surface_Area.csv")
BW <- read.csv("data/BW_tidy.csv") 
BW <- subset(BW, select=-c(1,4,5)) #remove some erroneous columns
TLE <- read.csv("data/TLE_summary.csv")
FCM <- read.csv("data/FCM_tidy.csv") %>% 
  rename(Placement_Code = PLACEMENT)

#metadata
PAC_meta <- read.csv("data/PAC_Coral_Codes.csv") 
PRU_meta <- read.csv("data/PRU_Coral_Codes.csv") 
```

```{r explanatory}
all_nut_long <- pivot_longer(
  data = all_nut,
  cols = c(Minimum_Salinity:CV_Ammonia_umolL)) 
names(all_nut_long)[names(all_nut_long) == "name"] <- "Parameter"

join <- explanatory_seasonal[,c(2:7)]
join <- unique(join)

explanatory_all <- left_join(join,all_nut_long, by = join_by(CowTagID))
explanatory_all_wide <- left_join(join, all_nut,by = join_by(CowTagID))
```

```{r response variables, echo=FALSE}
#meta takes a little massaging
PAC_meta <- PAC_meta[,c(3:10)]
PAC_meta$Species <- "Pocillopora acuta"
PAC_meta$Genotype <- gsub('PR21', 'PA21', PAC_meta$Genotype) #fix a typo in the genotypes
PRU_meta <- PRU_meta[,c(2:9)]
PRU_meta$Species <- "Porites rus"

meta <- rbind(PAC_meta, PRU_meta)
meta$Cage_Uncaged <- as.factor(meta$Cage_Uncaged)

#make a master response variable data sheet
response_data <- left_join(meta[,c(2:4,7,8)], BW[,c(2,4,5)], by = join_by(Placement_Code)) %>% 
  #select(!X) %>% #check that this doesn't need to be here
  left_join(FCM[,c(1,15,17)], by = join_by(Placement_Code)) %>% 
  left_join(TLE[,c(3:7)], by = join_by(Placement_Code)) 
response_data$CowTagID <- paste0("V",response_data$Pin_Number)

PRU_response_data <- response_data %>% 
  filter(Species == "Porites rus")

PRU_response_data_caged <- PRU_response_data %>% 
  filter(Cage_Uncaged == "C")

PAC_response_data <- response_data %>% 
  filter(Species == "Pocillopora acuta")

PAC_response_data_caged <- PAC_response_data %>% 
  filter(Cage_Uncaged == "C")
```

```{r all data seasonal, warning=FALSE}
#now make an all data that has both explanatory and response variables:
all_data_seasonal <- left_join(response_data, explanatory_seasonal, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
all_data_seasonal$Genotype <- as.factor(all_data_seasonal$Genotype)
all_PAC_data_seasonal <- all_data_seasonal %>% 
  filter(Species == "Pocillopora acuta")
```

```{r all data both seasons, warning=FALSE}
#now make an all data that has both explanatory and response variables:
all_data <- left_join(response_data, explanatory_all, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
all_data$Genotype <- as.factor(all_data$Genotype)
all_PAC_data <- all_data %>% 
  filter(Species == "Pocillopora acuta")

all_data_wide <- left_join(response_data, explanatory_all_wide, by = join_by(CowTagID))
```

## Data Exploration:
Data exploration focuses on the following points:
1. Outliers
2. Collinearity
3. Independence (Relationships between response and explanatory variables) 

```{r response data hist, fig.width=10, fig.height=10, echo=FALSE}
#png("output/response_data_hist.png", width = 1100, height = 1100)
par(mfrow = c(4,4), oma=c(1,1,3,1))
hist(PAC_response_data$Change_Over_Area, main="Change Over Area", col="lightblue", xlab="Change_Over_Area")
hist_columns <- c(7:13)
for (i in 1:length(hist_columns)) {
  hist(PAC_response_data[, hist_columns[i]], col="lightblue", 
       main = paste("PAC", colnames(PAC_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
hist(PRU_response_data$Change_Over_Area, main="Change Over Area", col="darkolivegreen3", xlab="Change_Over_Area")
for (i in 1:length(hist_columns)) {
  hist(PRU_response_data[, hist_columns[i]], col="darkolivegreen3", 
       main = paste("PRU", colnames(PRU_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
main_title <- "Response Data Exploration"
title(main = main_title, outer = TRUE)
#dev.off()
```

Ok so FSC.Events are both pretty right-skewed. CN says log10 transform these.

```{r wet season raw growth data plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
         plot = plot, 
         width = 15, height = 15)
}
```
Now, doing this again but for both seasons combined:
```{r both season raw growth data plots, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```

Same but let's do for FCM data

```{r wet season raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = FSC.Events_per_cm_2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  ggsave(filename = paste("output/wet_raw_symbiong_count_response_plot_", sp, ".png", sep = ""),
         plot = plot, 
         width = 15, height = 15)
}
```

And again, now for both seasons:
```{r both seasons raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = FSC.Events_per_cm_2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```


#### Look for Outliers:
Buoyant Weight
It looks like that 13 A is really off. We know it broke, so I am removing it and replotted next to it:
```{r BW Outlier Tidy, echo=FALSE}
# BW Plot
BW <- merge(BW, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
BW_no <- BW[!(BW$Placement_Code == "PRU V13 A"), ]
```

Now making a plot to look at outliers:

```{r BW Outlier Plot, echo=FALSE}
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
dotchart(BW$Change_Over_Area, main = "∆BW", 
         groups = BW$Cage_Uncaged)
dotchart(BW_no$Change_Over_Area, main = "∆BW removed erroneous data point", 
         groups = BW_no$Cage_Uncaged)
layout(1)
```
I removed PRU V13 A placement because this was broken in the field, creating a large negative change in buoyant weight. 

TLE
```{r TLE Plot, echo=FALSE}
TLE <- merge(TLE, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
layout(matrix(c(1, 2, 3, 4), 2, 2, byrow = TRUE)) # Setting up the layout
dotchart(TLE$Max_L_AVG, main = "Max L", groups = TLE$Cage_Uncaged) # Side Max L Plot
dotchart(TLE$Max_L_Basal, main = "Max L Basal", groups = TLE$Cage_Uncaged) # Basal Max L plot
dotchart(TLE$Area_AVG, main = "Area", groups = TLE$Cage_Uncaged) # Area plot
dotchart(TLE$Area_Basal, main = "Basal Area", groups = TLE$Cage_Uncaged) # Basal Area plot
layout(1) # Reset the layout
title("TLE Outlier Dotchart", line = -1, outer = TRUE, cex.main = 1.5)
```

### Symbionts

```{r Sym plot, echo=FALSE}
FCM <- FCM[complete.cases(FCM[, "Cage_Uncaged"]), ] #Remove T0 samples by only including rows where the "Cage_Uncaged" exists
FCM$Cage_Uncaged <- as.factor(FCM$Cage_Uncaged) 
FCM_no <- FCM[!(FCM$Placement_Code == "PRU V17 C"), ]
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
dotchart(FCM$FSC.Events_per_cm_2, main = "Symbionts", groups = FCM$Cage_Uncaged)
dotchart(FCM_no$FSC.Events_per_cm_2, main = "Symbionts without erroneous point", groups = FCM_no$Cage_Uncaged)
layout(1)
```
Jess will rerun PRU 49 (PRU V17 C) with her samples - I likely added twice as much as needed to the sample, but for now I will omit in my analyses (using dataframe FCM_no for "no outliers")

```{r FCM Violin Plot}
ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_cm_2), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1)+
  facet_wrap(~Species)+
  labs(title = expression(paste("Symbionts per ", cm^2, " by caging treatment")))+
  theme(plot.title = element_text(hjust = 0.5))
```
same but other normalization:
```{r FCM Violin Plot Dry Weight}
ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_g_dry_weight), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1) +
  facet_wrap(~Species)+
  labs(title = "Symbionts per g dry weight by caging treatment")+
  theme(plot.title = element_text(hjust = 0.5))
```

### Collinearity
need to make a dataframe that's long that has all the nutrient metrics and independent variables
Here's the thing: we KNOW the nutrients are collinear
For example, here's a pairs plot for the maximum of all the nutrient values:

```{r pairs plot collinearity}
nut_wide_no_seep<- nut_wide[nut_wide$CowTagID != 'VSEEP', ]
maximum_columns <- names(nut_wide)[grepl("Maximum_", names(nut_wide)) & !grepl("Salinity|Temperature|pH|Ammonia_umol", names(nut_wide))]

ggpairs(nut_wide_no_seep[, c(maximum_columns)],progress = FALSE)
```

```{r}

```


### Independence
```{r independence}

```

### Data Analysis


FCM Data Analysis:
```{r FCM vis, echo=FALSE}
anova_PRU <- aov(FSC.Events_per_cm_2 ~ Cage_Uncaged, 
                 data = FCM_no %>% 
                   filter(Species=="Porites_rus"))
summary(anova_PRU)

#caging did not effect P. rus

anova_PAC <- aov(FSC.Events_per_cm_2 ~ Cage_Uncaged, 
                 data = FCM_no %>% 
                   filter(Species=="Pocillopora_acuta"))
summary(anova_PAC)
#caging did not effect symbionts in P. acuta
#this I have a bit of a hard time believing 
```

### Models
#### Need to add a caged effect
#### Random effect of genotype

```{r}
minimum_columns <- names(all_data_wide)[grepl("Minimum_", names(all_data_wide)) & 
                                     grepl("Salinity|Temperature|pH", names(all_data_wide))]
maximum_columns <- names(all_data_wide)[grepl("Maximum_", names(all_data_wide)) & 
                                  !grepl("Salinity|Temperature|pH|Ammonia_umol",names(all_data_wide))]
mean_columns <- names(all_data_wide)[grepl("Mean_", names(all_data_wide)) & 
                                  !grepl("Salinity|Temperature|pH|Ammonia_umol",names(all_data_wide))]
CV_columns <- names(all_data_wide)[grepl("CV_", names(all_data_wide), names(all_data_wide))]

dredge.data <- na.omit(all_data_wide[, c(maximum_columns, 
                                    minimum_columns,
                                    mean_columns, #removed bc not enough data for the amount of variables
                                    CV_columns)])
```


```{r PAC dredge}
PAC_model_data <- na.omit(all_data_wide %>% 
        filter(Species == "Pocillopora acuta"))
#PAC_big_model <- 
#  glm(Percent_Change ~ CV_Phosphate_umolL + 
#        shore_dist_m + Mean_Silicate_umolL + Minimum_pH, 
#      data = PAC_model_data, na.action = na.pass)

# Extract column names for filtering
selected_columns <- c("Percent_Change", mean_columns, CV_columns)

# Filter the data and select the relevant columns
PAC_model_data <- PAC_model_data[, selected_columns]

# Build the GLM model
PAC_big_model <- glm(Percent_Change ~ ., data = PAC_model_data, na.action = na.pass)

PAC_big_model_dredge <- dredge(PAC_big_model)
model.sel(PAC_big_model_dredge)
```

```{r KE Dredge 2}
PAC.avg = model.avg(PAC_big_model_dredge, delta < 4)
summary(PAC.avg)
```

OK 




```{r KE Dredge PRU}
PRU_model_data <- na.omit(all_data_wide %>% 
        filter(Species == "Porites rus" & Cage_Uncaged == "C"))
PRU_big_model <- 
  glm(Percent_Change ~ CVSeasonal_Phosphate_umolL + 
        shore_dist_m + MeanSeasonal_Silicate_umolL + Maximum_pH, 
      data = PAC_model_data, na.action = na.pass)

PRU_big_model_dredge <- dredge(PRU_big_model)
model.sel(PRU_big_model_dredge)
```

```{r}
PRU.avg = model.avg(PRU_big_model_dredge, delta < 4)
summary(PRU.avg)
```


```{r}
bw_caged_PAC <-lmer(Percent_Change ~ shore_dist_m + (1 | Genotype), data = all_data %>%                     filter(Species == "Pocillopora acuta" 
                           & Cage_Uncaged == "C"))
anova(bw_caged_PAC)

data <- all_data %>% 
                     filter(Species == "Pocillopora acuta" 
                           & Cage_Uncaged == "C" 
                           & Parameter == "pc1") 
#this removes V13 because it doesn't have shore dist
ggplot(data, aes(x = X._loss, y = Percent_Change))+
  geom_point(aes(color = value))+
  scale_color_fish(option = "Coryphaena_hippurus")+
  geom_smooth(method = "lm", formula = 'y~poly(x,2)')+
  labs(title="Growth of Pocillopora acuta colonies against distance from shore")
```
So this is suggesting that distance from shore is significant, let's see if SGD is too:
```{r}
bw_caged_PAC <-lmer(Percent_Change ~ shore_dist_m + value + (1 | Genotype), data = all_data %>% 
                     filter(Species == "Pocillopora acuta" 
                           & Cage_Uncaged == "C" 
                           & Parameter == "pc1") ) 
anova(bw_caged_PAC)
```


Now that we've done our data exploration, let's make the data frame needed to build some models!

```{r starting models with megan}
#For models of only caged corals, we are running with Genotype as covariate instead of random effect
#Cannot run as random effect without singularity (as there is only one occurance of genotype for every occurence of the nutrient parameter)
#for example, see how this is singular:

bw_m1_caged_PAC <-lmer(Percent_Change ~ shore_dist_m + (1 | Genotype), data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "Pocillopora acuta" & Cage_Uncaged == "C")) 
anova(bw_m1_caged_PAC)
summary(bw_m1_caged_PAC)

##plots with megan
ggplot(all_PAC_data) +
  geom_point(data = subset(all_PAC_data, Cage_Uncaged == "C"),
             aes(x = Genotype, y = Percent_Change, color = shore_dist_m)) +
  scale_color_fish(option = "Coryphaena_hippurus")
#this shore distance looks like it could be a trend

#variance components analysis within genotype and between genotype variance
#looking at all the variance in the data, looking at the grand mean vs. the mean for the random effects
#how much is between vs within genotype variance
bw_m1_caged_PRU <-lmer(Percent_Change ~ shore_dist_m + (1 | Genotype), data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "Porites rus" & Cage_Uncaged == "C")) 
anova(bw_m1_caged_PRU)
summary(bw_m1_caged_PRU)

ggplot(all_PRU_data) +
  geom_point(data = subset(all_PRU_data, Cage_Uncaged == "C"),
             aes(x = Genotype, y = Percent_Change, color = shore_dist_m))+
  scale_color_fish(option = "Coryphaena_hippurus")
```
```{r WHY YOU GET SINGULARITY IN ACUTA, eval = FALSE, echo=FALSE}
dataPAC = all_data %>%
  filter(Parameters == "Salinity" & Species == "PAC" & Cage_Uncaged == "C")

dataPRU = all_data %>% 
  filter(Parameters == "Salinity" & Species == "PRU" & Cage_Uncaged == "C")

ggplot(dataPAC) +
  geom_point(aes(x = Genotype, y = Percent_Change, color = shore_dist_m))+
  scale_color_fish(option = "Coryphaena_hippurus")
#NEARLY NO VARIATION ATTRIBUTABLE TO GENOTYPE

ggplot(dataPRU) +
  geom_point(aes(x = Genotype, y = Percent_Change, color = shore_dist_m))+
  scale_color_fish(option = "Coryphaena_hippurus")

#Q from MJD for KE:
#if you get an estimate of 0 for your random effect what do you do
#do you keep bc the structure is important
#or do you drop because the model becomes simpler

#To move forward: drop the random effect for PAC models.
```

```{r starting models, eval = FALSE, echo=FALSE}
###
#Cannot run as random effect without singularity (as there is only one occurance of genotype for every occurence of the nutrient parameter)
bw_m1_caged_PAC <-lm(Percent_Change ~ CVSeasonal + shore_dist_m,
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PAC" & Cage_Uncaged == "C"))
anova(bw_m1_caged_PAC)

bw_m1_caged_PRU <-lmer(Percent_Change ~ CVSeasonal + shore_dist_m + (1 | Genotype),
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PRU" & Cage_Uncaged == "C"))
anova(bw_m1_caged_PRU)

#For models of all corals, we are running with Genotype as a random effect
bw_m1_PAC <- lmer(Percent_Change ~ CVSeasonal + (1 | Genotype) + Cage_Uncaged, 
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PAC"))
anova(bw_m1_PAC)

bw_m1_PRU <- lmer(Percent_Change ~ CVSeasonal + (1 | Genotype) + Cage_Uncaged, 
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PRU"))
anova(bw_m1_PRU)
summary(bw_m1_PRU)
```
Same models as before, but now with silicate instead, because the range in SiO32- was an order of magnitude higher on the reef than salinity allowing for a higher signal to noise ratio. I also used maximum silicate, as it is more indicative of the maximum amount of groundwater experienced

```{r starting models silicate, eval = FALSE, echo=FALSE}
#For models of only caged corals, we are running with Genotype as covariate instead of random effect
#Cannot run as random effect without singularity (as there is only one occurance of genotype for every occurence of the nutrient parameter)
#for example, see how this is singular:
bw_m1_caged_PAC <-lmer(Percent_Change ~ Maximum + (1 | Genotype), data = all_data %>% 
                    filter(Parameters == "Silicate_umolL" & Species == "PAC" & Cage_Uncaged == "C")) 
anova(bw_m1_caged_PAC)
                    
###
#Cannot run as random effect without singularity (as there is only one occurrence of genotype for every occurrence of the nutrient parameter)
bw_m1_caged_PAC <-lm(Percent_Change ~ Maximum + Genotype,
                  data = all_data %>% 
                    filter(Parameters == "Silicate_umolL" & 
                             Species == "PAC" & 
                             Cage_Uncaged == "C"))
anova(bw_m1_caged_PAC)

bw_m1_caged_PRU <-lm(Percent_Change ~ Maximum + Genotype,
                  data = all_data %>% 
                    filter(Parameters == "Silicate_umolL" & 
                             Species == "PRU" & 
                             Cage_Uncaged == "C"))
anova(bw_m1_caged_PRU)

#For models of all corals, we are running with Genotype as a random effect
bw_m1_PAC <- lmer(Percent_Change ~ Maximum + (1 | Genotype) + Cage_Uncaged, 
                  data = all_data %>% 
                    filter(Parameters == "Silicate_umolL" & 
                             Species == "PAC"))
anova(bw_m1_PAC)
summary(bw_m1_PAC)

bw_m1_PRU <- lmer(Percent_Change ~ Maximum + (1 | Genotype) + Cage_Uncaged, 
                  data = all_data %>% 
                    filter(Parameters == "Silicate_umolL" & 
                             Species == "PRU"))
anova(bw_m1_PRU)
summary(bw_m1_PRU)
```
This also only sees a correlation between all (not just caged) change in buoyant weight for PRU. PAC is not significant.

Now let's see those same models with symbiont counts
```{r starting models symb, eval = FALSE, echo=FALSE}
#again running with Genotype as covariate when only 
symb_m1_caged_PAC <-lm(Normalized_Symbiont_Counts ~ CVSeasonal + Genotype,
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PAC" & Cage_Uncaged == "C"))
anova(symb_m1_caged_PAC)

symb_m1_caged_PRU <-lm(Normalized_Symbiont_Counts ~ CVSeasonal + Genotype,
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PRU" & Cage_Uncaged == "C"))
anova(symb_m1_caged_PRU)

#but with all the corals we can use Genotype as random effect
symb_m1_PAC <- lmer(Normalized_Symbiont_Counts ~ CVSeasonal + (1 | Genotype) + Cage_Uncaged, 
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PAC"))
anova(symb_m1_PAC)

symb_m1_PRU <- lmer(Normalized_Symbiont_Counts ~ CVSeasonal + (1 | Genotype) + Cage_Uncaged, 
                  data = all_data %>% 
                    filter(Parameters == "Salinity" & Species == "PRU"))
anova(symb_m1_PRU)
summary(symb_m1_PRU)
```




