---
title: "PCA_Results"
author: "Callie Stephenson"
date: "2024-09-25"
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Question 1

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
```

## Loading the packages and data
```{r load packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(fishualize)
library(lme4)
library(ggplot2)
library(performance)
library(see)
library(dplyr)
library(car)
library(PerformanceAnalytics)
library(corrplot)
library(glmnet)
library(effects)
library(ggpubr)
```

```{r load data, echo=FALSE, message=FALSE}
setwd("/Users/calliestephenson/Documents/GitHub/MooreaCoralFish_CS")

#explanatory data
all_nut <- read.csv("data/All_Nutrients_Processed.csv")
shore_dist <- read.csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Shore_distance.csv") #using Danielle's
seep_dist <- read.csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Plate_Distance_to_Seep.csv") %>% 
  filter(Location == "Varari")
depth <- read.csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/Adj_Sandwich_Depth.csv")
clod <- read.csv("data/Clod_Cards.csv")
turb <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Biogeochem/August2021/Turb_NC.csv")

#Response data
response_data <- read.csv("data/response_data.csv")
response_data$Genotype <- as.factor(response_data$Genotype)
response_data$Cage_Uncaged <- as.factor(response_data$Cage_Uncaged)
response_data$Pin_Number <- as.factor(response_data$Pin_Number)

PRU_response_data <- response_data %>% 
  filter(Species == "Porites rus")

PRU_response_data_caged <- PRU_response_data %>% 
  filter(Cage_Uncaged == "C")

PAC_response_data <- response_data %>% 
  filter(Species == "Pocillopora acuta")

PAC_response_data_caged <- PAC_response_data %>% 
  filter(Cage_Uncaged == "C")

#metadata
meta <- read.csv("data/coral_metadata.csv")

#SI Stuff
SI_Zoop <- read.csv("data/SI_Zoop.csv")%>% 
  mutate(δ13C = δ13C............vs..VPDB.) %>% 
  mutate(δ15N = δ15N...........vs..AIR.) %>% 
  dplyr::select(-'δ13C............vs..VPDB.')%>% 
  dplyr::select(-'δ15N...........vs..AIR.')
```

```{r}
nut_no_seep <- all_nut[all_nut$CowTagID != 'VSEEP',]
nut_no_seep <- left_join(nut_no_seep, seep_dist[,c("CowTagID", "lat", "lon", "dist_to_seep_m")], by = join_by(CowTagID))
nut_no_seep <- left_join(nut_no_seep, shore_dist[,c("CowTagID", "shore_dist_m")], by = join_by(CowTagID))
```

```{r model data}
model_data <- left_join(response_data, nut_no_seep, by = join_by(CowTagID))
```

## Building a PCA
```{r}
pulse_columns <- names(all_nut)[grepl("Low_Tide_Mean_", names(all_nut)) & 
                            !grepl("Temperature|pH|Ammonia_umol|TA",names(all_nut))]
nut_no_seep_scaled <- nut_no_seep %>%
  mutate(across(all_of(pulse_columns), scale))

pca.data <- na.omit(nut_no_seep_scaled[, c(pulse_columns)]) 
pca = princomp(pca.data, cor=TRUE)
biplot(pca)
```

When I make a PCA of 
```{r}
names(pca.data)
```

we can explain 72.2% of the variation of these factors with 1 axis. 
```{r}
summary(pca)
loadings(pca)
```

```{r}
biplot(pca)
```

```{r failed autoplot}
# pca_plot <- autoplot(pca, loadings = TRUE, loadings.colour = '#47c16e',
#                      loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = '#47c16e') +
#   labs(
#     title = "Principal Component Analysis of Nutrient Data",
#     x = "Principal Component 1 (72.249%)",
#     y = "Principal Component 2 (15.019%)"
#   ) +
#   xlim(-0.5,0.5)+
# #  scale_x_reverse(-0.5,0.5) +
#   theme_minimal(base_size = 15) +  # Clean theme
#   theme(
#     plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),  # Center title
#     axis.title = element_text(size = 12)
#   )
# 
# pca_plot
#ggsave(pca_plot, filename = "PCA_Plot.png", dpi = 600, width = 7, height = 5)
```

```{r}
pulse_df <- as.data.frame(lapply(nut_no_seep_scaled[, c(pulse_columns)], as.numeric))
cor_matrix <- cor(pulse_df)
corrplot(cor_matrix, method = "number")
chart.Correlation(pulse_df)
```



### PCA of sd - doesn't work
```{r}
sd_columns <- names(all_nut)[grepl("sd_", names(all_nut)) & 
                                  !grepl("Phosphate_umolL|NN_umolL|Ammonia_umolL|NNP_umolL|Silicate_umolL",names(all_nut))]
sd.pca.data <- na.omit(nut_no_seep[, c(sd_columns)])
sd.pca = princomp(sd.pca.data, cor=TRUE)
names(sd.pca.data)
```

```{r}
summary(sd.pca)
loadings(sd.pca)
biplot(sd.pca)
```

As we can see from the biplot, on Comp 1, which is 72.2% of the variation, lower values on Comp 1 represent higher nutrient concentrations and a lower salinity value. This suggests that low values on Comp 1 represent pins that have high SGD values, and high numbers on Comp 1 represent low SGD values. I'm going to flop this so that 'high' numbers on comp1 mean 'high' nutrient environment.

Though Comp 2 helps capture another 20% of the variation, it doesn't appear to very interpretable to me. Silicate and Phosphate are higher with higher values, which are 'conserved', and therefore experiencing less 'drawdown'. Potentially this Comp 2 is higher values are retaining SGD after biological draw-down? This doesn't really play into my overall hypotheses, so I am sticking with Comp 1 here.

Here, the hollow yellow dot is the seep, and cooler colors have less nutrient influence, while more yellow colors have more nutrients at low tide. 

```{r}
nut_no_seep_scaled$pc1 = -1*(pca$scores[,1]) #what's this data's score on pc1 axis
nut_no_seep_scaled$pc2 = -1*(pca$scores[,2]) #what's this data's score on pc1 axis

ggplot(nut_no_seep_scaled, aes(x = lon, y = lat, color = pc1)) +
  geom_point() +
  labs(x = "Latitude", y = "Longitude", color = "PC1 of pulse") +
  scale_color_fish(option = "Coryphaena_hippurus", direction = -1) +
  geom_point(data = (seep_dist %>% filter(CowTagID == "VSEEP")), aes(x = lon, y = lat), shape = 21, size = 4, color = "#F1E700") +
  theme_minimal()
```

```{r}
nut_no_seep_scaled$sd_pc1 = sd.pca$scores[,1] #what's this data's score on pc1 axis
nut_no_seep_scaled$sd_pc2 = sd.pca$scores[,2] #what's this data's score on pc2 axis
ggplot(nut_no_seep_scaled, aes(x = lon, y = lat, color = sd_pc1)) +
  geom_point() +
  labs(x = "Latitude", y = "Longitude", color = "PC1 of SD (high numbers high sgd)") +
  scale_color_fish(option = "Antennarius_multiocellatus", direction = -1) +
  geom_point(data = (seep_dist %>% filter(CowTagID == "VSEEP")), aes(x = lon, y = lat), shape = 21, size = 4, color = "#F1E700") +
  theme_minimal()

ggplot(nut_no_seep_scaled, aes(x = lon, y = lat, color = sd_pc2)) +
  geom_point() +
  labs(x = "Latitude", y = "Longitude", color = "PC2 of SD (high numbers high sgd)") +
  scale_color_fish(option = "Antennarius_multiocellatus", direction = -1) +
  geom_point(data = (seep_dist %>% filter(CowTagID == "VSEEP")), aes(x = lon, y = lat), shape = 21, size = 4, color = "#F1E700") +
  theme_minimal()
```


```{r add pc to model data}
model_data <- left_join(model_data, nut_no_seep_scaled[,c("CowTagID", "pc1", "sd_pc1", "sd_pc2")], by = join_by(CowTagID))
```

## Growth

```{r}
PAC_model_data <- model_data %>%
  filter(Species == "Pocillopora acuta", !is.na(Pin_Number))

PRU_model_data <- left_join(PRU_response_data, nut_no_seep_scaled, by = join_by(CowTagID)) %>% 
  filter(!is.na(Percent_Change))

PRU_model_data_caged_only <- left_join(PRU_response_data_caged, nut_no_seep_scaled, by = join_by(CowTagID)) %>% 
  filter(!is.na(Percent_Change))
```



### Pocillopora acuta

```{r}
ggplot(PAC_model_data, aes(x = pc1, y = Percent_Change, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PCA - low numbers reflect high SGD influence", y = "Percent Change in Buoyant Weight", title = "Pocillopora acuta change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

#### linear model:

```{r pulse linear growth}
linear_growth_model_pac <- lmer(Percent_Change ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(linear_growth_model_pac)
#summary(linear_growth_model_pac)
Anova(linear_growth_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta


#### polynomial model: 

```{r}
polynomial_growth_model_pac <- lmer(log10(Percent_Change) ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(polynomial_growth_model_pac)
Anova(polynomial_growth_model_pac)
```


```{r}
prediction_data <- data.frame(pc1 = seq(min(PAC_model_data$pc1), 
                                   max(PAC_model_data$pc1), 
                                   length.out = 100))
# Add predicted values to the dataframe using the model without random effects
predictions <- predict(polynomial_growth_model_pac, newdata = prediction_data, re.form = NA, se.fit = TRUE)

#Add to new_data
prediction_data <- prediction_data %>%
  mutate(fit = predictions$fit,
         se.fit = predictions$se.fit,
         lower = fit - 1.96 * se.fit,  # 95% CI lower
         upper = fit + 1.96 * se.fit)   # 95% CI upper

# Create the plot with the fitted line
ggplot() +
  geom_point(data = PAC_model_data, aes(x = pc1, y = Percent_Change)) +
  geom_line(data = prediction_data, aes(x = pc1, y = fit), color = "blue") +  # Add the fitted line
  geom_ribbon(data = prediction_data, aes(x = pc1, ymin = lower, ymax = upper), alpha = 0.2, fill = "blue") +  # Add CI
  labs(title = "Pocillopora acuta Δ13C (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ13C") +
  theme_minimal()
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta

### Porites rus

```{r}
ggplot(PRU_model_data, aes(x = pc1, y = Percent_Change, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PCA - low numbers reflect high SGD influence", y = "Percent Change in Buoyant Weight", title = "Porites rus change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

```{r}
linear_growth_model_pru <- lmer(Percent_Change ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(linear_growth_model_pru)
Anova(linear_growth_model_pru)
```

We do not find support for pc1 affecting growth of Porites rus in a linear fashion

```{r}
polynomial_growth_model_pru <- lmer(Percent_Change ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_growth_model_pru)
```

We do not find support for pc1 affecting growth of Porites rus in a polynomial fashion

#### PRU but only caged

I also ran these models on a dataset with only caged corals:

```{r}
linear_growth_model_pru_c <- lmer(Percent_Change ~ pc1 +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(linear_growth_model_pru_c)
Anova(linear_growth_model_pru_c)
```

```{r}
polynomial_growth_model_pru_c <- lmer(Percent_Change ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(polynomial_growth_model_pru_c)
Anova(polynomial_growth_model_pru_c)
```

These model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta

When only looking at caged samples, we still don't find any influence of PC1 on calcification

## Symbiont Count

### Pocillopora acuta

```{r}
ggplot(PAC_model_data, aes(x = pc1, y = log10(FSC.Events_per_cm_2), color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PCA - low numbers reflect high SGD influence", y = "Log10 transformed Symbiont Abundance", title = "Pocillopora acuta change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

#### linear model
```{r}
linear_symb_model_pac <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(linear_symb_model_pac)
Anova(linear_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

#### polynomial model

```{r}
polynomial_symb_model_pac <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(polynomial_symb_model_pac)
Anova(polynomial_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

### Porites rus
```{r}
ggplot(PRU_model_data, aes(x = pc1, y = log10(FSC.Events_per_cm_2), color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PCA - low numbers reflect high SGD influence", y = "Log10 transformed Symbiont Abundance", title = "Porites rus change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

#### linear model:

```{r}
linear_symb_model_pru <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(linear_symb_model_pru)
Anova(linear_symb_model_pru)
```

We do not find support for pc1 affecting symb of Porites rus in a linear fashion

#### polynomial model:

```{r}
polynomial_symb_model_pru <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(polynomial_symb_model_pru)
Anova(polynomial_symb_model_pru)
```

We do not find support for pc1 affecting symb of Porites rus in a polynomial fashion

#### PRU but only caged

```{r}
linear_symb_model_pru_c <- lmer(log10(FSC.Events_per_cm_2) ~ pc1 +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(linear_symb_model_pru_c)
Anova(linear_symb_model_pru_c)
```


```{r}
polynomial_symb_model_pru_c <- lmer(log10(FSC.Events_per_cm_2) ~ poly(pc1, 2) +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(polynomial_symb_model_pru_c)
Anova(polynomial_symb_model_pru_c)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Porites rus

## Heterotrophy

### Exploratory plots
```{r}
SI_Zoop$Species <- SI_Zoop$POM_Zoop

SI_symb <- model_data[,c("Pin_Number", "Species", "lat", "lon", "δ13C_symb", "δ15N_symb", "pc1")] %>% 
  filter(!is.na(δ13C_symb)) %>% 
  filter(!is.na(pc1)) %>% 
  mutate(δ13C = δ13C_symb) %>% 
  mutate(δ15N = δ15N_symb) %>% 
  mutate(Species = ifelse(Species == "Pocillopora acuta", "Pocillopora acuta symbionts",
         ifelse(Species == "Porites rus", "Porites rus symbionts", NA))) %>% 
  filter(!is.na(δ13C))

SI_coral <- model_data[,c("Pin_Number", "Species", "lat", "lon", "δ13C_host", "δ15N_host", "pc1")] %>% 
  filter(!is.na(δ13C_host)) %>% 
  filter(!is.na(pc1)) %>% 
  mutate(δ13C = δ13C_host) %>% 
  mutate(δ15N = δ15N_host) %>% 
  filter(!is.na(δ15N))
```

to see how the coral host vs symb nitrogen looks against the potential sources (zooplankton and pom)
```{r}
SI_Zoop <- SI_Zoop %>%
  mutate(pc1 = ifelse(Seep_Out == "Out", (max(nut_no_seep_scaled$pc1)+0.5), 
                             ifelse(Seep_Out == "Seep", (min(nut_no_seep_scaled$pc1)-0.5), NA))) %>% 
  mutate(Species = POM_Zoop)

ggplot() +
  geom_jitter(data = SI_coral, aes(x = Species, y = δ13C, color = pc1), shape = 17) +
  geom_jitter(data = SI_symb, aes(x = Species, y = δ13C, color = pc1), shape = 16) +
  geom_jitter(data = SI_Zoop, aes(x = Species, y = δ13C, , color = pc1), shape = 15) +
  scale_color_fish(option = "Coryphaena_hippurus", direction = -1, name = "PCA of SGD nutrients")
```

Instead of coloring by the pc axis, this is using the pc as the x-axis with 13C as the response and colored/shaped by species:

```{r}
ggplot() +
  geom_point(data = SI_coral, aes(x = pc1, y = δ13C, color = Species), shape = 17) +
  geom_point(data = SI_symb, aes(x = pc1, y = δ13C, color = Species), shape = 16) +
  geom_point(data = SI_Zoop, aes(x = pc1, y = δ13C, , color = Species), shape = 15) +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1, name = "Source")
```

to see how the coral host vs symb nitrogen looks against the potential sources (zooplankton and pom)
```{r}
ggplot() +
  geom_jitter(data = SI_coral, aes(x = Species, y = δ15N, color = pc1), shape = 17) +
  geom_jitter(data = SI_symb, aes(x = Species, y = δ15N, color = pc1), shape = 16) +
  geom_jitter(data = SI_Zoop, aes(x = Species, y = δ15N, color = pc1), shape = 15) +
  scale_color_fish(option = "Coryphaena_hippurus", direction = -1, name = "PCA of SGD nutrients")
```

Instead of coloring by the pc axis, this is using the pc as the x-axis with 13C as the response and colored/shaped by species:
```{r}
ggplot() +
  geom_point(data = SI_coral, aes(x = pc1, y = δ15N, color = Species), shape = 17) +
  geom_point(data = SI_symb, aes(x = pc1, y = δ15N, color = Species), shape = 16) +
  geom_point(data = SI_Zoop, aes(x = pc1, y = δ15N, , color = Species), shape = 15) +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1, name = "Source")
```

```{r}
ggplot() +
  geom_point(data = SI_coral, aes(x = δ13C, y = δ15N, color = Species), shape = 17) +
  geom_point(data = SI_symb, aes(x = δ13C, y = δ15N, color = Species), shape = 16) +
  geom_point(data = SI_Zoop, aes(x = δ13C, y = δ15N, , color = Species), shape = 15) +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1, name = "Source")
```

## Carbon

### Pocillopora acuta

#### Exploratory plot:
```{r}
SI_PAC_model_data <- PAC_model_data %>% 
  filter(!is.na(Δ13C))
ggplot(data = SI_PAC_model_data, (aes(y=Δ13C, x= pc1))) +
         geom_point()
```

#### Linear model:
```{r}
linear_carbon_model_pac <- lmer(Δ13C ~ pc1 +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(linear_carbon_model_pac)
#summary(linear_carbon_model_pac)
Anova(linear_carbon_model_pac)
```
Not significant

#### Polynomial model:
```{r}
polynomial_carbon_model_pac <- lmer(Δ13C ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(polynomial_carbon_model_pac)
Anova(polynomial_carbon_model_pac)
```

WE FOUND SOMETHING SIGNIFICANT!!!!
```{r}
prediction_data <- data.frame(pc1 = seq(min(SI_PAC_model_data$pc1), 
                                   max(SI_PAC_model_data$pc1), 
                                   length.out = 100))
# Add predicted values to the dataframe using the model without random effects
predictions <- predict(polynomial_carbon_model_pac, newdata = prediction_data, re.form = NA, se.fit = TRUE)

#Add to new_data
prediction_data <- prediction_data %>%
  mutate(fit = predictions$fit,
         se.fit = predictions$se.fit,
         lower = fit - 1.96 * se.fit,  # 95% CI lower
         upper = fit + 1.96 * se.fit)   # 95% CI upper

# Create the plot with the fitted line
ggplot() +
  geom_point(data = SI_PAC_model_data, aes(x = pc1, y = Δ13C)) +
  geom_line(data = prediction_data, aes(x = pc1, y = fit), color = "blue") +  # Add the fitted line
  geom_ribbon(data = prediction_data, aes(x = pc1, ymin = lower, ymax = upper), alpha = 0.2, fill = "blue") +  # Add CI
  labs(title = "Pocillopora acuta Δ13C (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ13C") +
  theme_minimal()
```


```{r}
SI_PAC_model_data <- SI_PAC_model_data %>%
  mutate(predicted_Δ13C = predict(polynomial_carbon_model_pac))

# Create the plot with the fitted line
ggplot(data = SI_PAC_model_data, aes(x = pc1, y = Δ13C)) +
  geom_point() +
  geom_line(aes(y = predicted_Δ13C), color = "blue") +  # Add the fitted line
  labs(title = "Pocillopora acuta Δ13C (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ13C") +
  theme_minimal()
```

### Porites rus
#### Exploratory plot:
```{r}
SI_PRU_model_data <- PRU_model_data %>% 
  filter(!is.na(Δ13C))
ggplot(data = SI_PRU_model_data, (aes(y=Δ13C, x= pc1))) +
         geom_point()
```

#### Linear model:
```{r}
linear_carbon_model_PRU <- lmer(Δ13C ~ pc1 +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(linear_carbon_model_PRU)
#summary(linear_carbon_model_PRU)
Anova(linear_carbon_model_PRU)
```
Not significant

#### Polynomial model:
```{r}
polynomial_carbon_model_PRU <- lmer(Δ13C ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(polynomial_carbon_model_PRU)
Anova(polynomial_carbon_model_PRU)
```
Not significant

## Nitrogen

### Pocillopora acuta

#### Exploratory plot:
```{r}
SI_PAC_model_data <- PAC_model_data %>% 
  filter(!is.na(Δ15N))
ggplot(data = SI_PAC_model_data, (aes(y=Δ15N, x= pc1))) +
         geom_point()
```

#### Linear model:
```{r}
linear_nitrogen_model_pac <- lmer(Δ15N ~ pc1 +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(linear_nitrogen_model_pac)
#summary(linear_nitrogen_model_pac)
Anova(linear_nitrogen_model_pac)
```
Not significant

#### Polynomial model:
```{r}
polynomial_nitrogen_model_pac <- lmer(Δ15N ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(polynomial_nitrogen_model_pac)
Anova(polynomial_nitrogen_model_pac)
```
Not significant

### Porites rus
#### Exploratory plot:
```{r}
SI_PRU_model_data <- PRU_model_data %>% 
  filter(!is.na(Δ15N))
ggplot(data = SI_PRU_model_data, (aes(y=Δ15N, x= pc1))) +
         geom_point()
```

#### Linear model:
```{r}
linear_nitrogen_model_PRU <- lmer(Δ15N ~ pc1 +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(linear_nitrogen_model_PRU)
#summary(linear_nitrogen_model_PRU)
Anova(linear_nitrogen_model_PRU)
```
Not significant

#### Polynomial model:
```{r}
polynomial_nitrogen_model_PRU <- lmer(Δ15N ~ poly(pc1, 2) +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(polynomial_nitrogen_model_PRU)
Anova(polynomial_nitrogen_model_PRU)
```

Significant

```{r}
prediction_data <- data.frame(pc1 = seq(min(SI_PRU_model_data$pc1), 
                                   max(SI_PRU_model_data$pc1), 
                                   length.out = 100))
# Add predicted values to the dataframe using the model without random effects
predictions <- predict(polynomial_nitrogen_model_PRU, newdata = prediction_data, re.form = NA, se.fit = TRUE)

#Add to new_data
prediction_data <- prediction_data %>%
  mutate(fit = predictions$fit,
         se.fit = predictions$se.fit,
         lower = fit - 1.96 * se.fit,  # 95% CI lower
         upper = fit + 1.96 * se.fit)   # 95% CI upper

# Create the plot with the fitted line
ggplot() +
  geom_point(data = SI_PRU_model_data, aes(x = pc1, y = Δ15N), color = "grey") +
  geom_line(data = prediction_data, aes(x = pc1, y = fit), color = "#21908c") +  # Add the fitted line
  geom_ribbon(data = prediction_data, aes(x = pc1, ymin = lower, ymax = upper), alpha = 0.2, fill = "#21908c") +  # Add CI
  labs(title = "Porites rus Δ15N (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ15N") +
  theme_minimal()
```

```{r}
prediction_data <- data.frame(pc1 = seq(min(SI_PRU_model_data$pc1), 
                                   max(SI_PRU_model_data$pc1), 
                                   length.out = 100))
# Add predicted values to the dataframe using the model without random effects
predictions <- predict(polynomial_nitrogen_model_PRU, newdata = prediction_data, re.form = NA, se.fit = TRUE)

#Add to new_data
prediction_data <- prediction_data %>%
  mutate(fit = predictions$fit,
         se.fit = predictions$se.fit,
         lower = fit - 1.96 * se.fit,  # 95% CI lower
         upper = fit + 1.96 * se.fit)   # 95% CI upper

# Create the plot with the fitted line
ggplot() +
  geom_point(data = SI_PAC_model_data, aes(x = pc1, y = Δ13C)) +
  geom_line(data = prediction_data, aes(x = pc1, y = fit), color = "blue") +  # Add the fitted line
  geom_ribbon(data = prediction_data, aes(x = pc1, ymin = lower, ymax = upper), alpha = 0.2, fill = "blue") +  # Add CI
  labs(title = "Pocillopora acuta Δ13C (host minus symbiont, proxy for heterotrophy)",
       x = "PC1",
       y = "Δ13C") +
  theme_minimal()
```


# Question 2

## Collinearity
```{r}
chart.Correlation(na.omit(model_data[,c("sd_Salinity","sd_Temperature", "sd_TA", "sd_pH")]))
```


## Growth

### PAC
```{r PAC lasso}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("Percent_Change","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  mutate(sd_TA = scale(sd_TA))%>% 
  mutate(sd_pH = scale(sd_pH))%>% 
  mutate(sd_Salinity = scale(sd_Salinity))%>% 
  mutate(sd_Temperature = scale(sd_Temperature))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)
PAC_lasso$Cage_Uncaged <- as.factor(PAC_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_percent_change <- PAC_model_data_matrix[, -1]
y_pac_percent_change <- PAC_model_data_matrix[, 1]
lasso_PAC_growth <- cv.glmnet(x = x_pac_percent_change, y = y_pac_percent_change)
coef(lasso_PAC_growth)
```

#### Significant model

```{r}
multivariate_PAC_growth <- lmer(Percent_Change ~ sd_Temperature + sd_TA + sd_pH + 
  (1 | Genotype) + (1 | Cage_Uncaged),
  data = PAC_model_data
)
summary(multivariate_PAC_growth)
car::Anova(multivariate_PAC_growth)
plot(allEffects(multivariate_PAC_growth))

#check_model(multivariate_PAC_growth)
```

##### Sig model plot
```{r}
pred_temp1 <- ggpredict(multivariate_PAC_growth, terms = "sd_Temperature")
pred_ta <- ggpredict(multivariate_PAC_growth, terms = "sd_TA")
pred_ph <- ggpredict(multivariate_PAC_growth, terms = "sd_pH")

#annnotations
anova_results <- car::Anova(multivariate_PAC_growth)
p_values <- anova_results$`Pr(>Chisq)`[1:3]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1:3] 
```

```{r}
custom_theme <- theme_minimal(base_size = 10)+
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10)
  )
```

```{r}
# Plot for sd_Temperature
plot_temp <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = sd_Temperature, y = Percent_Change), color = "grey50", alpha = 0.6) +
  # Prediction line with confidence interval
  geom_line(data = pred_temp, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_temp, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "Standard Deviation of Temperature",
    y = "Percent Change in Buoyant Weight"
  ) +
  custom_theme+
  annotate("text", y = 1, x = max(pred_temp$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["sd_Temperature"], 6)), 
           hjust = 1.1, vjust = 1, size = 3)

# Plot for sd_TA
plot_ta <- ggplot() +
  geom_point(data = PAC_model_data, 
             aes(x = sd_TA, y = Percent_Change), color = "grey50", alpha = 0.6) +
  geom_line(data = pred_ta, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_ta, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "Standard Deviation of TA", y = " "
  ) +
  custom_theme+
  annotate("text", y = 1, x =max(pred_ta$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["sd_TA"], 6)), 
           hjust = 1.1, vjust = 1, size = 3)

# Plot for sd_pH
plot_ph <- ggplot() +
  geom_point(data = PAC_model_data, 
             aes(x = sd_pH, y = Percent_Change), color = "grey50", alpha = 0.6) +
  geom_line(data = pred_ph, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_ph, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "Standard Deviation of pH", y = " "
  ) +
  custom_theme+
  annotate("text", y = 1, x =max(pred_ph$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["sd_pH"], 9)), 
           hjust = 1.1, vjust = 1, size = 3)
```

```{r}
combined_plot <- ggarrange(plot_temp + 
                              labs(subtitle = "A") + 
                              theme(plot.subtitle = element_text(hjust = -0.2, size = 12)),
                            plot_ta + 
                              labs(subtitle = "B") + 
                              theme(plot.subtitle = element_text(hjust = -0.2, size = 12)),
                            plot_ph + 
                              labs(subtitle = "C") + 
                              theme(plot.subtitle = element_text(hjust = -0.2, size = 12)),
                            nrow = 1)+ 
  ggtitle(expression("Effects of Variability on Growth in" ~ italic("Pocillopora acuta"))) +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        plot.margin = margin(t = 20, r = 20, b = 20, l = 40))
combined_plot
ggsave(combined_plot, filename = "PAC_Percent_Change_Effects.png", dpi = 600, width = 10, height = 5)
```

### PRU
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("Percent_Change","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)
PRU_lasso$Cage_Uncaged <- as.factor(PRU_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_percent_change <- PRU_model_data_matrix[, -1]
y_PRU_percent_change <- PRU_model_data_matrix[, 1]
lasso_PRU_growth <- cv.glmnet(x = x_PRU_percent_change, y = y_PRU_percent_change)
coef(lasso_PRU_growth)
```

### PRU Caged Only
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso_c <- PRU_model_data_caged_only[,c("Percent_Change","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso_c$Genotype <- as.factor(PRU_lasso_c$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso_c)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso_c <- cbind(PRU_lasso_c, genotype_dummies)
PRU_lasso_c <- PRU_lasso_c %>% dplyr::select(-Genotype)

#lasso works on matrix
PRU_model_data_matrix_c <- as.matrix(PRU_lasso_c)

x_PRU_percent_change_c <- PRU_model_data_matrix_c[, -1]
y_PRU_percent_change_c <- PRU_model_data_matrix_c[, 1]
lasso_PRU_growth_c <- cv.glmnet(x = x_PRU_percent_change_c, y = y_PRU_percent_change_c)
coef(lasso_PRU_growth_c)
```

## Symb Abundance

### PAC
```{r}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("FSC.Events_per_cm_2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

PAC_lasso <- PAC_lasso %>% 
  mutate(FSC.Events_per_cm_2 = log10(FSC.Events_per_cm_2)) 

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)
PAC_lasso$Cage_Uncaged <- as.factor(PAC_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_symb <- PAC_model_data_matrix[, -1]
y_pac_symb <- PAC_model_data_matrix[, 1]
lasso_PAC_symb <- cv.glmnet(x = x_pac_symb, y = y_pac_symb)
coef(lasso_PAC_symb)
```

### PRU
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("FSC.Events_per_cm_2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

PRU_lasso <- PRU_lasso %>% 
  mutate(FSC.Events_per_cm_2 = log10(FSC.Events_per_cm_2)) %>% 
  filter(!is.na(FSC.Events_per_cm_2))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)
PRU_lasso$Cage_Uncaged <- as.factor(PRU_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_symb <- PRU_model_data_matrix[, -1]
y_PRU_symb <- PRU_model_data_matrix[, 1]
lasso_PRU_symb <- cv.glmnet(x = x_PRU_symb, y = y_PRU_symb)
coef(lasso_PRU_symb)
```

#### PRU Caged Only
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data_caged_only[,c("FSC.Events_per_cm_2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

PRU_lasso <- PRU_lasso %>% 
  mutate(FSC.Events_per_cm_2 = log10(FSC.Events_per_cm_2)) %>% 
  filter(!is.na(FSC.Events_per_cm_2))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_symb <- PRU_model_data_matrix[, -1]
y_PRU_symb <- PRU_model_data_matrix[, 1]
lasso_PRU_symb <- cv.glmnet(x = x_PRU_symb, y = y_PRU_symb)
coef(lasso_PRU_symb)
```

whole lotta nothing

## Heterotrophy

### Carbon

#### PAC
```{r}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("Δ13C","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ13C))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_13c <- PAC_model_data_matrix[, -1]
y_pac_13c <- PAC_model_data_matrix[, 1]
lasso_PAC_13c <- cv.glmnet(x = x_pac_13c, y = y_pac_13c)
coef(lasso_PAC_13c)
```

#### PRU

```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("Δ13C","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ13C))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_13c <- PRU_model_data_matrix[, -1]
y_PRU_13c <- PRU_model_data_matrix[, 1]
lasso_PRU_13c <- cv.glmnet(x = x_PRU_13c, y = y_PRU_13c)
coef(lasso_PRU_13c)
```

nothin again

### Nitrogen

#### PAC
```{r}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("Δ15N","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ15N))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_15N <- PAC_model_data_matrix[, -1]
y_pac_15N <- PAC_model_data_matrix[, 1]
lasso_PAC_15N <- cv.glmnet(x = x_pac_15N, y = y_pac_15N)
coef(lasso_PAC_15N)
```

#### PRU

```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("Δ15N","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ15N))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_15N <- PRU_model_data_matrix[, -1]
y_PRU_15N <- PRU_model_data_matrix[, 1]
lasso_PRU_15N <- cv.glmnet(x = x_PRU_15N, y = y_PRU_15N)
coef(lasso_PRU_15N)
```


# Playing with Danielle Data
```{r}
sp_data <- read.csv("data/Species_Abundances_wide.csv")
danielle <- read.csv("data/Full_Metadata.csv")

sp_data_pc <- left_join(sp_data, nut_no_seep_scaled[,c("CowTagID", "pc1", "sd_TA")],by = join_by(CowTagID))
sp_data_pc <- left_join(sp_data_pc, danielle[,c("CowTagID", "LiveCoral")])
sp_data_pc$Rusperlivecoral <- sp_data_pc$Porites.rus/sp_data_pc$LiveCoral
sp_data_pc$Acutaperlivecoral <- sp_data_pc$Pocillopora.acuta/sp_data_pc$LiveCoral


ggplot(sp_data_pc) +
  geom_line(aes(y=Porites.rus, x = pc1), color = "blue") +
  geom_line(aes(y=Pocillopora.acuta, x = pc1))

ggplot(sp_data_pc) +
  geom_line(aes(y=Rusperlivecoral, x = sd_TA), color = "blue") +
  geom_line(aes(y=Acutaperlivecoral, x = sd_TA))
```

# quick make a table:
```{r}
ltm_table <- nut_no_seep[,c("CowTagID", pulse_columns)]

# Select relevant columns and calculate summary statistics
ltm_table_sum <- data.frame(
  Variable = c("Phosphate_umolL", "Silicate_umolL", "NN_umolL", "NNP_umolL", "Salinity"),
  Min = sapply(ltm_table[, c("Low_Tide_Mean_Phosphate_umolL", "Low_Tide_Mean_Silicate_umolL", 
                             "Low_Tide_Mean_NN_umolL", "Low_Tide_Mean_NNP_umolL",
                             "Low_Tide_Mean_Salinity")], min, na.rm = TRUE),
  Max = sapply(ltm_table[, c("Low_Tide_Mean_Phosphate_umolL", "Low_Tide_Mean_Silicate_umolL", 
                             "Low_Tide_Mean_NN_umolL", "Low_Tide_Mean_NNP_umolL",
                             "Low_Tide_Mean_Salinity")], max, na.rm = TRUE),
  Mean = sapply(ltm_table[, c("Low_Tide_Mean_Phosphate_umolL", "Low_Tide_Mean_Silicate_umolL", 
                              "Low_Tide_Mean_NN_umolL", "Low_Tide_Mean_NNP_umolL",
                              "Low_Tide_Mean_Salinity")], mean, na.rm = TRUE)
)
```

```{r}
sd_table <- nut_no_seep[,sd_columns]

sd_table_sum <- data.frame(
  Variable = c("Salinity", "Temperature", "Total Alkalinity", "pH"),
  Min = sapply(sd_table[, c(sd_columns)], min, na.rm = TRUE),
  Max = sapply(sd_table[, c(sd_columns)], max, na.rm = TRUE),
  Mean = sapply(sd_table[, c(sd_columns)], mean, na.rm = TRUE)
)
```

