---
title: "LTM_Silicate_Results"
author: "Callie Stephenson"
date: "2024-09-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
```

#### Loading the packages and data
```{r load packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(fishualize)
library(lme4)
library(lmerTest) #figure out which lme package you need
library(ggplot2)
library(performance)
library(see)
library(dplyr)
library(car)
```

```{r load data, echo=FALSE, message=FALSE}
setwd("/Users/calliestephenson/Documents/GitHub/MooreaCoralFish_CS")

#explanatory data
all_nut <- read.csv("data/All_Nutrients_Processed.csv")
shore_dist <- read.csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Shore_distance.csv") #using Danielle's
seep_dist <- read.csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Plate_Distance_to_Seep.csv") %>% 
  filter(Location == "Varari")
depth <- read.csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/Adj_Sandwich_Depth.csv")
clod <- read.csv("data/Clod_Cards.csv")
turb <- read_csv("https://raw.githubusercontent.com/dbarnas/Community_Functional_Diversity/main/Data/Biogeochem/August2021/Turb_NC.csv")

#Response data
response_data <- read.csv("data/response_data.csv")
response_data$Genotype <- as.factor(response_data$Genotype)
response_data$Cage_Uncaged <- as.factor(response_data$Cage_Uncaged)
response_data$Pin_Number <- as.factor(response_data$Pin_Number)

PRU_response_data <- response_data %>% 
  filter(Species == "Porites rus")

PRU_response_data_caged <- PRU_response_data %>% 
  filter(Cage_Uncaged == "C")

PAC_response_data <- response_data %>% 
  filter(Species == "Pocillopora acuta")

PAC_response_data_caged <- PAC_response_data %>% 
  filter(Cage_Uncaged == "C")

#metadata
meta <- read.csv("data/coral_metadata.csv")

#SI Stuff
SI_Zoop <- read.csv("data/SI_Zoop.csv")%>% 
  mutate(δ13C = δ13C............vs..VPDB.) %>% 
  mutate(δ15N = δ15N...........vs..AIR.) %>% 
  dplyr::select(-'δ13C............vs..VPDB.')%>% 
  dplyr::select(-'δ15N...........vs..AIR.')
```

```{r}
nut_no_seep <- all_nut[all_nut$CowTagID != 'VSEEP',]
nut_no_seep <- left_join(nut_no_seep, seep_dist[,c("CowTagID", "lat", "lon", "dist_to_seep_m")], by = join_by(CowTagID))
nut_no_seep <- left_join(nut_no_seep, shore_dist[,c("CowTagID", "shore_dist_m")], by = join_by(CowTagID))
```

```{r model data}
model_data <- left_join(response_data, nut_no_seep, by = join_by(CowTagID))
```


```{r}
ggplot(nut_no_seep, aes(x = lon, y = lat, color = Low_Tide_Mean_Silicate_umolL)) +
  geom_point() +
  labs(x = "Latitude", y = "Longitude", color = "Low_Tide_Mean_Silicate_umolL") +
  scale_color_fish(option = "Coryphaena_hippurus", direction = -1) +
  geom_point(data = (seep_dist %>% filter(CowTagID == "VSEEP")), aes(x = lon, y = lat), shape = 21, size = 4, color = "#F1E700") +
  theme_minimal()
```

## Growth

```{r}
PAC_model_data <- model_data %>%
  filter(Species == "Pocillopora acuta", !is.na(Pin_Number))

PRU_model_data <- left_join(PRU_response_data, nut_no_seep, by = join_by(CowTagID)) %>% 
  filter(!is.na(FSC.Events_per_cm_2))

PRU_model_data_caged_only <- left_join(PRU_response_data_caged, nut_no_seep, by = join_by(CowTagID)) %>% 
  filter(!is.na(FSC.Events_per_cm_2))
```


### Pocillopora acuta

```{r}
ggplot(PAC_model_data, aes(x = Low_Tide_Mean_Silicate_umolL, y = FSC.Events_per_cm_2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "Low tide mean silicate - high numbers = high SGD", y = "Percent Change in Buoyant Weight", title = "Pocillopora acuta change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

#### linear model:

```{r}
linear_growth_model_pac <- lmer(FSC.Events_per_cm_2 ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(linear_growth_model_pac)
#summary(linear_growth_model_pac)
Anova(linear_growth_model_pac)
```

This model does not support the hypothesis SGD contributes to the variance in calcification in Pocillopora acuta

#### polynomial model: 

```{r}
polynomial_growth_model_pac <- lmer(FSC.Events_per_cm_2 ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(polynomial_growth_model_pac)
Anova(polynomial_growth_model_pac)
```

This model does not support the hypothesis SGD contributes to the variance in calcification in Pocillopora acuta

### Porites rus

```{r}
ggplot(PRU_model_data, aes(x = Low_Tide_Mean_Silicate_umolL, y = FSC.Events_per_cm_2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PCA - low numbers reflect high SGD influence", y = "Percent Change in Buoyant Weight", title = "Porites rus change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

```{r}
linear_growth_model_pru <- lmer(FSC.Events_per_cm_2 ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(linear_growth_model_pru)
Anova(linear_growth_model_pru)
```

We do not find support for Low_Tide_Mean_Silicate_umolL affecting growth of Porites rus in a linear fashion

```{r}
polynomial_growth_model_pru <- lmer(FSC.Events_per_cm_2 ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(polynomial_growth_model_pru)
Anova(polynomial_growth_model_pru)
```

We do not find support for Low_Tide_Mean_Silicate_umolL affecting growth of Porites rus in a polynomial fashion

#### PRU but only caged

I also ran these models on a dataset with only caged corals:

```{r}
linear_growth_model_pru_c <- lmer(FSC.Events_per_cm_2 ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(linear_growth_model_pru_c)
Anova(linear_growth_model_pru_c)
```

```{r}
polynomial_growth_model_pru_c <- lmer(FSC.Events_per_cm_2 ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(polynomial_growth_model_pru_c)
Anova(polynomial_growth_model_pru_c)
```

These model does not support the hypothesis that nutriencts from SGD contribute to the variance in calcification in Pocillopora acuta

When only looking at caged samples, we still don't find any influence of Low_Tide_Mean_Silicate_umolL on calcification

## Symbiont Count

### Pocillopora acuta

```{r}
ggplot(PAC_model_data, aes(x = Low_Tide_Mean_Silicate_umolL, y = FSC.Events_per_cm_2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PCA - low numbers reflect high SGD influence", y = "Percent Change in Buoyant Weight", title = "Pocillopora acuta change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

#### linear model
```{r}
linear_symb_model_pac <- lmer(FSC.Events_per_cm_2 ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(linear_symb_model_pac)
Anova(linear_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

#### polynomial model

```{r}
polynomial_symb_model_pac <- lmer(FSC.Events_per_cm_2 ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PAC_model_data)
#check_model(polynomial_symb_model_pac)
Anova(polynomial_symb_model_pac)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Pocillopora acuta

### Porites rus
```{r}
ggplot(PRU_model_data, aes(x = Low_Tide_Mean_Silicate_umolL, y = FSC.Events_per_cm_2, color = Genotype, shape = Cage_Uncaged)) +
  geom_point() +
  labs(x = "PCA - low numbers reflect high SGD influence", y = "Symbiont Density in Buoyant Weight", title = "Porites rus change in buoyant weight by SGD exposure") +
  scale_color_fish_d(option = "Coryphaena_hippurus", direction = -1) +
  theme_minimal()
```

#### linear model:

```{r}
linear_symb_model_pru <- lmer(FSC.Events_per_cm_2 ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(linear_symb_model_pru)
Anova(linear_symb_model_pru)
```

We do not find support for Low_Tide_Mean_Silicate_umolL affecting symb of Porites rus in a linear fashion

#### polynomial model:

```{r}
polynomial_symb_model_pru <- lmer(FSC.Events_per_cm_2 ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype) + (1|Cage_Uncaged), data = PRU_model_data)
#check_model(polynomial_symb_model_pru)
Anova(polynomial_symb_model_pru)
```

We do not find support for Low_Tide_Mean_Silicate_umolL affecting symb of Porites rus in a polynomial fashion

#### PRU but only caged

```{r}
linear_symb_model_pru_c <- lmer(FSC.Events_per_cm_2 ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(linear_symb_model_pru_c)
Anova(linear_symb_model_pru_c)
```


```{r}
polynomial_symb_model_pru_c <- lmer(FSC.Events_per_cm_2 ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype), data = PRU_model_data_caged_only)
#check_model(polynomial_symb_model_pru_c)
Anova(polynomial_symb_model_pru_c)
```

This model does not support the hypothesis that nutriencts from SGD contribute to the variance in symbiont density in Porites rus

## Heterotrophy

```{r}
SI_Zoop$Species <- SI_Zoop$POM_Zoop
SI_Zoop <- SI_Zoop %>%
  mutate(Low_Tide_Mean_Silicate_umolL = ifelse(Seep_Out == "Seep", max(nut_no_seep$Low_Tide_Mean_Silicate_umolL), 
                             ifelse(Seep_Out == "Out", min(nut_no_seep$Low_Tide_Mean_Silicate_umolL), NA)))

SI_symb <- model_data[,c("Pin_Number", "Species", "lat", "lon", "δ13C_symb", "δ15N_symb", "Low_Tide_Mean_Silicate_umolL")] %>% 
  filter(!is.na(δ13C_symb)) %>% 
  filter(!is.na(Low_Tide_Mean_Silicate_umolL)) %>% 
  mutate(δ13C = δ13C_symb) %>% 
  mutate(δ15N = δ15N_symb) %>% 
  mutate(Species = ifelse(Species == "Pocillopora acuta", "Pocillopora acuta symbionts",
         ifelse(Species == "Porites rus", "Porites rus symbionts", NA))) %>% 
  filter(!is.na(δ13C))

SI_coral <- model_data[,c("Pin_Number", "Species", "lat", "lon", "δ13C_host", "δ15N_host", "Low_Tide_Mean_Silicate_umolL")] %>% 
  filter(!is.na(δ13C_host)) %>% 
  filter(!is.na(Low_Tide_Mean_Silicate_umolL)) %>% 
  mutate(δ13C = δ13C_host) %>% 
  mutate(δ15N = δ15N_host) %>% 
  filter(!is.na(δ15N))
```

to see how the coral host vs symb nitrogen looks against the potential sources (zooplankton and pom)
```{r}
ggplot() +
  geom_jitter(data = SI_coral, aes(x = Species, y = δ13C, color = Low_Tide_Mean_Silicate_umolL), shape = 17) +
  geom_jitter(data = SI_symb, aes(x = Species, y = δ13C, color = Low_Tide_Mean_Silicate_umolL), shape = 16) +
  geom_jitter(data = SI_Zoop, aes(x = Species, y = δ13C, , color = Low_Tide_Mean_Silicate_umolL), shape = 15) +
  scale_color_fish(option = "Coryphaena_hippurus", direction = -1, name = "Silicate gradient")
```

to see how the coral host vs symb nitrogen looks against the potential sources (zooplankton and pom)
```{r}
ggplot() +
  geom_jitter(data = SI_coral, aes(x = Species, y = δ15N, color = Low_Tide_Mean_Silicate_umolL), shape = 17) +
  geom_jitter(data = SI_symb, aes(x = Species, y = δ15N, color = Low_Tide_Mean_Silicate_umolL), shape = 16) +
  geom_jitter(data = SI_Zoop, aes(x = Species, y = δ15N, color = Low_Tide_Mean_Silicate_umolL), shape = 15) +
  scale_color_fish(option = "Coryphaena_hippurus", direction = -1, name = "PCA of SGD nutrients")
```

## Carbon

### Pocillopora acuta

#### Exploratory plot:
```{r}
SI_PAC_model_data <- PAC_model_data %>% 
  filter(!is.na(Δ13C))
ggplot(data = SI_PAC_model_data, (aes(y=Δ13C, x= Low_Tide_Mean_Silicate_umolL))) +
         geom_point()
```

#### Linear model:
```{r}
linear_carbon_model_pac <- lmer(Δ13C ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(linear_carbon_model_pac)
#summary(linear_carbon_model_pac)
Anova(linear_carbon_model_pac)
```
Not significant

#### Polynomial model:
```{r}
polynomial_carbon_model_pac <- lmer(Δ13C ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(polynomial_carbon_model_pac)
Anova(polynomial_carbon_model_pac)
```

WE FOUND SOMETHING SIGNIFICANT!!!!

```{r}
SI_PAC_model_data <- SI_PAC_model_data %>%
  mutate(predicted_Δ13C = predict(polynomial_carbon_model_pac))

# Create the plot with the fitted line
ggplot(data = SI_PAC_model_data, aes(x = Low_Tide_Mean_Silicate_umolL, y = Δ13C)) +
  geom_point() +
  geom_line(aes(y = predicted_Δ13C), color = "blue") +  # Add the fitted line
  labs(title = "Pocillopora acuta Δ13C (host minus symbiont, proxy for heterotrophy)",
       x = "Low_Tide_Mean_Silicate_umolL",
       y = "Δ13C") +
  theme_minimal()
```

### Porites rus
#### Exploratory plot:
```{r}
SI_PRU_model_data <- PRU_model_data %>% 
  filter(!is.na(Δ13C))
ggplot(data = SI_PRU_model_data, (aes(y=Δ13C, x= Low_Tide_Mean_Silicate_umolL))) +
         geom_point()
```

#### Linear model:
```{r}
linear_carbon_model_PRU <- lmer(Δ13C ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(linear_carbon_model_PRU)
#summary(linear_carbon_model_PRU)
Anova(linear_carbon_model_PRU)
```
Not significant

#### Polynomial model:
```{r}
polynomial_carbon_model_PRU <- lmer(Δ13C ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(polynomial_carbon_model_PRU)
Anova(polynomial_carbon_model_PRU)
```
Not significant

## Nitrogen

### Pocillopora acuta

#### Exploratory plot:
```{r}
SI_PAC_model_data <- PAC_model_data %>% 
  filter(!is.na(Δ15N))
ggplot(data = SI_PAC_model_data, (aes(y=Δ15N, x= Low_Tide_Mean_Silicate_umolL))) +
         geom_point()
```

#### Linear model:
```{r}
linear_nitrogen_model_pac <- lmer(Δ15N ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(linear_nitrogen_model_pac)
#summary(linear_nitrogen_model_pac)
Anova(linear_nitrogen_model_pac)
```
Not significant

#### Polynomial model:
```{r}
polynomial_nitrogen_model_pac <- lmer(Δ15N ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype), data = SI_PAC_model_data)
#check_model(polynomial_nitrogen_model_pac)
Anova(polynomial_nitrogen_model_pac)
```
Not significant

### Porites rus
#### Exploratory plot:
```{r}
SI_PRU_model_data <- PRU_model_data %>% 
  filter(!is.na(Δ15N))
ggplot(data = SI_PRU_model_data, (aes(y=Δ15N, x= Low_Tide_Mean_Silicate_umolL))) +
         geom_point()
```

#### Linear model:
```{r}
linear_nitrogen_model_PRU <- lmer(Δ15N ~ Low_Tide_Mean_Silicate_umolL +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(linear_nitrogen_model_PRU)
#summary(linear_nitrogen_model_PRU)
Anova(linear_nitrogen_model_PRU)
```
Not significant

#### Polynomial model:
```{r}
polynomial_nitrogen_model_PRU <- lmer(Δ15N ~ poly(Low_Tide_Mean_Silicate_umolL, 2) +
                                  (1|Genotype), data = SI_PRU_model_data)
#check_model(polynomial_nitrogen_model_PRU)
Anova(polynomial_nitrogen_model_PRU)
```

Not significant


