---
title: "Fish_Ordination"
author: "Callie Stephenson"
date: "2025-03-15"
output: html_document
---
```{r}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Set-Up
```{r packages, message=FALSE}
library(dplyr)
library(readr)
library(here)
library(fishualize)
library(ggplot2)
library(vegan)
library(viridis)
library(tidyr)
library(stringr)
library(lme4)
library(mgcv)
library(gridExtra)
library(forcats) #needed for reversing factor levels
library(cowplot)
```

## Explanatory Data
```{r explanatory data}
setwd("/Users/calliestephenson/Documents/GitHub/MooreaCoralFish_CS")

explan<- read.csv("data/explanatory_all_wide.csv") %>% 
  mutate(Pin = as.numeric(sub("V", "", CowTagID)))
db <- read.csv("data/DB_Full_Metadata.csv") %>% 
  mutate_all(~ifelse(is.na(.), 0, .))#Rugosity data pulled from Danielle Barnas' measurements
turb <- read.csv("data/Turb_NC.csv")
```

CS added 6/19
```{r}
resid_explan <- read.csv(here("Data", "explan_complexity_resid.csv"))
```

### adding in turb and rugosity
```{r}
# explan <- left_join(explan, db[,c("CowTagID", "meanRugosity", "dist_to_seep_m")])
# explan <- left_join(explan, turb[,c("CowTagID", "del15N", "N_percent")])
water_quality_explan <- explan[,c("CowTagID", "dist_to_seep_m", "shore_dist_m", "Low_Tide_Mean_NN_umolL", "Low_Tide_Mean_Silicate_umolL", "Low_Tide_Mean_Phosphate_umolL", "Low_Tide_Mean_pH", "Low_Tide_Mean_TA", "Low_Tide_Mean_Salinity", "pulse_pc1")]

indirect_explan <- left_join(explan[,c("CowTagID", "dist_to_seep_m", "shore_dist_m")], turb[,c("CowTagID", "del15N", "N_percent", "C_percent", "C_N", "del13C")])
indirect_explan <- left_join(indirect_explan, db[,c("CowTagID", "meanRugosity")])
indirect_explan <- left_join(indirect_explan, water_quality_explan[,c("CowTagID", "pulse_pc1")])

#Putting them together
full_explan <- left_join(indirect_explan, water_quality_explan)
```

### Cleaning up explan
```{r}
indirect_explan <- indirect_explan %>%
   mutate(Pin = as.integer(sub("V", "", CowTagID)))
# explan <- left_join(map_data_join[,c("Pin", "alphabetical")], explan, by = join_by(Pin))
# explan$Pin <- explan$alphabetical 

water_quality_explan <- water_quality_explan %>%
   mutate(Pin = as.integer(sub("V", "", CowTagID)))
```

thinking about correlations for models:
```{r}
direct_cor_matrix <- cor(na.omit(water_quality_explan[,c(2:9)]))
direct_corplot <- ggcorrplot(direct_cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)

indirect_cor_matrix <- cor(na.omit(indirect_explan[,c(2:10)]))
indirect_corplot <- ggcorrplot(indirect_cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)
```

```{r}
numeric_cols <- sapply(full_explan, is.numeric)
all_cor_matrix <- cor(na.omit(full_explan[, numeric_cols]))
direct_corplot <- ggcorrplot(all_cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)
```


## Response Data
```{r}
fish_tidy <- read.csv("data/fish_tidy.csv")
functional_groups <- read.csv("data/fish_functional_groups.csv") %>% 
  mutate(Herbivore = ifelse(Fine_Trophic %in% c("Herbivore/Detritivore","Primary Consumer","Scraper","Cropper", 
                                                "Excavator", "Brusher", "Browser","Concealed Cropper", "Omnivore"), "Y", "N"))
```

```{r}
herbivores <- fish_tidy %>%
  left_join(functional_groups[,c("Species", "Fine_Trophic", "Herbivore")], by = join_by(Species)) %>% 
  filter(Herbivore == "Y") %>% 
#  filter(!(Pin == "14" & Block == "On 8")) %>%  #This block only has one chub and it's messing with the ordination
  filter(!(Pin == "14")) #This removes pin 14 entirely
```

# Alpha Diversity
```{r}
species_richness <- fish_tidy %>%
#  filter(!Pin == "13") %>% #Already removed
  filter(!Pin == "A") %>% #Pin 14
  group_by(Pin, On.Adjacent, Block) %>%
  summarise(Richness = n_distinct(Species[MaxN > 0]))
```

# Ordination
```{r}
#make it wide
wide_maxn_df <- pivot_wider(fish_tidy, 
                             id_cols = c("Pin", "On.Adjacent", "Block"),
                             names_from = Species, 
                             values_from = MaxN) %>% 
  mutate_all(~ifelse(is.na(.), 0, .))

#make the version without 14 wide
wide_herb_df <- pivot_wider(herbivores, 
                             id_cols = c("Pin", "On.Adjacent", "Block"),
                             names_from = Species, 
                             values_from = MaxN) %>% 
  mutate_all(~ifelse(is.na(.), 0, .))

speciescolumms <- names(wide_maxn_df[,c(4:ncol(wide_maxn_df))])
speciesdata <- wide_maxn_df[,c(4:ncol(wide_maxn_df))]

herbspeciescolumns <-names(wide_herb_df[,c(4:ncol(wide_herb_df))])
herbspeciesdata <- wide_herb_df[,c(4:ncol(wide_herb_df))]
```

```{r}
Fish_trans <- decostand(speciesdata, method = "max")
non_zero_rows <- rowSums(Fish_trans == 0) != ncol(Fish_trans)
species.data.use <- Fish_trans[non_zero_rows, ]
set.seed(123)
ord = metaMDS(species.data.use, dist = "bray", trymax = 500, k = 3, trace = FALSE)
ord
stressplot(ord) ##stressplot is just a diagnostic for your data, shows how your datapoints are distributed and should be somewhat linear
##Stress Value = how well doe your NMDS represent reality (<0.2)
plot(ord, type = "text")
```

```{r}
sppscores(ord) <- Fish_trans #noam says you have to add the species back in
```

## Envfit

### Turb
```{r}
explan_envfit <- left_join(wide_maxn_df, 
                          # explan[, c("Pin", "del15N", "N_percent", "meanRugosity", "shore_dist_m", "dist_to_seep_m")], 
                          indirect_explan[,c("Pin", "del15N", "N_percent", "C_N", "C_percent", "del13C", "meanRugosity", "dist_to_seep_m", "shore_dist_m")],
                           by = "Pin")

# indirect_fit <- envfit(ord ~ del15N + N_percent + shore_dist_m + dist_to_seep_m + meanRugosity, data = explan_envfit, na.rm = T, permutations = 999)
indirect_fit <- envfit(ord ~ shore_dist_m + dist_to_seep_m + meanRugosity + del15N + N_percent + C_percent + C_N + del13C, data = explan_envfit, na.rm = T, permutations = 999)
#remove non-significant del13c
indirect_fit <- envfit(ord ~ shore_dist_m + dist_to_seep_m + meanRugosity + del15N + N_percent + C_percent + C_N, data = explan_envfit, na.rm = T, permutations = 999)
indirect_fit

#to get r2:
indirect_fit$vectors$r

indirect_r2_df <- data.frame(
  Variable = rownames(indirect_fit$vectors$arrows),
  R2 = indirect_fit$vectors$r,
  P_value = indirect_fit$vectors$pvals
)
indirect_r2_df
```
For the turb exploration, I kept everything. And, you know what, it all was significant, except the del13c. Which is kind of crazy. 

```{r}
plot(ord)
plot(indirect_fit)
```

### Water Quality Envfit
```{r}
wc_explan_envfit <- left_join(wide_maxn_df, 
                           water_quality_explan, 
                           by = "Pin")

direct_fit <- envfit(ord ~ Low_Tide_Mean_NN_umolL + Low_Tide_Mean_Silicate_umolL + Low_Tide_Mean_Phosphate_umolL + Low_Tide_Mean_pH + Low_Tide_Mean_TA +Low_Tide_Mean_Salinity + shore_dist_m + dist_to_seep_m, data = wc_explan_envfit, na.rm = T, permutations = 999)
direct_fit
```


#### Plot it
```{r}
ord_scores <- as.data.frame(scores(ord, display = "sites"))
ord_scores2 <- as.data.frame(scores(ord, "site"))
envfit_vectors <- as.data.frame(direct_fit$vectors$arrows)  # Extract arrow coordinates
envfit_pvals <- direct_fit$vectors$pvals  # Extract p-values
envfit_vectors$Variable <- rownames(envfit_vectors)
envfit_vectors$pval <- envfit_pvals  # Add p-values to the dataframe

# Filter only significant vectors (p < 0.05)
sig_vectors <- envfit_vectors %>% filter(pval < 0.05)

# Scale arrows for better visualization (optional)
arrow_scale <- 1  # Adjust scaling factor
sig_vectors <- sig_vectors %>%
  mutate(NMDS1 = NMDS1 * arrow_scale, NMDS2 = NMDS2 * arrow_scale)

# Base NMDS plot
nmds <- ggplot(data = ord_scores, aes(x = NMDS1, y = NMDS2)) +
  geom_point() +
  theme_minimal()+
  xlim(-1.2, 1.2) +
  ylim(-1.2, 1.2)
  
nmds
# Create NMDS plot with envfit arrows
envfit <- ggplot(data = ord_scores, aes(x = NMDS1, y = NMDS2)) +
  geom_point() +
  geom_segment(data = sig_vectors, 
               aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "darkgreen") +
  geom_text(data = sig_vectors, 
            aes(x = 0.5*NMDS1, y = 0.5*NMDS2, label = Variable), 
            hjust = 1.2, vjust = 1.2, color = "darkgreen") +
  theme_minimal()+
  xlim(-1.2, 1.2) +
  ylim(-1.2, 1.2)

# Print the plot
(envfit)
```

## Residual 
```{r}
resid_explan_envfit <- left_join(wide_maxn_df, 
                          # explan[, c("Pin", "del15N", "N_percent", "meanRugosity", "shore_dist_m", "dist_to_seep_m")], 
                          resid_explan,
                           by = "Pin")

# resid_fit <- envfit(ord ~ del15N + N_percent + shore_dist_m + dist_to_seep_m + meanRugosity, data = explan_envfit, na.rm = T, permutations = 999)
resid_fit <- envfit(ord ~ res_dist_to_seep_m + res_shore_dist_m + res_del15N + res_N_percent + res_C_percent + res_C_N + res_del13C + res_Low_Tide_Mean_NN_umolL + res_Low_Tide_Mean_Silicate_umolL + res_Low_Tide_Mean_Phosphate_umolL + res_Low_Tide_Mean_pH + res_Low_Tide_Mean_TA + res_Low_Tide_Mean_Salinity + res_pulse_pc1, data = resid_explan_envfit, na.rm = T, permutations = 999)
resid_fit

#to get r2:
resid_fit$vectors$r

resid_r2_df <- data.frame(
  Variable = rownames(resid_fit$vectors$arrows),
  R2 = resid_fit$vectors$r,
  P_value = resid_fit$vectors$pvals
)
resid_r2_df

resid_r2_df_sig <- resid_r2_df %>% filter(P_value < 0.05)
```

The nice thing about envfit is that it's running permutations on each variable individually, not looking at it as a multivariate model. But, these variables are correlated. See:
```{r}
resid_cor_matrix <- cor(na.omit(resid_explan[,c("res_dist_to_seep_m", "res_shore_dist_m", "res_del15N", "res_C_percent", "res_Low_Tide_Mean_Silicate_umolL","res_Low_Tide_Mean_Salinity")]))
resid_corplot <- ggcorrplot(resid_cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)
resid_corplot
```


# Significant species
```{r}
#Start with making a dataframe that has what we want from the ordination
data.scores = as.data.frame(scores(ord, "site"))
#including the factors (Pin)
data.scores$Site = wide_maxn_df$Pin 
#data.scores$Site <- factor(data.scores$Site, levels = c(1:223))
# head(data.scores)

#envfit to see which species are making a difference here:
data.spp.fit <- envfit(data.scores, speciesdata, permutations = 999)

#check that went ok
#head(data.spp.fit)

spp.scrs <- as.data.frame(scores(data.spp.fit, display = "vectors")) #save species intrinsic values into dataframe
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs)) #add species names to dataframe
#head(spp.scrs)

# Extract p-values from the envfit result
pvals <- data.spp.fit$vectors$pvals
# Add the p-values to the spp.scrs dataframe
spp.scrs$pval <- pvals

# Subset data to show species significant at 0.05
sig.spp.scrs <- subset(spp.scrs, pval <= 0.05)
#make more strict because showing too many species
#sig.spp.scrs <- subset(spp.scrs, pval <= 0.01)
# sig.spp.scrs <- spp.scrs
# sig.spp.scrs <- subset(spp.scrs, pvals<=0.05) #subset data to show species significant at 0.05

head(sig.spp.scrs)
```

Plot with significant species
```{r}
envfit_sp <- ggplot(data = data.scores, aes(x = NMDS1, y = NMDS2)) +
  #geom_point() +
  ggrepel::geom_text_repel(
    data = sig.spp.scrs,
    aes(x = NMDS1, y = NMDS2, label = Species),
    cex = 2,
    direction = "both",
    segment.size = 0,
    max.overlaps = getOption("ggrepel.max.overlaps", default = 100)
  )+
  theme_minimal()+
  xlim(-1.2, 1.2) +
  ylim(-1.2, 1.2)
```

#### Put the envfit and species plots together:
```{r}
cowplot::plot_grid(nmds, envfit, envfit_sp, labels = c('A', 'B', 'C'), ncol = 3)
```

# Herbivore-only ordination
```{r}
Fish_trans_herb <- decostand(herbspeciesdata, method = "max")
non_zero_rows <- rowSums(Fish_trans_herb == 0) != ncol(Fish_trans_herb)
species.data.use <- Fish_trans_herb[non_zero_rows, ]

ordherb = metaMDS(species.data.use, dist = "bray", trymax = 500, k = 2, trace = FALSE)
ordherb 
stressplot(ord) ##stressplot is just a diagnostic for your data, shows how your datapoints are distributed and should be somewhat linear
##Stress Value = how well doe your NMDS represent reality (<0.2)
plot(ord, type = "text")
```

### Herbivore Envfit
For an envfit, we need the explanatory dataset in the size of the fish data:
```{r}
explan_envfit_herb <- left_join(wide_herb_df, 
                           indirect_explan[, c("Pin", "del15N", "N_percent", "meanRugosity", "shore_dist_m", "dist_to_seep_m")], 
                           by = "Pin")
```

```{r}
fitherb <- envfit(ordherb ~ del15N + N_percent + shore_dist_m + meanRugosity, data = explan_envfit_herb, na.rm = T, permutations = 999) 
fitherb
```

### Plot it
```{r}
ordherb_scores <- as.data.frame(scores(ordherb, display = "sites"))
ordherb_scores2 <- as.data.frame(scores(ordherb, "site"))
envfit_vectors <- as.data.frame(fitherb$vectors$arrows)  # Extract arrow coordinates
envfit_pvals <- fitherb$vectors$pvals  # Extract p-values
envfit_vectors$Variable <- rownames(envfit_vectors)
envfit_vectors$pval <- envfit_pvals  # Add p-values to the dataframe

# Filter only significant vectors (p < 0.05)
sig_vectors <- envfit_vectors %>% filter(pval < 0.05)

# Scale arrows for better visualization (optional)
arrow_scale <- 1.5  # Adjust scaling factor
sig_vectors <- sig_vectors %>%
  mutate(NMDS1 = NMDS1 * arrow_scale, NMDS2 = NMDS2 * arrow_scale)

# Base NMDS plot
nmds_herb <- ggplot(data = ordherb_scores, aes(x = NMDS1, y = NMDS2)) +
  geom_point() +
  theme_minimal()+
  xlim(-2.3, 2.3) +
  ylim(-1.5, 1.5)
  

# Create NMDS plot with envfit arrows
envfit_herb <- ggplot(data = ordherb_scores, aes(x = NMDS1, y = NMDS2)) +
  geom_point() +
  geom_segment(data = sig_vectors, 
               aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "darkgreen") +
  geom_text(data = sig_vectors, 
            aes(x = NMDS1, y = NMDS2, label = Variable), 
            hjust = 1.2, vjust = 1.2, color = "darkgreen") +
  theme_minimal()+
  xlim(-2.3, 2.3) +
  ylim(-1.5, 1.5)

# Print the plot
print(envfit_herb)
```

## Significant species
```{r}

#Start with making a dataframe that has what we want from the ordination
herbdata.scores = as.data.frame(scores(ordherb, "site"))
#including the factors (Pin)
herbdata.scores$Site = wide_herb_df$Pin 
#data.scores$Site <- factor(data.scores$Site, levels = c(1:223))
# head(data.scores)

#envfit to see which species are making a difference here:
data.spp.fit <- envfit(herbdata.scores, herbspeciesdata, permutations = 999)

#check that went ok
#head(data.spp.fit)

spp.scrs <- as.data.frame(scores(data.spp.fit, display = "vectors")) #save species intrinsic values into dataframe
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs)) #add species names to dataframe
#head(spp.scrs)

# Extract p-values from the envfit result
pvals <- data.spp.fit$vectors$pvals
# Add the p-values to the spp.scrs dataframe
spp.scrs$pval <- pvals

# Subset data to show species significant at 0.05
sig.spp.scrs <- subset(spp.scrs, pval <= 0.05)
#make more strict because showing too many species
#sig.spp.scrs <- subset(spp.scrs, pval <= 0.01)
# sig.spp.scrs <- spp.scrs
# sig.spp.scrs <- subset(spp.scrs, pvals<=0.05) #subset data to show species significant at 0.05

head(sig.spp.scrs)
```

Plot with significant species
```{r}
envfit_herb_sp <- ggplot(data = herbdata.scores, aes(x = NMDS1, y = NMDS2)) +
  #geom_point() +
  ggrepel::geom_text_repel(
    data = sig.spp.scrs,
    aes(x = NMDS1, y = NMDS2, label = Species),
    cex = 3,
    direction = "both",
    segment.size = 0
  )+
  theme_minimal()+
  xlim(-2.3, 2.3) +
  ylim(-1.5, 1.5)
```


#### Put the plots together:
```{r}
cowplot::plot_grid(nmds_herb, envfit_herb, envfit_herb_sp, labels = c('A', 'B', 'C'),ncol = 3)
```


GAMLAND
