---
title: "Buoyant_Weight_SGD"
author: "Callie Stephenson"
date: "2023-05-23"
output: html_document
---

### Description
This R Markdown file is meant to create a graph that evaluates the change in buoyant weight of corals from Callie's 2023 Moorea experiment against the SGD Gradient.

Data for this project is accessed from GitHub [here](https://github.com/donahuem/MooreaCoralFish_CS/tree/main/Coral_Data)
```{r}
#if needed, install tidyverse by removing the # in the line below:
#install.packages("tidyverse")

#add libraries
library(gdata)
library(reshape)
library(tidyverse)
library(ggplot2)

# adding Coral data
Coral_Dict <- read.csv("https://raw.githubusercontent.com/donahuem/MooreaCoralFish_CS/main/Coral_Data/PAC_Coral_Codes.csv")

Buoyant_weight_Pac <- read.csv("https://raw.githubusercontent.com/donahuem/MooreaCoralFish_CS/main/Coral_Data/PAC_Buoyant_Weight.csv")
No_NA_Buoyant_weight_Pac <- Buoyant_weight_Pac%>%
               drop_na(T1_Weight)

Merged_Coral_Data <- merge(Coral_Dict, No_NA_Buoyant_weight_Pac, by = "Vial_No_Only", all.x=FALSE, all.y = TRUE)

# adding Gradient data
BioGC.Info_raw <-read_csv("https://raw.githubusercontent.com/njsilbiger/MooreaSGD_site-selection/main/Data/August2021/Allbiogeochemdata_QC.csv")
BioGC.Info <- BioGC.Info_raw %>%
  filter(!Date == "2021-08-06") #filter out Varari high wave event
BioGC.Info <- BioGC.Info %>%
  filter(!Location == "Cabral") #filter out Cabral data
BioGC.Info <- BioGC.Info %>%
  filter(!Silicate_umolL == "NA") #filter out NA
BioGC.Info <- BioGC.Info %>%
  filter(!CowTagID == "Varari_Well") #filter out Varari_Well
BioGC.Info_no_Seep <- BioGC.Info %>%
  filter(!CowTagID == "VSEEP") #filter out Seep
BioGC.Info_no_Seep$Varari_Pin <-as.numeric(gsub("V","",BioGC.Info_no_Seep$CowTagID)) #adds column 'Varari_Pin' that removes the V in CowTagID so that pin placements are numbers not characters

```
Now that the data is loaded, this code displays the values for silicate along the gradient by tide.
```{r}
Silicate_Gradient <- ggplot(BioGC.Info, aes(
  x=factor(CowTagID, level=c('VSEEP','V1', 'V2', 'V3','V4','V5','V6','V7','V8','V9','V10','V11','V12','V13','V14','V15','V16','V17','V18','V19','V20')), 
  y =Silicate_umolL)) + geom_point (color="cyan4") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+xlab('Location')+ylab('Silicate (umol/L)')+stat_summary(
        geom = "point",
        fun = "mean",
        col = "darkolivegreen",
        size = 2,
        shape = 24,
        fill = "darkolivegreen") +labs(title='Silicate Gradient with Seep')
Silicate_Gradient
```
Since the seep is so much higher than the other values, and the seep was not used in this experiment, I graphed this data without the seep included:
```{r}
Silicate_Gradient_no_Seep <- ggplot(BioGC.Info_no_Seep, aes(
  x=Varari_Pin, 
  y =Silicate_umolL)) + 
  geom_point (aes(color=Tide)) +
  labs(title='Silicate Gradient without Seep') +
  xlab('Location')+
  ylab('Silicate (umol/L)')+ 
  scale_x_continuous(n.breaks=20)+
  scale_color_manual(values = c("steelblue", "lightskyblue"))+
  stat_summary(
        geom = "point",
        fun = "mean",
        fill = "darkolivegreen",
        size = 2,
        show.legend = NA,
        aes(shape="mean")) +scale_shape_manual("", values=c("mean"="x")) 
  
Silicate_Gradient_no_Seep 
```
Now that we have an understanding of what the gradient looks like, it is time to compare that to our change in buoyant weight

This code creates a Data Frame with the mean values displayed above
```{r}
Mean_Silicate <- setNames(aggregate(BioGC.Info_no_Seep$Silicate_umolL, list(BioGC.Info_no_Seep$CowTagID), FUN=mean),c("CowTagID","Mean"))
Mean_Silicate$Varari_Pin<- as.numeric(gsub("V","",Mean_Silicate$CowTagID))
Mean_Silicate =Mean_Silicate[order(Mean_Silicate$Mean),]
View(Mean_Silicate)
```
Now I graph just the mean values and change the orientation on the x-axis to lowest to highest value
```{r}
Silicate_Mean<- ggplot()+
  geom_line(data=Mean_Silicate, aes(
    x=reorder(Varari_Pin,Mean_Silicate$Mean), 
    y=Mean,
    group=1))+
  geom_point(data=Mean_Silicate, aes(
    x=reorder(Varari_Pin,Mean_Silicate$Mean), 
    y=Mean))+
  xlab('Location')
```
```{r}
Silicate_Mean
```
This code does the same thing, but for the Maximum Silicate Values
```{r}
Max_Silicate <- setNames(aggregate(BioGC.Info_no_Seep$Silicate_umolL, list(BioGC.Info_no_Seep$CowTagID), FUN=max),c("CowTagID","Maximum_Silicate"))
Max_Silicate$Varari_Pin <- as.numeric(gsub("V","",Max_Silicate$CowTagID))
Max_Silicate =Max_Silicate[order(Max_Silicate$Max),]
View(Max_Silicate)
```
```{r}
Silicate_Max<- ggplot()+
  geom_line(data=Max_Silicate, aes(
    x=reorder(Varari_Pin,Max_Silicate$Max), 
    y=Max_Silicate$Maximum_Silicate,
    group=1))+
  geom_point(data=Max_Silicate, aes(
    x=reorder(Varari_Pin,Max_Silicate$Maximum_Silicate), 
    y=Max_Silicate$Maximum_Silicate))+
  xlab('Location')
Silicate_Max
```

Plotting the Seep values to include the trend line for Mean Values above:
```{r}
Silicate_Mean_with_all_values <- Silicate_Mean +geom_point(data=BioGC.Info_no_Seep, aes(
    x=Varari_Pin, 
    y=Silicate_umolL, 
    color=Tide)) +
  scale_color_manual(values = c("steelblue", "lightskyblue","black"))+
  labs(title='Silicate Gradient without Seep')+
  xlab('Location')+
  ylab('Silicate (umol/L)')
```
```{r}
Silicate_Mean_with_all_values

```
Now, to compare the mean gradient to the change in buoyant weight
```{r}
Buoyant_Weight_on_Mean_Silicate_Gradient <- ggplot()+
  geom_line(data=Mean_Silicate, aes(
    x=reorder(Varari_Pin,Mean_Silicate$Mean), 
    y=Mean,
    group=1,
    color="Mean"))+
  geom_point(data=Mean_Silicate, aes(
    x=reorder(Varari_Pin,Mean_Silicate$Mean), 
    y=Mean,
    color="Mean"))+
  geom_line(data=Merged_Coral_Data, aes(
    x=Pin_Number,
    y=X._Change_100/6,
    color =Cage_Uncaged))+
  geom_point(data=Merged_Coral_Data, aes(
    x=Pin_Number,
    y=X._Change_100/6,
    color =Cage_Uncaged))+
  scale_y_continuous(name="Mean Silicate Concentration (umol/L)",
                       sec.axis = sec_axis(~.*6, name = "Change in Buoyant Weight"))+
  xlab('Varari Pin')+
  ylab('Silicate Concentration')+
  labs(title='Percent change in Buoyant Weight Relative to Mean Silicate Concentration')+
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07","black"),name="", labels = c("Uncaged", "Partially Caged", "Fully Caged","Mean Silicate Concentration"))

Buoyant_Weight_on_Mean_Silicate_Gradient
```
And, to compare the maximum values to the change in buoyant weight
```{r}
scale = 2.5
Buoyant_Weight_on_Max_Silicate_Gradient <- ggplot()+
  geom_line(data=Max_Silicate, aes(
    x=reorder(Varari_Pin,Max_Silicate$Max), 
    y=Max_Silicate$Maximum_Silicate,
    group=1))+
  geom_point(data=Max_Silicate, aes(
    x=reorder(Varari_Pin,Max_Silicate$Maximum_Silicate), 
    y=Max_Silicate$Maximum_Silicate))+
  xlab('Location') +
  geom_line(data=Merged_Coral_Data, aes(
    x=Pin_Number,
    y=X._Change_100/scale,
    color =Cage_Uncaged))+
  geom_point(data=Merged_Coral_Data, aes(
    x=Pin_Number,
    y=X._Change_100/scale,
    color =Cage_Uncaged))+
  scale_y_continuous(name="Mean Silicate Concentration (umol/L)",
                       sec.axis = sec_axis(~.*scale, name = "Change in Buoyant Weight"))+
  xlab('Varari Pin')+
  ylab('Silicate Concentration')+
  labs(title='Percent change in Buoyant Weight Relative to Maximum Silicate Concentration')+
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07","black"),name="", labels = c("Uncaged", "Partially Caged", "Fully Caged","Maximum Silicate Concentration"))

Buoyant_Weight_on_Max_Silicate_Gradient
```




Code for just a buoyant weight graph. I cannot figure out how to order the graph for the silicate concentration without plotting both on the same graph, which limits the axis
```{r}
PAC_Buoyant_Weight <- ggplot()+
  geom_point(data=Merged_Coral_Data, aes(
    x=Pin_Number,
    y=X._Change_100,
    color =Cage_Uncaged))+
  scale_y_continuous(name="Silicate Concentration",
                       sec.axis = sec_axis(trans=~.*10, name = "Change in Buoyant Weight"))+
  xlab('Varari Pin')+
  ylab('Silicate Concentration')+
  labs(title='Percent change in Buoyant Weight Relative to Silicate Concentration')+
  scale_color_discrete(name = "", labels = c("Uncaged", "Partially Caged", "Fully Caged","Mean Silicate Concentration"))
PAC_Buoyant_Weight
```

