---
title: "Calcification_Analysis"
author: "Callie Stephenson"
date: "2024-12-04"
output: html_document
---
# Setup
```{r}
library(dplyr)
library(ggplot2)
library(glmnet)
library(effects)
library(ggeffects)
library(performance)
library(ggpubr)
library(lmerTest)
```


```{r}
response_data <- read.csv("data/T1_response_data.csv")
explanatory_variables <- read.csv("/Users/calliestephenson/Documents/GitHub/MooreaCoralFish_CS/data/explanatory_all_wide.csv")
```

```{r}
model_data <- left_join(response_data, explanatory_variables, by = join_by(CowTagID))

PAC_model_data <- model_data %>%
  filter(Species == "Pocillopora acuta", !is.na(Pin_Number))

PRU_model_data <- model_data %>%
  filter(Species == "Porites rus", !is.na(Pin_Number))

PRU_model_data_caged_only <- PRU_model_data %>% 
  filter(Cage_Uncaged == "C")
```

## Collinearity
```{r}
chart.Correlation(na.omit(model_data[,c("sd_Salinity","sd_Temperature", "sd_TA", "sd_pH")]))
```

```{r}
chart.Correlation(na.omit(model_data[,c("sd_Salinity","sd_Temperature", "sd_pH")]))
```

Temperature variability, salinity variability, and pH variability are all significantly corrleated, but TA is not


Still, a PCA is not appropriate:
```{r}
sd_columns <- names(explanatory_variables)[grepl("sd_", names(explanatory_variables)) & 
                          !grepl("Phosphate_umolL|NN_umolL|Ammonia_umolL|NNP_umolL|Silicate_umolL",names(explanatory_variables))]
sd.pca.data <- na.omit(explanatory_variables[, c(sd_columns)])
sd.pca = princomp(sd.pca.data, cor=TRUE)
summary(sd.pca)
```

Therefore, we will use the LASSO method, which can do variable selection with shrinkage

# Lasso
## Growth

### P acuta
```{r}
linear_growth_model_pac <- lmer(Percent_Change ~ sd_TA +
                                  (1|Genotype), data = PAC_model_data)
#check_model(linear_growth_model_pru_c)
Anova(linear_growth_model_pac)
AIC(linear_growth_model_pac)

linear_growth_model_pac <- lmer(Percent_Change ~ sd_pH +
                                  (1|Genotype), data = PAC_model_data)
#check_model(linear_growth_model_pru_c)
Anova(linear_growth_model_pac)
AIC(linear_growth_model_pac)

linear_growth_model_pac <- lmer(Percent_Change ~ sd_pH + sd_TA +
                                  (1|Genotype), data = PAC_model_data)
#check_model(linear_growth_model_pru_c)
Anova(linear_growth_model_pac)
AIC(linear_growth_model_pac)
```


```{r PAC lasso}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("Percent_Change","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  mutate(sd_TA = scale(sd_TA))%>% 
  mutate(sd_pH = scale(sd_pH))%>% 
  mutate(sd_Salinity = scale(sd_Salinity))%>% 
  mutate(sd_Temperature = scale(sd_Temperature))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)
PAC_lasso$Cage_Uncaged <- as.factor(PAC_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_percent_change <- PAC_model_data_matrix[, -1]
y_pac_percent_change <- PAC_model_data_matrix[, 1]
lasso_PAC_growth <- cv.glmnet(x = x_pac_percent_change, y = y_pac_percent_change)
coef(lasso_PAC_growth)
```

#### Significant model
```{r}
multivariate_PAC_growth <- lmer(Percent_Change ~ (sd_Temperature) + (sd_TA) + (sd_pH) + 
  (1 | Genotype) + (1 | Cage_Uncaged),
  data = PAC_model_data
)
summary(multivariate_PAC_growth)
car::Anova(multivariate_PAC_growth)
plot(allEffects(multivariate_PAC_growth))
anova(multivariate_PAC_growth)
```

```{r}
multivariate_PAC_growth <- lmer(Percent_Change ~ scale(sd_Temperature) + scale(sd_TA) + scale(sd_pH) + 
  (1 | Genotype) + (1 | Cage_Uncaged),
  data = PAC_model_data
)
```

```{r}
# multivariate_PAC_growth <- lmer(Percent_Change ~ scale(sd_Temperature) + scale(sd_TA) + scale(sd_pH) + 
#   (1 | Genotype) + (1 | Cage_Uncaged),
#   data = PAC_model_data
# )
# summary(multivariate_PAC_growth)
# car::Anova(multivariate_PAC_growth)
# plot(allEffects(multivariate_PAC_growth))

#check_model(multivariate_PAC_growth)
```

```{r}
null_model <- lmer(Percent_Change ~ (1 | Genotype) + (1 | Cage_Uncaged), data = PAC_model_data)
anova(null_model, multivariate_PAC_growth)
```


##### Sig model plot
```{r}
pred_temp <- ggpredict(multivariate_PAC_growth, terms = "sd_Temperature")
pred_ta <- ggpredict(multivariate_PAC_growth, terms = "sd_TA")
pred_ph <- ggpredict(multivariate_PAC_growth, terms = "sd_pH")

#annnotations
anova_results <- car::Anova(multivariate_PAC_growth)
p_values <- anova_results$`Pr(>Chisq)`[1:3]  # Adjust index based on your model's fixed effects
names(p_values) <- rownames(anova_results)[1:3] 
```

```{r}
custom_theme <- theme_minimal(base_size = 10)+
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 10)
  )
```

```{r}
# Plot for sd_Temperature
plot_temp <- ggplot() +
  # Raw data points
  geom_point(data = PAC_model_data, 
             aes(x = sd_Temperature, y = Percent_Change), color = "grey50", alpha = 0.6) +
  # Prediction line with confidence interval
  geom_line(data = pred_temp, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_temp, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "Standard Deviation of Temperature",
    y = "Percent Change in Buoyant Weight"
  ) +
  custom_theme+
  annotate("text", y = 1, x = max(pred_temp$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["sd_Temperature"], 6)), 
           hjust = 1.1, vjust = 1, size = 3)

# Plot for sd_TA
plot_ta <- ggplot() +
  geom_point(data = PAC_model_data, 
             aes(x = sd_TA, y = Percent_Change), color = "grey50", alpha = 0.6) +
  geom_line(data = pred_ta, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_ta, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "Standard Deviation of TA", y = " "
  ) +
  custom_theme+
  annotate("text", y = 1, x =max(pred_ta$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["sd_TA"], 6)), 
           hjust = 1.1, vjust = 1, size = 3)

# Plot for sd_pH
plot_ph <- ggplot() +
  geom_point(data = PAC_model_data, 
             aes(x = sd_pH, y = Percent_Change), color = "grey50", alpha = 0.6) +
  geom_line(data = pred_ph, aes(x = x, y = predicted), color = "#21908c") +
  geom_ribbon(data = pred_ph, aes(x = x, ymin = conf.low, ymax = conf.high), 
              fill = "#21908c", alpha = 0.2) +
  scale_y_continuous(limits = c(-1.5,25)) +
  labs(
    x = "Standard Deviation of pH", y = " "
  ) +
  custom_theme+
  annotate("text", y = 1, x =max(pred_ph$x, na.rm=TRUE), 
           label = paste("p =", round(p_values["sd_pH"], 9)), 
           hjust = 1.1, vjust = 1, size = 3)
```

```{r}
combined_plot <- ggarrange(plot_temp + 
                              labs(subtitle = "A") + 
                              theme(plot.subtitle = element_text(hjust = -0.2, size = 12)),
                            plot_ta + 
                              labs(subtitle = "B") + 
                              theme(plot.subtitle = element_text(hjust = -0.2, size = 12)),
                            plot_ph + 
                              labs(subtitle = "C") + 
                              theme(plot.subtitle = element_text(hjust = -0.2, size = 12)),
                            nrow = 1)+ 
  ggtitle(expression("Effects of Variability on Growth in" ~ italic("Pocillopora acuta"))) +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        plot.margin = margin(t = 20, r = 20, b = 20, l = 40))
combined_plot
#ggsave(combined_plot, filename = "PAC_Percent_Change_Effects.png", dpi = 600, width = 10, height = 5)
```

### Porites rus
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("Percent_Change","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)
PRU_lasso$Cage_Uncaged <- as.factor(PRU_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_percent_change <- PRU_model_data_matrix[, -1]
y_PRU_percent_change <- PRU_model_data_matrix[, 1]
lasso_PRU_growth <- cv.glmnet(x = x_PRU_percent_change, y = y_PRU_percent_change)
coef(lasso_PRU_growth)
```

### Caged Only Porites rus
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso_c <- PRU_model_data_caged_only[,c("Percent_Change","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso_c$Genotype <- as.factor(PRU_lasso_c$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso_c)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso_c <- cbind(PRU_lasso_c, genotype_dummies)
PRU_lasso_c <- PRU_lasso_c %>% dplyr::select(-Genotype)

#lasso works on matrix
PRU_model_data_matrix_c <- as.matrix(PRU_lasso_c)

x_PRU_percent_change_c <- PRU_model_data_matrix_c[, -1]
y_PRU_percent_change_c <- PRU_model_data_matrix_c[, 1]
lasso_PRU_growth_c <- cv.glmnet(x = x_PRU_percent_change_c, y = y_PRU_percent_change_c)
coef(lasso_PRU_growth_c)
```

## Symb Abundance

### P acuta
```{r}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("FSC.Events_per_cm_2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

PAC_lasso <- PAC_lasso %>% 
  mutate(FSC.Events_per_cm_2 = log10(FSC.Events_per_cm_2)) 

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)
PAC_lasso$Cage_Uncaged <- as.factor(PAC_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_symb <- PAC_model_data_matrix[, -1]
y_pac_symb <- PAC_model_data_matrix[, 1]
lasso_PAC_symb <- cv.glmnet(x = x_pac_symb, y = y_pac_symb)
coef(lasso_PAC_symb)
```

### Porites rus
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("FSC.Events_per_cm_2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

PRU_lasso <- PRU_lasso %>% 
  mutate(FSC.Events_per_cm_2 = log10(FSC.Events_per_cm_2)) %>% 
  filter(!is.na(FSC.Events_per_cm_2))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)
PRU_lasso$Cage_Uncaged <- as.factor(PRU_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_symb <- PRU_model_data_matrix[, -1]
y_PRU_symb <- PRU_model_data_matrix[, 1]
lasso_PRU_symb <- cv.glmnet(x = x_PRU_symb, y = y_PRU_symb)
coef(lasso_PRU_symb)
```

### Caged Only Porites rus
```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data_caged_only[,c("FSC.Events_per_cm_2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")]

PRU_lasso <- PRU_lasso %>% 
  mutate(FSC.Events_per_cm_2 = log10(FSC.Events_per_cm_2)) %>% 
  filter(!is.na(FSC.Events_per_cm_2))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_symb <- PRU_model_data_matrix[, -1]
y_PRU_symb <- PRU_model_data_matrix[, 1]
lasso_PRU_symb <- cv.glmnet(x = x_PRU_symb, y = y_PRU_symb)
coef(lasso_PRU_symb)
```

whole lotta nothing

## Chlorophyll

### P acuta
```{r PAC lasso}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("Chl_ug.cm.2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  mutate(sd_TA = scale(sd_TA))%>% 
  mutate(sd_pH = scale(sd_pH))%>% 
  mutate(sd_Salinity = scale(sd_Salinity))%>% 
  mutate(sd_Temperature = scale(sd_Temperature))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)
PAC_lasso$Cage_Uncaged <- as.factor(PAC_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_Chl_ug.cm.2 <- PAC_model_data_matrix[, -1]
y_pac_Chl_ug.cm.2 <- PAC_model_data_matrix[, 1]
lasso_PAC_chla <- cv.glmnet(x = x_pac_Chl_ug.cm.2, y = y_pac_Chl_ug.cm.2)
coef(lasso_PAC_chla)
```

### P rus
```{r PAC lasso}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("Chl_ug.cm.2","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  mutate(sd_TA = scale(sd_TA))%>% 
  mutate(sd_pH = scale(sd_pH))%>% 
  mutate(sd_Salinity = scale(sd_Salinity))%>% 
  mutate(sd_Temperature = scale(sd_Temperature))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)
PRU_lasso$Cage_Uncaged <- as.factor(PRU_lasso$Cage_Uncaged)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

#remove NA values
complete_cases <- complete.cases(y_PRU_Chl_ug.cm.2)
x_PRU_Chl_ug.cm.2 <- x_PRU_Chl_ug.cm.2[complete_cases, ]
y_PRU_Chl_ug.cm.2 <- y_PRU_Chl_ug.cm.2[complete_cases]

# x_PRU_Chl_ug.cm.2 <- PRU_model_data_matrix[, -1]
# y_PRU_Chl_ug.cm.2 <- PRU_model_data_matrix[, 1]
lasso_PRU_chla <- cv.glmnet(x = x_PRU_Chl_ug.cm.2, y = y_PRU_Chl_ug.cm.2)
coef(lasso_PRU_chla)
```

## Heterotrophy

### Carbon

#### PAC
```{r}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("Δ13C","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ13C))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_13c <- PAC_model_data_matrix[, -1]
y_pac_13c <- PAC_model_data_matrix[, 1]
lasso_PAC_13c <- cv.glmnet(x = x_pac_13c, y = y_pac_13c)
coef(lasso_PAC_13c)
```

#### PRU

```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("Δ13C","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ13C))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies, cage_uncaged_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_13c <- PRU_model_data_matrix[, -1]
y_PRU_13c <- PRU_model_data_matrix[, 1]
lasso_PRU_13c <- cv.glmnet(x = x_PRU_13c, y = y_PRU_13c)
coef(lasso_PRU_13c)
```

nothin again

### Nitrogen

#### PAC
```{r}
# Remove the Pin_Number and Species columns
PAC_lasso <- PAC_model_data[,c("Δ15N","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ15N))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PAC_lasso$Genotype <- as.factor(PAC_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PAC_lasso)
cage_uncaged_dummies <- model.matrix( ~ Cage_Uncaged - 1, data = PAC_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PAC_lasso <- cbind(PAC_lasso, genotype_dummies, cage_uncaged_dummies)
PAC_lasso <- PAC_lasso[, !names(PAC_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PAC_model_data_matrix <- as.matrix(PAC_lasso)

x_pac_15N <- PAC_model_data_matrix[, -1]
y_pac_15N <- PAC_model_data_matrix[, 1]
lasso_PAC_15N <- cv.glmnet(x = x_pac_15N, y = y_pac_15N)
coef(lasso_PAC_15N)
```

#### PRU

```{r}
# Remove the Pin_Number and Species columns
PRU_lasso <- PRU_model_data[,c("Δ15N","Cage_Uncaged","Genotype","sd_TA", "sd_pH", "sd_Salinity", "sd_Temperature")] %>% 
  filter(!is.na(Δ15N))

# Convert Genotype and Cage_Uncaged to factors if they are not already
PRU_lasso$Genotype <- as.factor(PRU_lasso$Genotype)

# Create dummy variables
genotype_dummies <- model.matrix( ~ Genotype - 1, data = PRU_lasso)

# Combine the dummy variables with the original data frame (excluding the original columns)
PRU_lasso <- cbind(PRU_lasso, genotype_dummies)
PRU_lasso <- PRU_lasso[, !names(PRU_lasso) %in% c("Genotype", "Cage_Uncaged")]

#lasso works on matrix
PRU_model_data_matrix <- as.matrix(PRU_lasso)

x_PRU_15N <- PRU_model_data_matrix[, -1]
y_PRU_15N <- PRU_model_data_matrix[, 1]
lasso_PRU_15N <- cv.glmnet(x = x_PRU_15N, y = y_PRU_15N)
coef(lasso_PRU_15N)
```