---
title: "Moorea Coral Data Exploration"
author: "Callie Stephenson"
date: "2024-05-29"
output: html_document
---
## Introduction

This is a R Markdown file in which I hope to write out my analyses. I will use this document to process all the R scripts found in the R folder and complete my project. 

#### Loading the data
```{r load packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(ggrepel)
library(patchwork)
library(tidytext)
library(AICcmodavg)
library(kableExtra)
library(PNWColors)
library(fishualize)
library(lme4)
library(lmerTest) #figure out which lme package you need
library(ggpp)
library(ggplot2)
library(GGally)
library(ggeffects)
library(arm) #discrete hist
library(MuMIn) #dredge
library(effects) #effects plots for dredged models
library(curl) #curl
library(gridExtra)
library(vegan) #pcoa
```


```{r}
setwd("/Users/calliestephenson/Documents/GitHub/MooreaCoralFish_CS/")
#explanatory data
explanatory_all_wide <- read.csv("data/explanatory_all_wide.csv")
explanatory_seasonal <- read.csv("data/Explanatory_variables.csv")

#Response data
BW <- read.csv("data/BW_tidy.csv")
TLE <- read.csv("data/TLE_summary.csv")
FCM <- read.csv("data/FCM_tidy.csv") %>% 
  rename(Placement_Code = PLACEMENT)
response_data <- read.csv("data/response_data.csv")


PRU_response_data <- response_data %>% 
  filter(Species == "Porites rus")

PRU_response_data_caged <- PRU_response_data %>% 
  filter(Cage_Uncaged == "C")

PAC_response_data <- response_data %>% 
  filter(Species == "Pocillopora acuta")

PAC_response_data_caged <- PAC_response_data %>% 
  filter(Cage_Uncaged == "C")

#metadata
meta <- read.csv("data/coral_metadata.csv")
```


```{r all data seasonal, warning=FALSE}
#now make an all data that has both explanatory and response variables:
all_data_seasonal <- left_join(response_data, explanatory_seasonal, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
all_data_seasonal$Genotype <- as.factor(all_data_seasonal$Genotype)
all_PAC_data_seasonal <- all_data_seasonal %>% 
  filter(Species == "Pocillopora acuta")
```

```{r all data both seasons, warning=FALSE}
#now make an all data that has both explanatory and response variables:
#all_data <- left_join(response_data, explanatory_all, by = join_by(CowTagID))
#suppress warnings because there will be a many-to-many relationship between x and y
#all_data$Genotype <- as.factor(all_data$Genotype)
#all_PAC_data <- all_data %>% 
#  filter(Species == "Pocillopora acuta")

all_data_wide <- left_join(response_data, explanatory_all_wide, by = join_by(CowTagID))
```


## Data Exploration:
Data exploration focuses on the following points:
1. Outliers
2. Collinearity
3. Independence (Relationships between response and explanatory variables) 

```{r response data hist, fig.width=10, fig.height=10, echo=FALSE}
#png("output/response_data_hist.png", width = 1100, height = 1100)
par(mfrow = c(4,4), oma=c(1,1,3,1))
hist(PAC_response_data$Change_Over_Area, main="Change Over Area", col="lightblue", xlab="Change_Over_Area")
hist_columns <- c(7:13)
for (i in 1:length(hist_columns)) {
  hist(PAC_response_data[, hist_columns[i]], col="lightblue", 
       main = paste("PAC", colnames(PAC_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
hist(PRU_response_data$Change_Over_Area, main="Change Over Area", col="darkolivegreen3", xlab="Change_Over_Area")
for (i in 1:length(hist_columns)) {
  hist(PRU_response_data[, hist_columns[i]], col="darkolivegreen3", 
       main = paste("PRU", colnames(PRU_response_data)[hist_columns[i]], sep=" - "),
       xlab=colnames(data)[hist_columns[i]])
}
main_title <- "Response Data Exploration"
title(main = main_title, outer = TRUE)
#dev.off()
```

Ok so FSC.Events are both pretty right-skewed. CN says log10 transform these.


```{r wet season raw growth data plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
#  ggsave(filename = paste("output/wet_raw_growth_data_response_plot_", sp, ".png", sep = ""),
#         plot = plot, 
#         width = 15, height = 15)
}
```

Now, doing this again but for both seasons combined:
```{r both season raw growth data plots, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = Percent_Change)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```

Same but let's do for FCM data

```{r wet season raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data_seasonal$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data_seasonal %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = log10(FSC.Events_per_cm_2))) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
#  ggsave(filename = paste("output/wet_raw_symbiong_count_response_plot_", sp, ".png", sep = ""),
#         plot = plot, 
#         width = 15, height = 15)
}
```

And again, now for both seasons:
```{r both seasons raw symbiont count plots, echo=FALSE, fig.width=15, fig.height=15}
species <- unique(all_data$Species)
for (sp in species) {
  # Filter data for the current species and only caged
  species_data <- all_data %>%
    filter(Species == sp, Cage_Uncaged == "C")
  
  # Create plot for the current species
  plot <- species_data %>%
    ggplot(aes(x = value, y = log10(FSC.Events_per_cm_2))) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ Parameter, scales = "free_x") +
    ggtitle(paste(sp)) +  #Add species name to the plot title
    theme_bw()
  # Print and save the plot
  print(plot)
  #ggsave(filename = paste("output/raw_growth_data_response_plot_", sp, ".png", sep = ""),
  #       plot = plot, 
  #       width = 15, height = 15)
}
```

```{r}
rus_growth_model <- response_data %>%
  dplyr::filter(Species == "Porites rus")
 
ggplot(rus_growth_model, aes(x = Percent_Change, y = Cage_Uncaged)) +
  geom_boxplot() +
  geom_jitter(width = 0, height = 0.2, alpha = 0.5) #+
#  geom_text(aes(label = Pin_Number))

rus_growth_model <- lme4::lmer(Percent_Change ~ Cage_Uncaged + (1|Genotype), data = rus_growth_model)
Anova(rus_growth_model)
```

```{r}
rus_growth_model <- response_data %>%
  dplyr::filter(Species == "Pocillopora acuta")
 
ggplot(rus_growth_model, aes(x = Percent_Change, y = Cage_Uncaged)) +
  geom_boxplot()+
  geom_jitter(width = 0, height = 0.2, alpha = 0.5) +
  geom_text(aes(label = Pin_Number))

rus_growth_model <- lme4::lmer(Percent_Change ~ Cage_Uncaged + (1|Genotype), data = rus_growth_model)
Anova(rus_growth_model)
```



```{r growth change by symbionts per gm}
symb_growth_model <- response_data %>%
  filter(Cage_Uncaged == "C") %>%
  group_by(Species) %>% 
  do(model = lmer(Percent_Change ~ log10(FSC.Events_per_g_dry_weight) + (1|Genotype), data = .))

# Extract model objects and print the anovas
model_objects <- symb_growth_model$model
anova_tables <- lapply(model_objects, function(model) {
  car::Anova(model)
})
anova_tables
#car::Anova(symb_growth_model$model[[1]]) #less code with small number of models
car::Anova(symb_growth_model$model[[2]])

# Add a new column to specify which Species should have the smoothing line
response_data <- response_data %>%
  mutate(smooth_line = ifelse(Species == "Porites rus", TRUE, FALSE))

# Filter and plot
response_data %>% 
  filter(Cage_Uncaged == "C") %>% 
  ggplot(aes(y = Percent_Change, x = log10(FSC.Events_per_g_dry_weight))) +
  geom_point() +
  geom_smooth(data = . %>% filter(smooth_line), method = "lm") +
  ggtitle("Change in BW by Number of Symbionts") +
#  geom_text(aes(label = Pin_Number)) +
  facet_wrap(~ Species, scales = "free") -> growth_v_symb_plot

growth_v_symb_plot

#ggsave("output/caged_growth_v_symb.png", width = 9, height = 6, plot = growth_v_symb_plot)
#########
```


```{r growth change by symbionts per cm}
symb_growth_model <- response_data %>%
  filter(Cage_Uncaged == "C") %>%
  group_by(Species) %>% 
  do(model = lmer(Percent_Change ~ log10(FSC.Events_per_cm_2) + (1|Genotype), data = .))

# Extract model objects and print the anovas
model_objects <- symb_growth_model$model
anova_tables <- lapply(model_objects, function(model) {
  car::Anova(model)
})
anova_tables
#car::Anova(symb_growth_model$model[[1]]) #less code with small number of models
car::Anova(symb_growth_model$model[[2]])

# Add a new column to specify which Species should have the smoothing line
response_data <- response_data %>%
  mutate(smooth_line = ifelse(Species == "Porites rus", TRUE, FALSE))

# Filter and plot
response_data %>% 
  filter(Cage_Uncaged == "C") %>% 
  ggplot(aes(y = Percent_Change, x = log10(FSC.Events_per_cm_2))) +
  geom_point() +
  geom_smooth(data = . %>% filter(smooth_line), method = "lm", formula = y ~ poly(x, 2)) +
  ggtitle("Change in BW by Number of Symbionts") +
#  geom_text(aes(label = Pin_Number)) +
  facet_wrap(~ Species, scales = "free") -> growth_v_symb_plot

growth_v_symb_plot

#ggsave("output/caged_growth_v_symb.png", width = 9, height = 6, plot = growth_v_symb_plot)
#########
```

OK, I don't want this relationship driven by 14 for rus, which I know had a good B coral and a diseased C coral

```{r}
response_data_no_14C_14B <- response_data %>%
  filter(Species == "Porites rus" & (Cage_Uncaged == "C" | (Cage_Uncaged == "B" & Pin_Number == 14))) %>% 
  filter(!(Pin_Number == 14 & Cage_Uncaged == "C"))

response_data_no_14C <- response_data %>%
  filter(Species == "Porites rus" & 
         (Cage_Uncaged == "C")) %>%
  filter(!(Pin_Number == 14 & Cage_Uncaged == "C"))

response_data_PAC <- response_data %>%
  filter(Species == "Pocillopora acuta")

model = lmer(Percent_Change ~ log10(FSC.Events_per_cm_2) + (1|Genotype), data = response_data_no_14C)
car::Anova(model)

plot1 <- response_data_PAC %>% ggplot(aes(y = Percent_Change, x = log10(FSC.Events_per_cm_2)))+
  geom_point() +
#  geom_smooth(method = "lm", formula = "y~poly(x,2)") +
  ggtitle("Pocillopora acuta")

plot2 <- response_data_no_14C %>% ggplot(aes(y = Percent_Change, x = log10(FSC.Events_per_cm_2)))+
  geom_point() +
#  geom_smooth(method = "lm", formula = "y~poly(x,2)") +
  ggtitle("Porites rus")

grid.arrange(plot1, plot2, ncol = 2)

```




```{r bw change in rus vs acuta}
BW_long <- response_data %>% 
#  filter(Cage_Uncaged == "C") %>%
  filter(!(Species == "Porites rus" & Cage_Uncaged == "A" & Pin_Number == "13")) %>%
  dplyr::select(Species, Percent_Change,Pin_Number, Cage_Uncaged) %>%
  group_by(Species, Pin_Number,Cage_Uncaged)%>%
  summarise(Percent_ChangeM  = mean(Percent_Change, na.rm = TRUE)) %>% 
  pivot_wider(names_from = "Species",
              values_from = "Percent_ChangeM")

PRU_genotype <- response_data %>%
  filter(Species == "Porites rus") %>%
  distinct(Pin_Number, Genotype) %>%
  rename(PRU_Genotype = Genotype)

PAC_genotype <- response_data %>%
  filter(Species == "Pocillopora acuta") %>%
  distinct(Pin_Number, Genotype) %>%
  rename(PAC_Genotype = Genotype)

BW_long <- response_data %>%
  # filter(Cage_Uncaged == "C") %>%
  filter(!(Species == "Porites rus" & Cage_Uncaged == "A" & Pin_Number == "13")) %>%
  dplyr::select(Species, Percent_Change, Pin_Number, Cage_Uncaged) %>%
  group_by(Species, Pin_Number, Cage_Uncaged) %>%
  summarise(Percent_ChangeM = mean(Percent_Change, na.rm = TRUE)) %>%
  pivot_wider(names_from = "Species", values_from = "Percent_ChangeM") %>%
  # Join with genotypes
  left_join(PRU_genotype, by = "Pin_Number") %>%
  left_join(PAC_genotype, by = "Pin_Number")

BW_long$Cage_Uncaged <- as.factor(BW_long$Cage_Uncaged)
BW_long$PRU_Genotype <- as.factor(BW_long$PRU_Genotype)
BW_long$Source_Colony <- as.integer(BW_long$PRU_Genotype)
BW_long$Source_Colony <- as.factor(BW_long$Source_Colony)
```

```{r}
custom_labels <- c("A" = "Uncaged", "B" = "Cage Control", "C" = "Caged")

growth_v_sp_plot_all <- ggplot(data = na.omit(BW_long), aes(y = `Pocillopora acuta`, x = `Porites rus`)) +
  geom_point(aes(shape = Source_Colony, color = Cage_Uncaged)) +
  coord_fixed()+
  theme_linedraw() +
  labs(
    shape = "Source Colony",
    color = "Cage Treatment"  # New legend title for the color scale
  ) +
  scale_x_continuous(breaks = seq(0, max(BW_long$`Pocillopora acuta`, na.rm = TRUE), by = 5)) +
  scale_y_continuous(breaks = seq(0, max(BW_long$`Porites rus`, na.rm = TRUE), by = 5)) +
  scale_color_fish_d(option = "Coryphaena_hippurus", labels = custom_labels) +
  theme(
    axis.title.x = element_text(face = "italic", size = 12),
    axis.title.y = element_text(face = "italic", size = 12),
    plot.title = element_text(hjust = 0.5, size = 12)) +
  annotate("text", x = Inf, y = Inf, 
           label = paste("p =", "0.71", "\nR² =", "0.25"),
           hjust = 1.1, vjust = 1.1, size = 4)

x_range <- range(BW_long$`Porites rus`)
y_range <- range(BW_long$`Pocillopora acuta`)

# Calculate the aspect ratio
aspect_ratio <- diff(x_range) / diff(y_range)

print(growth_v_sp_plot_all)

ggsave(filename = "../output/growth_v_sp_plot_all1.png",height = 5, width = 4 * (1/aspect_ratio), plot = growth_v_sp_plot_all)
```


```{r}
(cor(BW_long$`Pocillopora acuta`, BW_long$`Porites rus`))^2 #how to calculate r and r^2
cor(BW_long$`Pocillopora acuta`, BW_long$`Porites rus`, method = "spearman")


residuals_acuta <- residuals(lmer(BW_long$`Pocillopora acuta` ~ BW_long$`Porites rus` + (1|Cage_Uncaged) + (1|PRU_Genotype) + (1|PAC_Genotype), data = BW_long))
residuals_rus <- residuals(lmer(BW_long$`Porites rus` ~ BW_long$`Pocillopora acuta` + (1|Cage_Uncaged) + (1|PRU_Genotype) + (1|PAC_Genotype), data = BW_long))

correlation <- cor(residuals_acuta, residuals_rus)
correlation
```

#### Look for Outliers:
Buoyant Weight
It looks like that 13 A is really off. We know it broke, so I am removing it and replotted next to it:
```{r BW Outlier Tidy, echo=FALSE}
# BW Plot
BW <- merge(BW, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
BW_no <- BW[!(BW$Placement_Code == "PRU V13 A"), ]
```

Now making a plot to look at outliers:
```{r BW Outlier Plot, echo=FALSE}
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
#dotchart(BW$Change_Over_Area, main = "∆BW", 
#         groups = BW$Cage_Uncaged)
#dotchart(BW_no$Change_Over_Area, main = "∆BW removed erroneous data point", 
#         groups = BW_no$Cage_Uncaged)
layout(1)
```
I removed PRU V13 A placement because this was broken in the field, creating a large negative change in buoyant weight. 

TLE
```{r TLE Plot, echo=FALSE}
TLE <- merge(TLE, meta[, c("Placement_Code", "Cage_Uncaged")], by = "Placement_Code")
layout(matrix(c(1, 2, 3, 4), 2, 2, byrow = TRUE)) # Setting up the layout
#dotchart(na.omit(TLE$Max_L_AVG), main = "Max L", groups = na.omit(TLE$Cage_Uncaged)) # Side Max L Plot
#dotchart(TLE$Max_L_Basal, main = "Max L Basal", groups = na.omit(TLE$Cage_Uncaged)) # Basal Max L plot
#dotchart(TLE$Area_AVG, main = "Area", groups = na.omit(TLE$Cage_Uncaged)) # Area plot
#dotchart(TLE$Area_Basal, main = "Basal Area", groups = na.omit(TLE$Cage_Uncaged)) # Basal Area plot
layout(1) # Reset the layout
#title("TLE Outlier Dotchart", line = -1, outer = TRUE, cex.main = 1.5)
```

### Symbionts

```{r Sym plot, echo=FALSE}
FCM <- FCM[complete.cases(FCM[, "Cage_Uncaged"]), ] #Remove T0 samples by only including rows where the "Cage_Uncaged" exists
FCM$Cage_Uncaged <- as.factor(FCM$Cage_Uncaged) 
FCM_no <- FCM[!(FCM$Placement_Code == "PRU V17 C"), ]
par(mar = c(2.5, 1, 1, 1) + 0.1)
layout(matrix(1:2, nrow = 2), heights = c(3, 3))
dotchart(log10(FCM$FSC.Events_per_cm_2), main = "Symbionts", groups = FCM$Cage_Uncaged)
dotchart(log10(FCM_no$FSC.Events_per_cm_2), main = "Symbionts without erroneous point", groups = FCM_no$Cage_Uncaged)
layout(1)
```
Jess will rerun PRU 49 (PRU V17 C) with her samples - I likely added twice as much as needed to the sample, but for now I will omit in my analyses (using dataframe FCM_no for "no outliers")

```{r FCM Violin Plot}
FCM_violin <- ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_cm_2), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
#  geom_text(aes(label = Pin_Number)) +
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1)+
  facet_wrap(~Species)+
  labs(title = expression(paste("Symbionts per ", cm^2, " by caging treatment")))+
  theme(plot.title = element_text(hjust = 0.5))

FCM_violin
#ggsave("output/FCM_violin.png", width = 9, height = 6, plot = FCM_violin)
```

same but other normalization:

```{r FCM Violin Plot Dry Weight}
ggplot(FCM, aes(x = Cage_Uncaged, y = log10(FSC.Events_per_g_dry_weight), fill = Species)) +
  geom_violin() +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) +  # Adding jitter
  scale_fill_fish_d(option = "Acanthurus_olivaceus", alpha = 0.4, begin = 0.3, end = 1) +
  facet_wrap(~Species)+
  labs(title = "Symbionts per g dry weight by caging treatment")+
  theme(plot.title = element_text(hjust = 0.5))
```

