---
title: "gllvm"
author: "Callie Stephenson"
date: "2025-07-03"
output: html_document
---
```{r}
library(mvabund)
library(gllvm)
library(dplyr)
library(tidyr)
library(ggcorrplot)
```

# Set-up
## Explanatory Data
```{r explanatory data}
setwd("/Users/calliestephenson/Documents/GitHub/MooreaCoralFish_CS")

explan<- read.csv("data/explanatory_all_wide.csv") %>% 
  mutate(Pin = as.numeric(sub("V", "", CowTagID))) %>% 
  filter(!Pin == "13") %>% 
  filter(!Pin == "2")

db <- read.csv("data/DB_Full_Metadata.csv") %>% 
  mutate_all(~ifelse(is.na(.), 0, .))#Rugosity data pulled from Danielle Barnas' measurements
# turb <- read.csv("data/Turb_NC.csv")

explan <- explan %>% 
  left_join(db[,c("CowTagID", "meanRugosity")])
```

### Correlations to choose variables:
#### salinity
```{r}
salinity_data <- explan %>%
  select(contains("Salinity"), dist_to_seep_m)

cor_matrix <- cor(salinity_data, use = "pairwise.complete.obs")

ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)
```

#### silicate
```{r}
Silicate_data <- explan %>%
  select(contains("Silicate"), dist_to_seep_m)

cor_matrix <- cor(Silicate_data, use = "pairwise.complete.obs")

ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)
```

Really, none of these are great. But I want to include salinity, because it is something the fish can feel and respond to directly. 

#### spatial covarince
```{r}
# spatial_corr <- explan_model %>%
#   select(-Pin, -scale_sqrt_dist_seep_m, -scale_log_shore_dist_m)
# cor_matrix <- cor(spatial_corr, use = "pairwise.complete.obs")
# ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower",
#           lab = TRUE)
```


### Look at skew

```{r}
par(mfrow=c(2,2))
hist(explan$dist_to_seep_m) #left skew
hist(explan$shore_dist_m) #right skew
hist(explan$N_percent) #fine
hist(explan$Low_Tide_Mean_Salinity, breaks = 9) #fine?
```
#### Transform and scale
dist_to_seep_m
```{r}
par(mfrow = c(1,2))
hist(explan$dist_to_seep_m) #left skew
hist(sqrt(explan$dist_to_seep_m)) #less skew
explan$scale_sqrt_dist_seep_m <- scale(sqrt(explan$dist_to_seep_m))
```

shore_dist_m
```{r}
par(mfrow = c(1,2))
hist(explan$shore_dist_m) #left skew
hist(log(explan$shore_dist_m)) #less skew
explan$scale_log_shore_dist_m <- scale(log(explan$dist_to_seep_m))
```

N_percent
```{r}
par(mfrow = c(1,2))
hist(explan$N_percent) #no skew

explan$scale_N_percent <- scale((explan$N_percent))
```

Low_Tide_Mean_Salinity
```{r}
hist(explan$Low_Tide_Mean_Salinity, breaks = 9) #fine?
hist(scale(explan$Low_Tide_Mean_Salinity), breaks = 9) #not fine?
hist(log(explan$Low_Tide_Mean_Salinity), breaks = 9) #better?
hist(scale(log(explan$Low_Tide_Mean_Salinity)), breaks = 9) #fine?

explan$scale_log_ltm_salinity <- scale(log(explan$Low_Tide_Mean_Salinity))
```

meanRugosity
```{r}
par(mfrow = c(1,1))
hist(explan$meanRugosity) #no skew

explan$scale_meanRugosity <- scale((explan$meanRugosity))
```
### finish explan
```{r}
explan_variables <- explan %>% 
  select(Pin, lat, lon, scale_sqrt_dist_seep_m, scale_log_shore_dist_m, scale_N_percent, scale_log_ltm_salinity,scale_meanRugosity)

rm(db, explan)
```

```{r}
predictor_data <- explan_variables[,c(4:8)]

cor_matrix <- cor(predictor_data, use = "pairwise.complete.obs")

ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)
```

## Response Data
```{r}
fish_tidy <- read.csv("data/fish_tidy.csv")
#make it wide
wide_maxn_df <- pivot_wider(fish_tidy, 
                             id_cols = c("Pin", "On.Adjacent", "Block"),
                             names_from = Species, 
                             values_from = MaxN) %>% 
  mutate_all(~ifelse(is.na(.), 0, .))

# explan_fish <- left_join(wide_maxn_df, 
#                            explan[, c("Pin", "del15N", "N_percent", "shore_dist_m", "dist_to_seep_m", "Low_Tide_Mean_Salinity")], 
#                            by = "Pin")
# 
# predictors <- c("dist_to_seep_m", "shore_dist_m", "N_percent", "Low_Tide_Mean_Salinity")
# explan_fish <- explan_fish %>%
#   mutate(across(all_of(predictors),
#                 ~ as.numeric(scale(.)),
#                 .names = "{.col}_scale"))
```


```{r}
# Pull just species columns (exclude metadata columns)
meta_cols <- c("Pin", "On.Adjacent", "Block")

fish_matrix <- wide_maxn_df %>%
  select(-all_of(meta_cols)) %>%
  as.matrix()

# # Pull out scaled explanatory columns
# explan_vars <- explan_fish %>%
#   select(dist_to_seep_m_scale,
#          shore_dist_m_scale,
#          N_percent_scale,
#          Low_Tide_Mean_Salinity_scale,
#          Pin) %>% 
#   mutate(Pin = as.factor(Pin))
# 
# # One df
# gllvm_data <- explan_fish %>%
#   select(Pin,
#          dist_to_seep_m_scale,
#          shore_dist_m_scale,
#          N_percent_scale,
#          Low_Tide_Mean_Salinity_scale)
```

### Explanatory matrix 
```{r}
explan_model <- explan_variables %>% 
  left_join(wide_maxn_df[,c("Pin")]) %>% 
  mutate(Pin = as.factor(Pin))
```


# GLLVM Models

```{r}
# Model without predictors:
y <- fish_matrix
no_predictors <- gllvm(y, family = "negative.binomial")
summary(no_predictors)
coef(no_predictors)
plot(no_predictors)
```

## Including environmental variables
```{r}
X <- as.matrix(explan_model %>% select(-Pin, -lat, -lon, -scale_meanRugosity))
# Model where environmental variables, that is, all variables in X are included:
env <- gllvm(y, X, family = "negative.binomial")
```

One time this gave a lot of warnings, but now it is not. My understanding is the warnings it did give me don't matter

Which fits, but because there isn't the 'formula' argument, it doesn't know how to interpret model
```{r}
# env_filt <- gllvm(
#   y,
#   X,
#   family = "negative.binomial",
#   num.lv = 0,
#   formula = ~ scale_sqrt_dist_seep_m + scale_log_shore_dist_m + scale_N_percent + scale_log_ltm_salinity,
#   seed = 1234
# )
# 
# gllvm::coefplot(env_filt, cex.ylab = 0.7, mar = c(4, 9, 2, 1), mfrow=c(1,1))
```

## Compare AIC values across different num.lv
```{r}
# criteria <- NULL
# for(i in 1:5){
#   fiti <- gllvm(y, X, family = "negative.binomial", num.lv = i, sd.errors = FALSE,
#                 formula = ~ scale_sqrt_dist_seep_m + scale_log_shore_dist_m + scale_N_percent + scale_log_ltm_salinity, seed = 1234)
#   criteria[i] <- summary(fiti)$AICc
#   names(criteria)[i] = i
# }
# 
# criteria
```
Based on these tools a model with NB distributed responses and 1 latent variable was chosen.
```{r}
# env_filt <- gllvm(
#   y,
#   X,
#   family = "negative.binomial",
#   num.lv = 1,
#   formula = ~ scale_sqrt_dist_seep_m + scale_log_shore_dist_m + scale_N_percent + scale_log_ltm_salinity,
#   seed = 1234
# )
# 
# gllvm::coefplot(env_filt, cex.ylab = 0.7, mar = c(4, 9, 2, 1), mfrow=c(1,1))
```

HOWEVER, this was just following the examples (which worked! Great), now we need to include random effects:

## Include Random Effects

No latent variables, but keep the random effect:
```{r}
# y <- fish_matrix # species abundance
# X <- as.matrix(explan_model %>% select(!Pin, !lat, !lon, !scale_meanRugosity)) #keeping all the explanatory variables, but removing survey design
# 
# 
# gllvm_nolv <- gllvm(
#   y       = y,
#   X       = X,
#   formula = ~(1 | Pin), # random intercept for each Pin
#   family  = "negative.binomial",
#   num.lv  = 0, #we'll play with this if it fits
#   data    = explan_model # for the formula to find Pin
# )
# 
# summary(gllvm_nolv)
# coef(gllvm_nolv)
# plot(gllvm_nolv)
```

Now, trying with studyDesign argument

```{r}
# y <- fish_matrix # species abundance
# X <- as.matrix(explan_model %>% select(-Pin, -lat, -lon, -scale_log_shore_dist_m)) #keeping all the explanatory variables, but removing survey design
# study_design <- data.frame(Pin = explan_model$Pin)
# env_studydesign <- gllvm(
#   y = y,
#   X = X,
#   formula = ~ scale_sqrt_dist_seep_m + 
#               scale_N_percent + scale_log_ltm_salinity + scale_meanRugosity,
#   family = "negative.binomial",
#   studyDesign = study_design,
#   row.eff = ~(1|Pin),
#   seed = 1234
# )
```

Not showing it is overfitted, but some of the effects are... weird
```{r}
# env_studydesign
# 
# #plot the effects:
# gllvm::coefplot(env_studydesign, cex.ylab = 0.7)
```

need to reset the par
```{r}
# dev.off()
```

Variance Partitioning in this model:
```{r}
# VP1 <- VP(env_studydesign)
# plot(VP1, args.legend = list(cex=0.7), col=hcl.colors(7, "Roma"))
```

The latent variables are doing ... nothing. So I'll remove them but keep the study design:

```{r}
# y <- fish_matrix # species abundance
# X <- explan_model %>% 
#   select(scale_sqrt_dist_seep_m, 
#          #scale_log_shore_dist_m, 
#          scale_N_percent, 
#          scale_log_ltm_salinity,
#          scale_meanRugosity) #keeping all the explanatory variables, but removing pin as survey design
# study_design <- data.frame(Pin = explan_model$Pin)
# env_studydesign_nolv <- gllvm(
#   y = y,
#   X = X,
#   formula = ~ scale_sqrt_dist_seep_m + 
#               scale_N_percent + scale_log_ltm_salinity + scale_meanRugosity,
#   family = "negative.binomial",
#   studyDesign = study_design,
#   row.eff = ~(1|Pin),
#   num.lv = 0,
#   seed = 1234
# )
```

Is it better:
```{r}
# anova(env_studydesign, env_studydesign_nolv)
```

Coefficient plot:
```{r}
# gllvm::coefplot(env_studydesign_nolv, cex.ylab = 0.7, mar = c(4, 9, 2, 1), mfrow=c(3,2))
```

```{r}
# par(mfrow=c(1,1))
# VP1 <- VP(env_studydesign_nolv)
# plot(VP1, args.legend = list(cex=0.7), col=hcl.colors(5, "Roma"))
```


Kind of obvious, but these coefficients are WILD and have to be wrong??
```{r}
# #make into dataframe
# model_coef<-coef(env_studydesign)
# est_df<-data.frame(model_coef$Intercept)
# est_df2<-data.frame(model_coef$Xcoef) 
# est_df3<-merge(est_df, est_df2, by = 0)
```

Come here

## Subset species model:
### subset sp
```{r}
species_presence <- colSums(fish_matrix > 0)

# Quick histogram
hist(species_presence, main = "Species prevalence (non-zero counts)", xlab = "# of sites")
```

```{r}
#species_present <- colSums(fish_matrix > 0) >= round(0.05 * nrow(fish_matrix)) #5% of samples aka 7 samples
species_present <- colSums(fish_matrix > 0) >= 6
fish_matrix_filtered <- fish_matrix[, species_present]
y1 <- fish_matrix_filtered
study_design <- data.frame(Pin = explan_model$Pin)
X <- explan_model %>% 
  select(scale_sqrt_dist_seep_m, 
         #scale_log_shore_dist_m, 
         scale_N_percent, 
         scale_log_ltm_salinity,
         scale_meanRugosity) #keeping all the explanatory variables, but removing pin as survey design
```

### Model
model, but how many latent variables
```{r}
criteria <- numeric(6)  # to store AICc values for num.lv = 0 to 5

for (i in 0:5) {
  cat("Fitting model with", i, "latent variables...\n")
  #Fit model
  fit_i <- gllvm(
    y = y1,
    X = X,
    formula = ~ scale_sqrt_dist_seep_m + 
                scale_N_percent + scale_log_ltm_salinity + 
                scale_meanRugosity,
    family = "negative.binomial",
    studyDesign = study_design,
    row.eff = ~(1 | Pin),
    num.lv = i,
    seed = 1234
  )
  
  # Take the AICc of that model
  criteria[i + 1] <- summary(fit_i)$AICc
  names(criteria)[i + 1] <- paste0("num.lv = ", i)
}

criteria
names(criteria)[which.min(criteria)]
```
Suggests the best model has 0 latent variables

```{r}
filtered_env_studydesign <- gllvm(
  y = y1,
  X = X,
  formula = ~ scale_sqrt_dist_seep_m + 
              scale_N_percent + scale_log_ltm_salinity + scale_meanRugosity,
  family = "negative.binomial",
  studyDesign = study_design,
  row.eff = ~(1|Pin),
  num.lv = 0,
  seed = 1234
)
```

```{r}
summary <- summary(filtered_env_studydesign)
coef(filtered_env_studydesign)
plot(filtered_env_studydesign)

coefs_model <- as.data.frame(summary$Coef.tableX) %>%
  tibble::rownames_to_column(var = "VariableSpecies") %>%
  separate(VariableSpecies, into = c("Variable", "Species"), sep = ":", extra = "merge")

sig_coefs_model <- coefs_model %>% 
  filter(`Pr(>|z|)`< 0.05)
```

#### Plots:
Coeffiient plot: first check how big the coefficients are
```{r}
#make into dataframe
model_coef<-coef(filtered_env_studydesign)
est_df<-data.frame(model_coef$Intercept)
est_df2<-data.frame(model_coef$Xcoef) #this one is the best one
est_df3<-merge(est_df, est_df2, by = 0)
```

```{r}
gllvm::coefplot(filtered_env_studydesign, cex.ylab = 0.7, mar = c(4, 9, 2, 1), mfrow=c(2,2), xlim.list= list(c(-2.5,2.5), c(-2.5,2.5), c(-2.5,2.5),c(-2.5,2.5)))

# gllvm::coefplot(filtered_env_studydesign, cex.ylab = 0.7, mar = c(4, 9, 2, 1), mfrow=c(2,2), xlim.list= list(c(-2.5,2.5), c(-2.5,2.5), c(-2.5,2.5),c(-2.5,2.5)), show = "all")
```
Variance Partitioning Plot:
```{r}
par(mfrow=c(1,1))
VP1 <- VP(filtered_env_studydesign)
plot(VP1, args.legend = list(cex=0.7), col=hcl.colors(5, "Roma"))
```

#### dropping N_percent
```{r}
X_reduced <- X %>%
  select(-scale_N_percent)
```
reduce model
```{r}
reduced_model <- gllvm(
  y = y1,
  X = X_reduced,
  formula = ~ scale_sqrt_dist_seep_m + scale_log_ltm_salinity + scale_meanRugosity,
  family = "negative.binomial",
  studyDesign = study_design,
  row.eff = ~(1 | Pin),
  num.lv = 0,
  seed = 1234
)
```
compare AIC
```{r}
AIC(filtered_env_studydesign)
AIC(reduced_model)
```
It is better
Check num latent variables
```{r}
criteria_reduced <- numeric(6)

for (i in 0:5) {
  cat("Fitting model with", i, "latent variables (reduced predictors)...\n")
  
  fit_i <- gllvm(
    y = y1,
    X = X_reduced,   # your reduced set of predictors
    formula = ~ scale_sqrt_dist_seep_m + scale_log_ltm_salinity + scale_meanRugosity,  # adjusted formula
    family = "negative.binomial",
    studyDesign = study_design,
    row.eff = ~(1 | Pin),
    num.lv = i,
    seed = 1234
  )
  
  criteria_reduced[i + 1] <- summary(fit_i)$AICc
  names(criteria_reduced)[i + 1] <- paste0("num.lv = ", i)
}

criteria_reduced
names(criteria_reduced)[which.min(criteria_reduced)]
```
still 0 latent variables

```{r}
gllvm::coefplot(reduced_model, cex.ylab = 0.7, mar = c(4, 9, 2, 1), mfrow=c(2,2), xlim.list= list(c(-2.5,2.5), c(-2.5,2.5), c(-2.5,2.5),c(-2.5,2.5)))
```

```{r}
par(mfrow=c(1,1))
VP1 <- VP(reduced_model)
plot(VP1, args.legend = list(cex=0.7), col=hcl.colors(5, "Roma"))
```



```{r}
summary <- summary(reduced_model)
coef(reduced_model)
plot(reduced_model)

coefs_model <- as.data.frame(summary$Coef.tableX) %>%
  tibble::rownames_to_column(var = "VariableSpecies") %>%
  separate(VariableSpecies, into = c("Variable", "Species"), sep = ":", extra = "merge")

sig_coefs_model <- coefs_model %>% 
  filter(`Pr(>|z|)`< 0.05)
```



# Spatial autocorrelation with lat / lon
For my next trick, I'd like to levee this:
"If one would like to set eg. spatial correlation structure for the site level random effects, it could be defined for instance by row.eff = ~corExp(1|site) and spatial coordinates given to dist as a matrix."

And then removing the distance to seep and distance to shore variables, only looking at salinity and N_percent

```{r}
coords <- explan_model %>%
  select(Pin, lat, lon) 
# Convert to a matrix of Euclidean distances (or haversine if large-scale)
spatial_dist <- as.matrix(dist(cbind(coords$lat, coords$lon)))
```

but, i don't think ^ works because of repeated measures, so
```{r}
coords <- explan_model %>%
  select(Pin, lat, lon) %>%
  distinct(Pin, .keep_all = TRUE) %>%
  arrange(Pin)  # order needs to be the same as Pin because Pin will be removed for distance

# Step 2: distance matrix for site-level spatial random effect
spatial_dist <- as.matrix(dist(cbind(coords$lat, coords$lon)))
```


```{r}
y <- fish_matrix # species abundance
X <- explan_model %>% 
  select(scale_sqrt_dist_seep_m, 
         scale_log_shore_dist_m, 
         scale_N_percent, 
         scale_log_ltm_salinity,
         meanRugosity) #keeping all the explanatory variables, but removing pin as survey design
study_design <- data.frame(Pin = factor(explan_model$Pin, levels = coords$Pin)) #added in levels to make sure it matches coords

# #Spatial GLM model
# gllvm_spatial <- gllvm(
#   y = y,
#   X = X,
#   formula = ~ scale_sqrt_dist_seep_m + scale_log_shore_dist_m +
#               scale_N_percent + scale_log_ltm_salinity + meanRugosity,
#   family = "negative.binomial",
#   studyDesign = study_design,
#   row.eff = ~ corExp(1 | Pin),
#   dist = spatial_dist,
#   seed = 1234,
#   num.lv = 0 #no need for latent variables after checking VP
# )

#Spatial GLM model 2
gllvm_spatial <- gllvm(
  y = y,
  X = X,
  formula = ~ scale_N_percent + scale_log_ltm_salinity + meanRugosity,
  family = "negative.binomial",
  studyDesign = study_design,
  row.eff = ~ corExp(1 | Pin),
  dist = spatial_dist,
  seed = 1234,
  num.lv = 0 #no need for latent variables after checking VP
)
```

Warning message:
In doTryCatch(return(expr), name, parentenv, handler) :
  Determinant of the variance-covariance matix is zero. Please double check your model for e.g. overfitting or lack of convergence. 
  
```{r}
gllvm_spatial

# sds for random row effects and Correlation for consecutive years
gllvm_spatial$params$sigma

#coefficients
gllvm::coefplot(gllvm_spatial, cex.ylab = 0.7, mar = c(4, 9, 2, 1), xlim.list= list(NULL, NULL, c(-100,100)))
```

Variance Partitioning in this model:
```{r}
par(mfrow=c(1,1))
VPspatial <- VP(gllvm_spatial)
plot(VPspatial, args.legend = list(cex=0.7), col=hcl.colors(5, "Roma"))
```


