---
title: "MVABUND"
author: "Callie Stephenson"
date: "2025-07-09"
output: html_document
---
```{r}
library(dplyr)
library(tidyr)
library(mvabund)
library(skimr)
library(ggcorrplot)
library(reshape2)
```

# Set-up
## Explanatory Data
```{r explanatory data}
setwd("/Users/calliestephenson/Documents/GitHub/MooreaCoralFish_CS")

explan<- read.csv("data/explanatory_all_wide.csv") %>% 
  mutate(Pin = as.numeric(sub("V", "", CowTagID))) %>% 
  filter(!Pin == "13") %>% 
  filter(!Pin == "2") %>% 
  mutate(Pin = as.factor(Pin))

db <- read.csv("data/DB_Full_Metadata.csv") %>% 
  mutate_all(~ifelse(is.na(.), 0, .))#Rugosity data pulled from Danielle Barnas' measurements
# turb <- read.csv("data/Turb_NC.csv")

explan <- explan %>% 
  left_join(db[,c("CowTagID", "meanRugosity")])

rm(db)

# skimr::skim(explan) # Great function to get an overview of the data
```

### Correlations to choose variables:
#### salinity
```{r}
salinity_data <- explan %>%
  select(contains("Salinity"), dist_to_seep_m)

cor_matrix <- cor(salinity_data, use = "pairwise.complete.obs")

ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)

rm(salinity_data, cor_matrix)
```

#### silicate
```{r}
Silicate_data <- explan %>%
  select(contains("Silicate"), dist_to_seep_m)

cor_matrix <- cor(Silicate_data, use = "pairwise.complete.obs")

ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower",
          lab = TRUE)
rm(Silicate_data, cor_matrix)
```

Really, none of these are great. But I want to include salinity, because it is something the fish can feel and respond to directly. 

#### spatial covarince
```{r}
# spatial_corr <- explan_model %>%
#   select(-Pin, -scale_sqrt_dist_seep_m, -scale_log_shore_dist_m)
# cor_matrix <- cor(spatial_corr, use = "pairwise.complete.obs")
# ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower",
#           lab = TRUE)

# rm(spatial_corr, cor_matrix)
```


### Look at skew

```{r}
par(mfrow=c(3,2))
hist(explan$dist_to_seep_m) #left skew
hist(explan$shore_dist_m) #right skew
hist(explan$N_percent) #fine
hist(explan$Low_Tide_Mean_Salinity, breaks = 9) #fine?
hist(explan$meanRugosity, breaks = 9) #fine?
```

#### Transform and scale
dist_to_seep_m
```{r}
par(mfrow = c(1,2))
hist(explan$dist_to_seep_m) #left skew
hist(sqrt(explan$dist_to_seep_m)) #less skew
explan$scale_sqrt_dist_seep_m <- scale(sqrt(explan$dist_to_seep_m))
```

shore_dist_m
```{r}
par(mfrow = c(1,2))
hist(explan$shore_dist_m) #left skew
hist(log(explan$shore_dist_m)) #less skew
explan$scale_log_shore_dist_m <- scale(log(explan$dist_to_seep_m))
```

N_percent
```{r}
par(mfrow = c(1,2))
hist(explan$N_percent) #no skew

explan$scale_N_percent <- scale((explan$N_percent))
```

Low_Tide_Mean_Salinity
```{r}
hist(explan$Low_Tide_Mean_Salinity, breaks = 9) #fine?
hist(scale(explan$Low_Tide_Mean_Salinity), breaks = 9) #not fine?
hist(log(explan$Low_Tide_Mean_Salinity), breaks = 9) #better?
hist(scale(log(explan$Low_Tide_Mean_Salinity)), breaks = 9) #fine?

explan$scale_log_ltm_salinity <- scale(log(explan$Low_Tide_Mean_Salinity))
```

```{r}
par(mfrow = c(1,1))
hist(explan$meanRugosity) #no skew

explan$scale_meanRugosity <- scale((explan$meanRugosity))
```

### finish explan
```{r}
explan_variables <- explan %>% 
  select(Pin, lat, lon, scale_sqrt_dist_seep_m, scale_log_shore_dist_m, scale_N_percent, scale_log_ltm_salinity, scale_meanRugosity)
```

## Response Data
```{r}
fish_tidy <- read.csv("data/fish_tidy.csv")
#make it wide
wide_maxn_df <- pivot_wider(fish_tidy, 
                             id_cols = c("Pin", "On.Adjacent", "Block"),
                             names_from = Species, 
                             values_from = MaxN) %>% 
  mutate_all(~ifelse(is.na(.), 0, .)) %>% 
  mutate(Pin = as.factor(Pin))

# explan_fish <- left_join(wide_maxn_df, 
#                            explan[, c("Pin", "del15N", "N_percent", "shore_dist_m", "dist_to_seep_m", "Low_Tide_Mean_Salinity")], 
#                            by = "Pin")
# 
# predictors <- c("dist_to_seep_m", "shore_dist_m", "N_percent", "Low_Tide_Mean_Salinity")
# explan_fish <- explan_fish %>%
#   mutate(across(all_of(predictors),
#                 ~ as.numeric(scale(.)),
#                 .names = "{.col}_scale"))
```

# MVABUND

## model data
make an mvabund object
```{r}
predictors <- c("scale_N_percent", "scale_log_ltm_salinity", "scale_sqrt_dist_seep_m", "scale_log_shore_dist_m", "scale_meanRugosity")

explan_fish <- left_join(wide_maxn_df,
                           explan %>% select(all_of(predictors), "Pin"),
                           by = "Pin")

meta_cols <- c("Pin","On.Adjacent","Block")
fish_mv   <- mvabund(wide_maxn_df %>%  select(-all_of(meta_cols)))
```

## Model
Null
```{r}
model_mv_null <- manyglm(fish_mv ~ 1, 
                    family = "negative.binomial", 
                    data = explan_fish)
```


Predictors
```{r}
model_mv <- manyglm(fish_mv ~ scale_log_shore_dist_m + scale_sqrt_dist_seep_m + scale_N_percent + scale_log_ltm_salinity + scale_meanRugosity, 
                    family = "negative.binomial", 
                    data = explan_fish)
```

```{r}
model_mv_Pin <- manyglm(fish_mv ~ scale_log_shore_dist_m + scale_sqrt_dist_seep_m + scale_N_percent + scale_log_ltm_salinity + scale_meanRugosity + Pin, 
                    family = "negative.binomial", 
                    data = explan_fish)
```

### Check model assumptions:
```{r}
plot(model_mv)
```
### Check mean-variance:
```{r}
#meanvar.plot(fish_mv~, col = treatment)
```

### ANOVA
```{r}
set.seed(123)  
anova_full <- mvabund::anova.manyglm(
  model_mv,
  resamp = "montecarlo",             # or "pit.trap"
  block  = explan_fish$Pin,                  # repeat‑measure structure
  nBoot  = 999,                      # increase for more precision, takes time (~28 min 35 sec)
  test   = "LR",                     # likelihood‑ratio test
  p.uni  = "adjusted"                # univariate p‑values too
)
```

If you do not have a specific hypothesis in mind that you want to test, and are instead interested in which model terms are statistically significant, then the summary() function will come in handy. However results aren’t quite as trustworthy as for anova(). The reason is that re-samples are taken under the alternative hypothesis for summary(), where there is a greater chance of fitted values being zero, especially for rarer taxa (e.g. if there is a treatment combination in which a taxon is never present). Abundances don’t re-sample well if their predicted mean is zero.

### Model Summary
```{r}
summary(model_mv)
```

### Total Deviance
```{r}
# Sum deviance across all species for null and full model
total_dev_null <- sum(deviance(model_mv_null))
total_dev_full <- sum(deviance(model_mv))

total_dev_explained <- total_dev_null - total_dev_full
total_dev_explained

# Predictor deviances from anova output
predictor_dev <- c(447.6, 450.2, 172.3, 177.8)  

# Calculate proportion of explained deviance by each predictor
prop_explained <- predictor_dev / total_dev_explained

names(prop_explained) <- c("Distance shore", "Distance seep", "Nitrogen", "Salinity")

prop_explained

```

```{r}
total_dev_null
total_dev_full
total_dev_explained
```

### Species-Level Coefficient Heatmap
```{r}
# Extract coefficients
coef_mat <- coef(model_mv)  # matrix of coefficients [species x predictors]

# Turn into tidy format
coef_df <- melt(coef_mat, varnames = c("Predictor", "Species"), value.name = "Coefficient")

# Plot heatmap
library(viridis)
mid_color <- "gray90"  # neutral center
low_color <- viridis(4, option = "D")[1]   # dark purple-blue (negative)
high_color <- viridis(5, option = "D")[5]  # bright yellow-green (positive)

ggplot(coef_df, aes(x = Predictor, y = Species, fill = Coefficient)) +
  geom_tile() +
  scale_fill_gradient2(low = low_color, mid = mid_color, high = high_color, midpoint = 0) +
  theme_minimal(base_size = 12) +
  labs(title = "Species-Level Coefficients by Predictor")
  
```

## Model without 14
```{r}
predictors <- c("scale_N_percent", "scale_log_ltm_salinity", "scale_sqrt_dist_seep_m", "scale_log_shore_dist_m", "scale_meanRugosity")
wide_maxn_df_no14 <- wide_maxn_df %>% 
  filter(!Pin == "14") %>% 
  mutate(Pin = droplevels(as.factor(Pin)))

explan_fish_no14 <- left_join(wide_maxn_df_no14,
                           explan %>% select(all_of(predictors), "Pin"),
                           by = "Pin")%>% 
  mutate(Pin = droplevels(as.factor(Pin)))

meta_cols <- c("Pin","On.Adjacent","Block")
fish_mv_no14   <- mvabund(wide_maxn_df_no14 %>%  select(-all_of(meta_cols)))
```
Null
```{r}
model_mv_null_no14 <- manyglm(fish_mv_no14 ~ 1, 
                    family = "negative.binomial", 
                    data = explan_fish_no14)
```


Predictors
```{r}
model_mv_no14 <- manyglm(fish_mv_no14 ~ scale_sqrt_dist_seep_m + scale_N_percent + scale_log_ltm_salinity + scale_meanRugosity, 
                    family = "negative.binomial", 
                    data = explan_fish_no14)

model_mv_no14_Pin <- manyglm(fish_mv_no14 ~ scale_sqrt_dist_seep_m + scale_N_percent + scale_log_ltm_salinity + scale_meanRugosity + Pin, 
                    family = "negative.binomial", 
                    data = explan_fish_no14)

mvabund::anova.manyglm(model_mv_no14_Pin)
```

### Check model assumptions:
```{r}
plot(model_mv_no14)
```
### Check mean-variance:
```{r}
#meanvar.plot(fish_mv~, col = treatment)
```

### ANOVA
```{r}
set.seed(123)  
anova_full <- mvabund::anova.manyglm(
  model_mv_no14,
  resamp = "montecarlo",             # or "pit.trap"
  block  = explan_fish_no14$Pin,                  # repeat‑measure structure
  nBoot  = 999,                      # increase for more precision, takes time (~28 min 35 sec)
  test   = "LR",                     # likelihood‑ratio test
  p.uni  = "adjusted"                # univariate p‑values too
)

anova_full
```


### Model Summary
```{r}
summary(model_mv_no14)
```

### Total Deviance
```{r}
# Sum deviance across all species for null and full model
total_dev_null <- sum(deviance(model_mv_null))
total_dev_full <- sum(deviance(model_mv_no14))

total_dev_explained <- total_dev_null - total_dev_full
total_dev_explained

# Predictor deviances from anova output
# predictor_dev <- c(447.6, 450.2, 172.3, 177.8)  
# 
# # Calculate proportion of explained deviance by each predictor
# prop_explained <- predictor_dev / total_dev_explained
# 
# names(prop_explained) <- c("Distance shore", "Distance seep", "Nitrogen", "Salinity")
# 
# prop_explained

```
